---
title: "Cleaned Cupid Cohort Analysis"
author: "Tallulah Story"
date: "2026-02-08"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

This code was authored originally by LTS from 6/2/2025-8/1/2025. The "cleaned" version of this code used for publication anlysis and visualization was completed from 2/8/2026-2/20/2026. Both versions are available via ... Data is also available via ...

## Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Packages Used
library(tidyverse)
library(lubridate)        #Formatting dates and times (tidyverse add-on)
library(ggplot2)          #plotting vehicle
library(patchwork)        #Combining ggplots
library(ggtext)           #Additional plot aesthetics
library(ggthemes)         #Additional ggplot themes
library(ggpubr)           #Plots containing statistical information
library(pastecs)          #For descriptive statistics relating to time-series data
library(car)              #Testing assumptions for statistical analysis
library(rstatix)          #Pipe-friendly statistical analysis/post-hoc testing
library(mgcv)             #Mixed GAMM computational vehicle
library(MASS)             #For mvnorm in simulated GAM
library(marginaleffects)  #Compute and plot_predictions from models
library(fastDummies)      #Dummy variable coding for SLR/MLR

```

## Data import

```{r data import}
## Fitness data set (experiments from 2/13/25 to 2/25/25)
fitness <- read.csv("Fitness Test.csv")

## Growth data (from measurements and counts taken 12/31/24 to 1/2/25)
growth <- read.csv("Growth Cruves Dec to Feb.csv") #wide format data
growth_combined <- read.csv("growth curves 2.csv") #long format data

## Settlement data (from settlement counts 3/26/24 to 7/9/24)
settlement_raw_wide <- read.csv("Settlement raw counts wide.csv")
settlement_raw_long <- read.csv ("Settlement raw count long.csv")
survivorship <- read.csv("Settlement total counts.csv")
settlement_cohort <- read.csv("Settlement cupid cohort.csv")
juvenile_settlement <- read.csv("Juvenile settlment survival.csv")

## Specimen Measurements (from Measurements done on 7/7/2025 to 7/9/2025)
specimens <- read.csv("Pycnopodia specimen measurements.csv")
specimens_alive <- read.csv("Pycnopodia specimen measurements - Live Stars.csv")

## Larval growth data collected from ImageJ measurements taken of larvae DPF 4-47
larval_growth <- read.csv("Larval Growth Curves.csv")
```

## Data Cleaning, fixing date and time data with lubridate

```{r lubridate}
#growth data
growth$Measurement_date_1 <- mdy(growth$Measurement_date_1) #jan sampling
class(growth$Measurement_date_1) #check it worked, output should be "Date"

growth$Measurement_date_2 <- mdy(growth$Measurement_date_2) #repeat for Feb sampling
class(growth$Measurement_date_2) 

#fitness data
fitness$Flip_time_seconds <- period_to_seconds(hms(fitness$Flip_time)) #time as seconds vs hms
fitness$Flip_time <- hms(fitness$Flip_time)
fitness$Flip_time <- hms::hms(fitness$Flip_time) #time as hours, mins, seconds
class(fitness$Flip_time) #should be hms, difftime

#settlement data
settlement_raw_wide$Date <- mdy(settlement_raw_wide$Date) 
class(settlement_raw_wide$Date) 
settlement_raw_long$Date <- mdy(settlement_raw_long$Date)
class(settlement_raw_long$Date)

#specimen data
specimens$Date_measured <- mdy(specimens$Date_measured)
class(specimens$Date_measured)
```


```{r recoding density}
fitness$Density <- factor(
  fitness$Density, 
  levels = c("1larva/ml", "2larva/ml", "5then2larva/ml", "Bucket"),
  labels = c("1 larvae/mL", "2 larvae/mL", "5 larvae/mL", "15-20 larvae/mL"))

settlement_raw_long$Density <- factor(
  settlement_raw_long$Density, 
  levels = c("1larva/ml", "2larva/ml", "5then2larva/ml", "Bucket"),
  labels = c("1 larvae/mL", "2 larvae/mL", "5 larvae/mL", "15-20 larvae/mL"))

survivorship$Density <- factor(
  survivorship$Density, 
  levels = c("1larva/ml", "2larva/ml", "5then2larva/ml", "Bucket"),
  labels = c("1 larvae/mL", "2 larvae/mL", "5 larvae/mL", "15-20 larvae/mL"))

juvenile_settlement$Density <- factor(
  juvenile_settlement$Density, 
  levels = c("1larva/ml", "2larva/ml", "5then2larva/ml", "Bucket"),
  labels = c("1 Larvae/mL", "2 larvae/mL", "5 larvae/mL", "15-20 larvae/mL"))
```

## Larval growth overtime from ImageJ photo meansurements

```{r plotting}
par(mfrow = c(1,1))

#first pass visualization
plot(data = larval_growth, Size ~ DPF)

#ordering densities
larval_growth$Density <- 
  factor(larval_growth$Density, levels = c("1 larvae/mL",
                                           "2 larvae/mL",
                                           "5 larvae/mL",
                                           "15-20 larvae/mL"))

#working df
Larval_growth <- larval_growth %>% arrange(desc(Density)) %>%
  group_by(DPF, Density) %>%
  summarise(mean_size = mean(Size))

#first pass plot
ggplot(Larval_growth, aes(x = DPF, y = mean_size, colour = Density)) +
  geom_point() +
  geom_line() +
  labs(y= "Average size (µm)", x = "Days post fertilization (DPF)") +
  theme_bw()
#day 42 is an extreme outlier, given each group sees dramatic decreases in average size

#exclude day 42, plot
Figure_2 <- Larval_growth %>% filter(!DPF == 42) %>%
  ggplot(aes(x = DPF, y = mean_size, colour = Density)) +
  geom_point() +
  geom_line() +
  scale_color_manual(values = c("#006b3c", "#D2BAAD", "#9F8FAC", "#f3be77")) +
  labs(y= "Average size (µm)", x = "Days post fertilization (DPF)",
       colour = "Larval Density") +
  xlim(0,50) +
  theme_bw() +
  theme(legend.position = 'top')

ggsave('Figure 2.png', Figure_2, path = "/Users/loganstory/Desktop/SSI/Data/Cupid Cohort/Figures", width = 7, height = 5, units = c("in"), dpi = 300)
```


## Larval settlement data 

```{r cumulative settlement plotting}
#cumulative settlement by larval density 
Density_settlement <- settlement_raw_long %>% 
  group_by(Density, DPF) %>% #initially group by DPF and density
  summarise(Settlers = sum(Settlers)) %>% #sum the settler count by day for each density
  arrange(Density, DPF) %>% #arrange df by density and DPF
  group_by(Density) %>% #cumulative sum only by density 
  mutate(Cum_sum = cumsum(Settlers)) #cumulative sum by density group over-time

#line plot of culuative settlement over DPF, NOT NORMALIZED
ggplot(Density_settlement, aes(x = DPF, y = Cum_sum, colour = Density)) +
  geom_line(linewidth = 0.7) +
  labs(x = "Days post fertilization (DPF)",y = "Approximate larval settlers", 
       colour = "Larval density") +
  scale_x_continuous(breaks = seq(0, 150, by = 10)) +
  scale_color_manual(values = c("#006b3c", "#D2BAAD", "#9F8FAC", "#f3be77")) +
  theme_bw() +
  theme(legend.position = 'top')

#cum settlement by density normalized according to # of replicates in density groups
replicate_counts <- c("1 larvae/mL" = 8, #replicates in each density trtment
                      "2 larvae/mL" = 4, 
                      "5 larvae/mL" = 4, 
                      "15-20 larvae/mL" = 4)

#normalized daily settlement cum df
Density_settlement_normalized <- settlement_raw_long %>%
  group_by(Density, DPF) %>% #summarize by density and DPF
  summarise(Daily_settlers = sum(Settlers), .groups = "drop") %>% #sum by daily settlers
  arrange(Density, DPF) %>% #rearrange df
  group_by(Density) %>% #calculate and mutate by density groups
  mutate(Cum_sum = cumsum(Daily_settlers),
    Normalized_cum_sum = Cum_sum / replicate_counts[Density]) #divide cumsum by # replicates

#normalized cumulative settlement over DPF, plot
Figure_4a <- ggplot(Density_settlement_normalized, 
       aes(x = DPF, y = Normalized_cum_sum, colour = Density)) +
  geom_line(linewidth = 0.7) +
  labs(x = "Days post fertilization (DPF)", 
       y = "Cumulative daily settlement",
       colour = "Larval density", subtitle = "(A)") +
  scale_color_manual(values = c("#006b3c", "#D2BAAD", "#9F8FAC", "#f3be77")) +
  scale_x_continuous(breaks = seq(0, 150, by = 10)) +
  theme_bw() +
  theme(legend.position = 'none')

ggsave('Figure 4a.png', Figure_4a, path = "/Users/loganstory/Desktop/SSI/Data/Cupid Cohort/Figures", width = 7, height = 5, units = c("in"), dpi = 300)
```

```{r GAM non normalized}
Settlement_dpf <- settlement_raw_long %>%
  mutate(Density = as.factor(Density),
         Group = as.factor(Group),
         Settlers = as.integer(Settlers), 
         DPF_scaled = scale(DPF)) 

#GAM 
settler_gam <- gam(Settlers ~ s(DPF, by = Density) + Density, 
                 family = nb(), data = Settlement_dpf)
summary(settler_gam)

#new df for settlement prediction across DPF range for each density group
new_data <- Settlement_dpf %>% group_by(Density) %>%
  reframe(DPF = seq(min(DPF), max(DPF), length.out = 100)) 

#adding the predicted fits on a response scale
new_data$Predicted <- predict(settler_gam, newdata = new_data, type = "response")

#plot
ggplot(Settlement_dpf, aes(x = DPF, y = Settlers, color = Density)) +
  geom_point(alpha = 0.4) +
  geom_rug(colour = "black") +
  geom_line(data = new_data, aes(x = DPF, y = Predicted, color = Density), size = 1.2) +
  labs(y = "Daily Settler Count (Observed + Predicted)",
       x = "Days Post Fertilization") +
  theme_bw() +
  theme(legend.position = 'top')

#plot GAM smooths apart from graphed data
plot(settler_gam, pages = 1, se = TRUE, shade = TRUE)
```

```{r GAM normalized}
Settlement_dpf_normalized <- settlement_raw_long %>%
  group_by(Density, DPF) %>% 
  arrange(Density, DPF) %>% 
  group_by(Density) %>% 
  mutate(Normalized_settlers = Settlers / replicate_counts[Density],
         Density = as.factor(Density),
         Group = as.factor(Group),
         Settlers = as.integer(Settlers))

#density as factor
Settlement_dpf_normalized$Density <- as.factor(Settlement_dpf_normalized$Density)

#nromalized GAM
settler_norm_gam <- gam(Normalized_settlers ~ s(DPF, by = Density) + Density, 
                 family = nb(), data = Settlement_dpf_normalized)
summary(settler_norm_gam)
coef(settler_norm_gam)

#visualizing GAM outputs:
par(mfrow = c(2,2))
plot(settler_norm_gam, shade = TRUE) #4 splines output
abline(h = 0, lty = 'dashed') #adding line at zero
par(mfrow = c(1,1))
```

```{r plot predictions GAM normalized}
#plot predictions of avg effects of DPF using marginal effects
#aggregated across density treatments
plot_predictions(settler_norm_gam, condition = 'DPF', type = 'link') +
  labs(y = "Linear predictor (link scale)") +
  scale_x_continuous(breaks = seq(0, 150, by = 10)) +
  theme_clean()

#plot predictions by density treatment
plot_predictions(settler_norm_gam, condition = c('DPF', 'Density'), type = 'link') +
  labs(y = "Linear predictor (link scale)") +
  scale_x_continuous(breaks = seq(0, 150, by = 10)) +
  scale_color_manual(values = c("#006b3c", "#D2BAAD", "#9F8FAC", "#f3be77")) +
  theme_bw()

#plotting GAM effect on response scale
#aggregated across larval densities
plot_predictions(settler_norm_gam, condition = 'DPF', type = 'response') +
  labs(y = "Expected response") +
  scale_x_continuous(breaks = seq(0, 150, by = 10)) +
  theme_bw()

#plot predictions by density treatment (without obvs)
plot_predictions(settler_norm_gam, condition = c('DPF', 'Density'), 
                 type = 'response') +
  labs(y = "Expected response") +
  scale_color_manual(values = c("#006b3c", "#D2BAAD", "#9F8FAC", "#f3be77")) +
  scale_x_continuous(breaks = seq(0, 150, by = 10)) +
  theme_bw() +
  theme(legend.position = "none")

#plot with regular observations on a smooth curve
ggplot(Settlement_dpf_normalized, aes(x = DPF, y = Normalized_settlers, 
                                      colour = Density)) +
  geom_smooth(aes(fill = Density), alpha = 0.1, size = 0.5) +
   labs(y = "Observed response", 
        title = "Observed effect of DPF on settlement",
       subtitle = "Aggregated by larval density") +
  scale_color_manual(values = c("#006b3c", "#D2BAAD", "#9F8FAC", "#f3be77")) +
  scale_fill_manual(values = c("#006b3c", "#D2BAAD", "#9F8FAC", "#f3be77")) +
  scale_x_continuous(breaks = seq(0, 150, by = 10)) +
  coord_cartesian(ylim = c(0,40)) + 
  theme_bw()

#plot predictions by density treatment WITH obvs barcoded
plot_predictions(settler_norm_gam, condition = c('DPF', 'Density'), 
                 type = 'response', points = 0.5, rug = TRUE) +
  labs(y = "Expected response") +
  scale_color_manual(values = c("#006b3c", "#D2BAAD", "#9F8FAC", "#f3be77")) +
  scale_fill_manual(values = c("#006b3c", "#D2BAAD", "#9F8FAC", "#f3be77")) +
  theme_bw() +
  theme(legend.position = 'top')

#new df for settlement prediction across DPF range for each density group
Normalized_settlement_prediction <- Settlement_dpf_normalized %>% 
  group_by(Density) %>%
  reframe(DPF = seq(min(DPF), max(DPF), length.out = 100)) 

#adding the predicted fits on a response scale
Normalized_settlement_prediction$Predicted <- predict(
  settler_norm_gam, newdata = Normalized_settlement_prediction, type = "response")

#plot it: GAM-predicted larval settlement trajectories by density
Suppfig_1 <- ggplot(Settlement_dpf_normalized, aes(x = DPF, y = Normalized_settlers, 
                                      color = Density)) +
  geom_point(alpha = 0.4) + #make actual obvs more translucent
  geom_line(data = Normalized_settlement_prediction, #plotting GAM predictions
            aes(x = DPF, y = Predicted, color = Density), size = 1) +
  labs(y = "Normalized daily settler count (observed + predicted)",
       x = "Days post fertilization") +
  scale_color_manual(values = c("#006b3c", "#D2BAAD", "#9F8FAC", "#f3be77")) +
  theme_bw() +
  theme(legend.position = 'top')

ggsave('Supp Fig 1.png', Suppfig_1, path = "/Users/loganstory/Desktop/SSI/Data/Cupid Cohort/Supplementary Figures", width = 7, height = 5, units = c("in"), dpi = 300)

#plot GAM smooths apart from graphed data
plot(settler_norm_gam, pages = 1, se = TRUE, shade = TRUE)
```

```{r simulating from fitted GAM}
#simulating from fitted GAM
#using linear predictor matrix Xlp
xp <- predict.gam(settler_norm_gam, type = 'lpmatrix')
dim(xp) #dimensions of prediction- 40 columns of basis functions for smooth terms in GAM
colnames(xp) #see smooth terms of spline

#fitted GAM coefficients
beta <- coef(settler_norm_gam) 

#covariance matrix of coefficients
Vb <- vcov(settler_norm_gam)

#number of simulated curves
n_sims <- 1000
set.seed(123)
beta_sims <- mvrnorm(n = n_sims, mu = beta, Sigma = Vb)

#multiply each sim coef vector by linear predictor
sim_preds <- beta_sims %*% t(xp)  
#results: rows = simulations, columns = predicted points

#mean and 95% confidence intervals
pred_mean  <- apply(sim_preds, 2, mean)
pred_lower <- apply(sim_preds, 2, quantile, probs = 0.025)
pred_upper <- apply(sim_preds, 2, quantile, probs = 0.975)

#combine with prediction grid for plotting
pred_grid <- data.frame(
  DPF   = settler_norm_gam$model$DPF,
  mean  = pred_mean,
  lower = pred_lower,
  upper = pred_upper)

#plot
ggplot(pred_grid, aes(x = DPF, y = mean)) +
  geom_line(color = "black", size = 0.8) +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.2, fill = "gray") +
  labs(x = "Days post fertilization (DPF)",
       y = "Predicted normalized settlers") +
  theme_bw()
```

```{r end settlement statistical analysis}
Density_settlement_norm_group <- settlement_raw_long %>%
  group_by(Density, DPF, Group) %>% 
  summarise(Daily_settlers = sum(Settlers)) %>%
  arrange(Density, DPF) %>%
  group_by(Group) %>% 
  mutate(Cum_sum = cumsum(Daily_settlers),
    Normalized_cum_sum = Cum_sum / replicate_counts[Density])

#df for testing, filtering out to final day of settlement period
Norm_sum <- Density_settlement_norm_group %>% 
  filter(DPF >= 146) %>%
  group_by(Density) %>%
  mutate(se = sd(Normalized_cum_sum)/sqrt(length(Normalized_cum_sum)))

#mean for below graph
Norm_mean_sum <- Density_settlement_norm_group %>% 
  filter(DPF >= 146) %>%
  group_by(Density) %>%
  summarise(Mean_settle = mean(Normalized_cum_sum),
            se = sd(Normalized_cum_sum)/sqrt(length(Normalized_cum_sum)))

#descriptive stats
by(Norm_sum$Normalized_cum_sum, Norm_sum$Density, stat.desc)

#visually check diffs in mean
ggplot(Norm_mean_sum, aes(x = Density, y = Mean_settle, group = 1)) +
  geom_line() +
  geom_point() +
  geom_jitter(data = Norm_sum, aes(y = Normalized_cum_sum, 
                                         colour = Density, shape = Density)) +
  geom_errorbar(width = 0.1, colour = "red", 
                aes(ymin = Mean_settle - se, ymax = Mean_settle + se)) +
  labs(x = "Larval Density", y = "Mean normalized end settlement") +
  theme_bw() +
  theme(axis.title.x = element_blank())

leveneTest(Norm_sum$Normalized_cum_sum, Norm_sum$Density, center = median) 
#appropriate, p = 0.88

ggplot(Norm_sum, aes(sample = Normalized_cum_sum)) +
     geom_qq() +
     geom_qq_line() +
     facet_wrap(~Density, scales = "free_y")

shapiro.test(Norm_sum$Normalized_cum_sum) 
#p = 0.623, overall normal, proceed with anova given normality from NON normalized data

#ANOVA
ds_norm_aov <- Norm_sum %>%
  ungroup() %>%
  anova_test(Normalized_cum_sum ~ Density)
ds_norm_aov # p < 0.05

#tukey test
ds_norm_ph <- Norm_sum %>%
  ungroup() %>%
  tukey_hsd(Normalized_cum_sum ~ Density)
ds_norm_ph #signf between 1-5* and 1-15/20***

#plot workflow
ds_norm_ph <- ds_norm_ph %>% arrange(p.adj)
ds_norm_ph <- add_xy_position(test = ds_norm_ph,
  data = Norm_sum %>% as_tibble(.name_repair = "unique"),  
  formula = Normalized_cum_sum ~ Density, fun = "mean", x = "Density")
x_levels <- levels(factor(Norm_sum$Density))
x_map <- setNames(seq_along(x_levels), x_levels)
ds_norm_ph <- ds_norm_ph %>%
  mutate(xmin = x_map[group1], xmax = x_map[group2])
ds_norm_ph <- ds_norm_ph
ds_norm_ph$y.position[1] <- 630 
ds_norm_ph$y.position[2] <- 570
```

```{r plotting}
#cumulative settlement at end of settlement period (reg box plot)
ggplot(Norm_sum, aes(x = Density, y = Normalized_cum_sum)) +
  geom_boxplot() +
  geom_jitter(aes(colour = Density, shape = Density), size = 2) +
  stat_boxplot(geom = 'errorbar', width = 0.2) +
  labs(x = "Larval Density", y = "Total Cumulative Settlers (normailized)") +
  scale_shape_manual(values = c(15, 16, 17, 18)) +
  scale_color_manual(values = c("#006b3c", "#D2BAAD", "#9F8FAC", "#f3be77")) +
  theme_bw() +
  theme(legend.position = 'none')

#plot with stats
Figure_4b <- ggplot(Norm_sum, aes(x = Density, y = Normalized_cum_sum)) +
  stat_boxplot(geom = "errorbar", width = 0.1) +
  geom_boxplot(outlier.shape = 19) +
  geom_jitter(aes(x = Density, y = Normalized_cum_sum, 
                  colour = Density),
              data = Norm_sum, width = 0.2, alpha = 1, size = 2) +
  stat_pvalue_manual(ds_norm_ph, label = "p.adj.signif", hide.ns = TRUE) +
  stat_compare_means(method = "anova", label.y = 680, label.x = "1 larvae/mL") +
  labs(x = "Larval density", y = "End cumulative settlement",
       subtitle = "(B)", colour = "Larval density") +
  scale_shape_manual(values = c(15, 16, 17, 18)) +
  scale_color_manual(values = c("#006b3c", "#D2BAAD", "#9F8FAC", "#f3be77")) +
  coord_cartesian(ylim = c(0, 700)) +  #extend y-axis to fit signf brackets
  theme_bw() +
  theme(legend.position = "none")

ggsave('Figure 4b.png', Figure_4b, path = "/Users/loganstory/Desktop/SSI/Data/Cupid Cohort/Figures", width = 7, height = 5, units = c("in"), dpi = 300)

Figure_4 <- Figure_4a + Figure_4b

ggsave('Figure 4.png', Figure_4, path = "/Users/loganstory/Desktop/SSI/Data/Cupid Cohort/Figures", width = 10, height = 5, units = c("in"), dpi = 300)
```

```{r survivorship statistical analysis}
#calculate survivorship as % rather than decimal
survivorship$Survivorship <- survivorship$Survivorship * 100

#descriptive stats
by(survivorship$Survivorship, survivorship$Density, stat.desc)

#best and worst performing buckets
survivorship %>%
  filter(Survivorship == min(Survivorship)) #low surv is 0.5% from Scorpius (bucket)
survivorship %>%
  filter(Survivorship == max(Survivorship)) #hig surv is 23% in Merope (1 larv)

#homogeniety of variance
leveneTest(survivorship$Survivorship, survivorship$Density, center = median)
#not appropriate, p = 0.01

#normality of dist
survivorship %>% shapiro_test(Survivorship)
#not normally dist

#kruskal wallis non parametric alaternative
ss_kw <- survivorship %>%
  kruskal_test(Survivorship ~ Density)
ss_kw #p < 0.002*

#post-hoc dunn test
ss_ph <- survivorship %>%
  dunn_test(Survivorship ~ Density, p.adjust.method = "bonferroni")
ss_ph #signf between 1/15-20 **

ss_ph <- ss_ph %>% arrange(p.adj)
ss_ph <- ss_ph %>% add_xy_position(x = "Density")

```

```{r plotting}
#without stats
ggplot(survivorship, aes(x = Density, y = Survivorship)) +
  stat_boxplot(geom = 'errorbar', width = 0.2) +
  geom_boxplot() +
  geom_jitter(aes(shape = Density, colour = Density)) +
  scale_shape_manual(values = c(15, 16, 17, 18)) +
  scale_color_manual(values = c("#006b3c", "#D2BAAD", "#9F8FAC", "#f3be77")) +
  theme_bw() +
  labs(y = "Larval settlement rate (%)", x = "Larval Density") +
  theme_bw() +
  theme(legend.position = "none")

#with stats
Figure_5 <- ggplot(survivorship, aes(x = Density, y = Survivorship)) +
  stat_boxplot(geom = 'errorbar', width = 0.2) +
  geom_boxplot() +
  geom_jitter(aes(shape = Density, colour = Density)) +
  stat_pvalue_manual(ss_ph, label = "p.adj.signif", 
                     hide.ns = TRUE, y.position = "y.position") +
  stat_compare_means(label.y = 31, label.x = "1 larvae/mL") +
  scale_y_continuous(labels = function(x) paste0(x, "%")) +
  scale_shape_manual(values = c(15, 16, 17, 18)) +
  scale_color_manual(values = c("#006b3c", "#D2BAAD", "#9F8FAC", "#f3be77")) +
  theme_bw() +
  labs(y = "Larval Settlement Rate (%)", x = "Larval Density") +
  theme_bw() +
  theme(legend.position = "none")

ggsave('Figure 5.png', Figure_5, path = "/Users/loganstory/Desktop/SSI/Data/Cupid Cohort/Figures", width = 7, height = 5, units = c("in"), dpi = 300)
```

## Juvenile survival/settlement data and diffs in growth

```{r statistical analysis}
#calculative settlement rate (#juveniles on feb 25 / #settlers on july 24)
juvenile_settlement <- juvenile_settlement %>% 
  dplyr::select(Group, Density, Juvenile_total, Larval_total) %>%
  group_by(Density) %>%
  mutate(Juv_survival_rate = Juvenile_total / Larval_total)

#survival as %
juvenile_settlement$Juv_survival_rate <- juvenile_settlement$Juv_survival_rate * 100

#descriptive stats
by(juvenile_settlement$Juv_survival_rate, juvenile_settlement$Density, stat.desc)

#best and worst performing buckets
juvenile_settlement %>%
  filter(Juv_survival_rate == min(Juv_survival_rate)) 
  #low surv is 0.05% from Keig (1 larv/ml)
juvenile_settlement %>%
  filter(Juv_survival_rate == max(Juv_survival_rate)) 
  #hig surv is 1.79% in Betelgeuse (1 larv/ml)

#homogeniety of variance
leveneTest(juvenile_settlement$Juv_survival_rate, juvenile_settlement$Density, center = median) #appropraite, p = 0.07

#normality of dist
juvenile_settlement %>% group_by(Density) %>%
  shapiro_test(Juv_survival_rate) #normally dist, p > 0.05 for all groups

#anova
js_aov <- juvenile_settlement %>%
  ungroup() %>%
  anova_test(Juv_survival_rate ~ Density)
js_aov #p = 0.5, non signf

#post-hoc dunn test
js_ph <- juvenile_settlement %>%
  ungroup() %>%
  tukey_hsd(Juv_survival_rate ~ Density)
js_ph #no between group signf

js_ph <- js_ph %>% arrange(p.adj) 
js_ph <- add_xy_position(test = js_ph,
  data = juvenile_settlement %>% as_tibble(.name_repair = "unique"),
  formula = Juv_survival_rate ~ Density, fun = "mean", x = "Density")

x_levels <- levels(factor(juvenile_settlement$Density))
x_map <- setNames(seq_along(x_levels), x_levels)
js_ph <- js_ph %>%
  mutate(xmin = x_map[group1], xmax = x_map[group2])

js_ph <- js_ph %>% mutate(y.position = y.position + 40)
js_ph$y.position[2] <- 95 #specific y position of 2nd signf bracket

```

```{r plotting}
juvenile_settlement$Juv_survival_rate <- juvenile_settlement$Juv_survival_rate / 100

#no stat plot
Figure_6 <- ggplot(juvenile_settlement, aes(x = Density, y = Juv_survival_rate)) +
  stat_boxplot(geom ='errorbar', width = 0.3) +
  geom_boxplot() +
  geom_jitter(aes(shape = Density, colour = Density)) +
  scale_y_continuous(labels = scales::percent) +
  scale_shape_manual(values = c(15, 16, 17, 18)) +
  scale_colour_manual(values = c("#006b3c", "#D2BAAD", "#9F8FAC", "#f3be77")) +
  theme_bw() +
  labs(y = "Survival from end of settlement \n(5 months post-fertilization) to 1 year post-fertilization",
       x = "Larval Density") +
  theme(legend.position = "none")

ggsave('Figure 6.png', Figure_6, path = "/Users/loganstory/Desktop/SSI/Data/Cupid Cohort/Figures", width = 7, height = 5, units = c("in"), dpi = 300)
```

## Larval density and juvenile weight

*exploration and analysis*
```{r statistical anlaysis}
#working df for length/weight analysis
Density_size <- fitness %>% 
  dplyr::select(Group, Individual, Density, Length_mm, Weight_g) %>%
  drop_na(Density, Length_mm) %>%
  group_by(Density) %>% 
   mutate(mean_length = mean(Length_mm),
         se_length = sd(Length_mm)/sqrt(length(Length_mm)),
         mean_weight = mean(Weight_g),
         se_weight = sd(Weight_g)/sqrt(length(Weight_g))) 

#Largest and smallest star
Density_size %>%
  filter(Weight_g == min(Weight_g),
         Length_mm == min(Length_mm)) #smallest star is 12mm and 0.2g in bucket group
Density_size %>%
  filter(Weight_g == max(Weight_g),
         Length_mm == max(Length_mm)) #largest star is 70mm and 12.1 in 1 larv group

#plotted diff in mean length
ggplot(Density_size, aes(x = Density, y = mean_weight, group = 1)) +
  geom_line() +
  geom_point() +
  geom_errorbar(width = 0.1, colour = "red", 
                aes(ymin = mean_weight - se_weight, 
                    ymax = mean_weight + se_weight)) #bucket group is signf smaller than other 3, 1/5 larv are relatively similar

#descriptive stats
by(Density_size$Weight_g, Density_size$Density, stat.desc)

#identify outliers, save in object to view whole list
Density_weight_outliers <- Density_size %>% 
  group_by(Density) %>%
  identify_outliers(Weight_g) #4 outliers, 1 is extreme

#exclude extreme outlier
Density_weight_flip_no_outliers <- Density_size %>%
  filter(!Weight_g == 12.1) #only get rid of indiv 1 from betelgeuse

#visualization of variation
ggqqplot(Density_weight_flip_no_outliers, "Weight_g") +
  facet_grid(~Density) + #multiple outliers in 1 larv group (big big boys)
  theme_bw() 

#checking assumption of homogeniety 
leveneTest(Density_weight_flip_no_outliers$Weight_g, 
           Density_weight_flip_no_outliers$Density, center = median) 
#appropriate, p = 0.16

#shapiro test
Density_weight_flip_no_outliers %>% group_by(Density) %>%
  shapiro_test(Weight_g) #2/5 groups inappropriate
Density_weight_flip_no_outliers %>%
  ungroup() %>%
  shapiro_test(Weight_g) #overall normality is very inappropriate p < 0.001

#non parametric analysis, given above testing
#KW
dw_kw <- Density_weight_flip_no_outliers %>%
  ungroup() %>%
  anova_test(Weight_g ~ Density) 
dw_kw #p < 0.004**

#post-hoc tukey test
dw_ph <- Density_weight_flip_no_outliers %>% 
  ungroup() %>%
  dunn_test(Weight_g ~ Density, p.adjust.method = 'bonferroni') 
dw_ph

#work flow for publication-ready plot
dw_ph <- dw_ph %>% arrange(p.adj) #arrange
dw_ph <- add_xy_position(test = dw_ph,
  data = Density_weight_flip_no_outliers %>% as_tibble(.name_repair = "unique"),
  formula = Weight_g ~ Density, fun = "mean", x = "Density")

#factor levels in the order used in the plot
x_levels <- levels(factor(Density_weight_flip_no_outliers$Density))

#vector to convert group names to x-axis positions
x_map <- setNames(seq_along(x_levels), x_levels)

#map to assign numeric values to xmin and xmax
dw_ph <- dw_ph %>%
  mutate(xmin = x_map[group1], xmax = x_map[group2])

#adjusting y.position to fit the signf brackets
dw_ph <- dw_ph %>% mutate(y.position = y.position - 40)
dw_ph$y.position[1] <- 13
dw_ph$y.position[2] <- 14 #specific y position of 2nd signf bracket
```

*plotting*
```{r plotting}
#reg box plot
ggplot(Density_weight_flip_no_outliers, aes(x = Density, y = Weight_g)) +
  stat_boxplot(geom='errorbar', width = 0.2) +
  geom_boxplot(colour = "black") +
  geom_jitter(aes(colour = Density, shape = Density), size = 2) +
  labs(x = "Larval density", y = "Weight (g)") +
  coord_cartesian(ylim = c(0, 15)) +  
  scale_shape_manual(values = c(15, 16, 17, 18)) +
  scale_colour_manual(values = c("#006b3c", "#D2BAAD", "#9F8FAC", "#f3be77")) +
  theme_bw() +
  theme(legend.position = "none")

#box blot with signf bars and stats
Suppfig_2b <- ggplot(Density_weight_flip_no_outliers, aes(x = Density, y = Weight_g)) +
  stat_boxplot(geom='errorbar', width = 0.2) +
  geom_boxplot(colour = "black") +
  geom_jitter(aes(colour = Density, shape = Density), size = 2) +
  stat_pvalue_manual(dw_ph, label = "p.adj.signif", hide.ns = TRUE) +
  stat_compare_means(label.y = 14.3, label.x = "1 larvae/mL") +
  labs(x = "Larval density", y = "Weight (g)") +
  coord_cartesian(ylim = c(0, 15)) +  
  scale_shape_manual(values = c(15, 16, 17, 18)) +
  scale_colour_manual(values = c("#006b3c", "#D2BAAD", "#9F8FAC", "#f3be77")) +
  theme_bw() +
  theme(legend.position = "none")

ggsave('Supp fig 2b.png', Suppfig_2b, path = "/Users/loganstory/Desktop/SSI/Data/Cupid Cohort/Supplementary Figures", width = 7, height = 5, units = c("in"), dpi = 300)
```

## Larval density and juvenile length 

*exploration and analysis*
```{r statistical analysis}
#working df for length/weight analysis
Density_size <- fitness %>% 
  dplyr::select(Group, Individual, Density, Length_mm, Weight_g) %>%
  drop_na(Density, Length_mm) %>%
  group_by(Density) %>% 
   mutate(mean_length = mean(Length_mm),
         se_length = sd(Length_mm)/sqrt(length(Length_mm)),
         mean_weight = mean(Weight_g),
         se_weight = sd(Weight_g)/sqrt(length(Weight_g))) 

#Largest and smallest star
Density_size %>%
  filter(Weight_g == min(Weight_g),
         Length_mm == min(Length_mm)) #smallest star is 12mm and 0.2g in bucket group
Density_size %>%
  filter(Weight_g == max(Weight_g),
         Length_mm == max(Length_mm)) #largest star is 70mm and 12.1 in 1 larv group

#plotted diff in mean length
ggplot(Density_size, aes(x = Density, y = mean_length, group = 1)) +
  geom_line() +
  geom_point() +
  geom_errorbar(width = 0.1, colour = "red", 
                aes(ymin = mean_length - se_length, 
                    ymax = mean_length + se_length)) #bucket group is signf smaller than other 3

#descriptive stats
by(Density_size$Length_mm, Density_size$Density, stat.desc)

#identify outliers, save in object to view whole list
Density_size_outliers <- Density_size %>% 
  group_by(Density) %>%
  identify_outliers(Length_mm) #1 outlier, not extreme

#checking assumption of homogeniety 
leveneTest(Density_size$Length_mm, 
           Density_size$Density, center = median) #appropriate, p = 0.81

#shapiro test
Density_size %>% group_by(Density) %>%
  shapiro_test(Length_mm) #all groups appropriate, p > 0.05

#visualization of variation
ggqqplot(Density_size, "Length_mm") +
  facet_grid(~Density) + #single outlier in 5 larv group, not extreme
  theme_bw() 

#parametric analysis, assumed independence of observations
#anova
dl_aov <- Density_size %>%
  ungroup() %>%
  anova_test(Length_mm ~ Density) 
dl_aov #p < 0.05**

#post-hoc tukey test
dl_ph <- Density_size %>% 
  ungroup() %>%
  tukey_hsd(Length_mm ~ Density) 
dl_ph

#work flow for publication-ready plot
dl_ph <- dl_ph %>% arrange(p.adj) #arrange
dl_ph <- add_xy_position(test = dl_ph,
  data = Density_size %>% as_tibble(.name_repair = "unique"),  #unique column names
  formula = Length_mm ~ Density, fun = "mean", x = "Densities")

#factor levels in the order used in the plot
x_levels <- levels(factor(Density_size$Density))

#vector to convert group names to x-axis positions
x_map <- setNames(seq_along(x_levels), x_levels)

#map to assign numeric values to xmin and xmax
dl_ph <- dl_ph %>%
  mutate(xmin = x_map[group1], xmax = x_map[group2])

#adjusting y.position to fit the signf brackets
dl_ph <- dl_ph %>% mutate(y.position = y.position + 40)
dl_ph$y.position[2] <- 95 #specific y position of 2nd signf bracket
```

*plotting*
```{r plotting}
#reg box plot
ggplot(Density_size, aes(x = Density, y = Length_mm)) +
  stat_boxplot(geom='errorbar', width = 0.2) +
  geom_boxplot(colour = "black") +
  geom_jitter(aes(colour = Density, shape = Density), size = 2) +
  labs(x = "Larval density", y = "Length (mm)") +
  coord_cartesian(ylim = c(0, 80)) +  
  scale_shape_manual(values = c(15, 16, 17, 18)) +
  scale_colour_manual(values = c("#006b3c", "#D2BAAD", "#9F8FAC", "#f3be77")) +
  theme_bw() +
  theme(legend.position = "none")

#box blot with signf bars and stats
Suppfig_2a <- ggplot(Density_size, aes(x = Density, y = Length_mm)) +
  stat_boxplot(geom='errorbar', width = 0.2) +
  geom_boxplot(colour = "black") +
  geom_jitter(aes(colour = Density, shape = Density), size = 2) +
  stat_pvalue_manual(dl_ph, label = "p.adj.signif", hide.ns = TRUE) +
  stat_compare_means(method = "anova", label.y = 101, label.x = "1 larvae/mL") +
  labs(x = "Larval density", y = "Length (mm)") +
  coord_cartesian(ylim = c(0, 105)) +  
  scale_shape_manual(values = c(15, 16, 17, 18)) +
  scale_colour_manual(values = c("#006b3c", "#D2BAAD", "#9F8FAC", "#f3be77")) +
  theme_bw() +
  theme(legend.position = "none")

ggsave('Supp fig 2a.png', Suppfig_2a, path = "/Users/loganstory/Desktop/SSI/Data/Cupid Cohort/Supplementary Figures", width = 7, height = 5, units = c("in"), dpi = 300)

Suppfig_2 <- Suppfig_2a + Suppfig_2b

ggsave('Supp fig 2.png', Suppfig_2, path = "/Users/loganstory/Desktop/SSI/Data/Cupid Cohort/Supplementary Figures", width = 10, height = 5, units = c("in"), dpi = 300)
```

## Larval density and juvenile flip time

*exploration and statistical analysis*
```{r statistical analysis}
#working vector for  density/flip exploration
Density_flip <- fitness %>% 
  dplyr::select(Group, Individual, Density, Flip_time, Flip_time_seconds) %>% 
  drop_na(Density, Flip_time) %>% #remove unknowns
  group_by(Density) %>% 
  mutate(mean_time = mean(Flip_time_seconds),
         se = sd(Flip_time_seconds)/sqrt(length((Flip_time_seconds))))

#largest and smallest flip_time
Density_flip %>%
  filter(Flip_time == min(Flip_time)) 
#fastest was 49 seconds from quasar indiv 2 (15-20 larv)
Density_flip %>%
  filter(Flip_time == max(Flip_time))
#slowest was 30 minutes, both electra stars 

#visual diff in mean time to flip
ggplot(Density_flip, aes(x = Density, y = mean_time, group = 1)) +
  geom_line() +
  geom_point() +
  geom_errorbar(width = 0.1, colour = "red", 
                aes(ymin = mean_time - se, ymax = mean_time + se))

#descriptive stats
by(Density_flip$Flip_time_seconds, Density_flip$Density, stat.desc)

#identify outliers, save in object to view whole list
Density_flip_outliers <- Density_flip %>% 
  group_by(Density) %>%
  identify_outliers(Flip_time) #8 outliers, 5 are extreme

#remove 5 extreme outliers from analysis
Density_flip_no_outliers <- Density_flip %>% 
   group_by(Density) %>%
   filter(!Flip_time_seconds == 1800,
          !Flip_time_seconds == 1322,
          !Flip_time_seconds == 1132,
          !Flip_time_seconds == 992)

#checking assumption of homogenity for testing
leveneTest(Density_flip_no_outliers$Flip_time_seconds, 
           Density_flip_no_outliers$Density, center = median) #appropriate, p = 0.40

#shapiro test
Density_flip_no_outliers %>% group_by(Density) %>%
  shapiro_test(Flip_time_seconds) #3/4 groups non-appropriate, p < 0.05

#visualization of variation
ggqqplot(Density_flip_no_outliers, "Flip_time_seconds") +
  facet_grid(~Density) +
  theme_bw() 

#Kruskal-walis non-parametric alternative to ANOVA
df_kw <- Density_flip_no_outliers %>%
  ungroup() %>%
  kruskal_test(Flip_time ~ Density)
df_kw #p = 0.317

#Post-hoc testing
df_ph <- Density_flip_no_outliers %>%
  ungroup() %>% 
  dunn_test(Flip_time_seconds ~ Density, p.adjust.method = "bonferroni")
df_ph

#arrange by adjusted p-value for plotting
df_ph <- df_ph %>% arrange(p.adj)
df_ph <- df_ph %>% add_xy_position(x = "Density")
```

*plotting*
```{r plotting}
#box plot, removed extreme outliers post outlier analysis
Figure_7a <- ggplot(Density_flip_no_outliers, aes(x = Density, y = Flip_time)) +
  stat_boxplot(geom = 'errorbar', width = 0.2) +
  geom_boxplot() +
  geom_jitter(aes(colour = Density, shape = Density), size = 2) +
  labs(x="Larval density", y="Flip time") +
  scale_shape_manual(values = c(15, 16, 17, 18)) +
  scale_colour_manual(values = c("#006b3c", "#D2BAAD", "#9F8FAC", "#f3be77")) +
  theme_bw() +
  theme(legend.position = "none")

ggsave('Figure 7a.png', Figure_7a, path = "/Users/loganstory/Desktop/SSI/Data/Cupid Cohort/Figures", width = 7, height = 5, units = c("in"), dpi = 300)

#plot (with outliers)
ggplot(Density_flip, aes(x = Density, y = Flip_time)) +
  stat_boxplot(geom = 'errorbar', width = 0.2) +
  geom_boxplot() +
  geom_jitter(aes(colour = Density, shape = Density), size = 2) +
  labs(x="Larval density (larvae/mL)", y="Flip time") +
  scale_shape_manual(values = c(15, 16, 17, 18)) +
  scale_colour_manual(values = c("#006b3c", "#D2BAAD", "#9F8FAC", "#f3be77")) +
  theme_bw() +
  theme(legend.position = "none")

#plot with stats
ggboxplot(Density_flip_no_outliers, x = "Density", y = "Flip_time", outliers = FALSE) + 
  geom_jitter(aes(colour = Density, shape = Density), size = 2) +
  stat_boxplot(geom = 'errorbar', width = 0.2) +
  stat_pvalue_manual(df_ph, label = "p.adj.signif", hide.ns = TRUE) +
  stat_compare_means(label.y = 1270, label.x = "2 larvae/mL") +
  labs(x = "Larval Density", y = "Flip Time") +
  scale_shape_manual(values = c(15, 16, 17, 18)) +
  scale_colour_manual(values = c("#006b3c", "#D2BAAD", "#9F8FAC", "#f3be77")) +
  theme_bw() +
  theme(legend.position = "none", axis.title.x = element_blank())
```

## Larval density and juvenile length and flip time

*linear regression model*
```{r linear regression no outliers}
#working vector
Density_size_flip <- fitness %>%
  dplyr::select(Group, Individual, Density, Length_mm, Flip_time, Flip_time_seconds) %>%
  drop_na(Density, Length_mm) %>%
  filter(!Flip_time_seconds == 1132)

#find and remove outliers (also completed in chunk XX)
Density_size_flip_outliers <- Density_size_flip %>% 
  group_by(Density) %>%
  identify_outliers(Flip_time) #7 outliers, 4 are extreme

#remove 4 extreme outliers from analysis
Density_size_flip_no_outliers <- Density_size_flip %>% 
   group_by(Density) %>%
   filter(!Flip_time_seconds == 1800,
          !Flip_time_seconds == 1322,
          !Flip_time_seconds == 1132)

Density_size_flip_no_outliers$Density <- as.factor(Density_size_flip_no_outliers$Density)

#simple linear regression
slr <- lm(Flip_time_seconds ~ Length_mm, data = Density_size_flip_no_outliers)
summary(slr)

#dummy variables
test <- Density_size_flip_no_outliers %>% 
  dplyr::select(Density, Length_mm, Flip_time, Flip_time_seconds)
dv <- dummy_cols(test)
dv

#interacting model
slr_density <- lm(Flip_time_seconds ~ Length_mm:`Density_2 larvae/mL` + 
                                      Length_mm:`Density_5 larvae/mL` + 
                                      Length_mm:`Density_15-20 larvae/mL`,
                  data = dv)
summary(slr_density)

#compare models
anova(slr_density) #all p ns
anova(slr, slr_density) #p = 0.26

#compare models with null
null <- lm(Flip_time_seconds ~ 1, data = Density_size_flip_no_outliers)
summary(null)
anova(slr, null) #p = 0.08
anova(slr_density, null) #p = 0.15
```

```{r linear regression outliers}
Density_size_flip$Density <- as.factor(Density_size_flip$Density)

#simple linear regression w/ outliers
slr_o <- lm(Flip_time_seconds ~ Length_mm, data = Density_size_flip)
summary(slr_o) #p = 0.1, overall F test p = 0.1

#dv outliers
test <- Density_size_flip %>% 
  dplyr::select(Density, Length_mm, Flip_time, Flip_time_seconds)
dv_o <- dummy_cols(test)
dv_o

#interacting model w/ outliers
slr_density_o <- lm(Flip_time_seconds ~ Length_mm:`Density_2 larvae/mL` + 
                                      Length_mm:`Density_5 larvae/mL` + 
                                      Length_mm:`Density_15-20 larvae/mL`,
                  data = dv_o)
summary(slr_density_o) #ns, overall F test p = 0.188
```

*plotting*
```{r plotting}
#plot with lm line
Figure_7b <- ggplot(Density_size_flip_no_outliers, aes(x = Length_mm, y = Flip_time)) +
  geom_smooth(data = Density_size_flip_no_outliers,
              method = lm, colour = "black", se = TRUE) +
  geom_point(aes(shape = Density, colour = Density), size = 2) +
  scale_shape_manual(values = c(15, 16, 17, 18),
                     name = "Larval density") +
  scale_color_manual(values = c("#006b3c", "#D2BAAD", "#9F8FAC", "#f3be77"),
                     name = "Larval density") +
  labs(x = "Length (mm)",
       y = "Flip time", subtitle = "(B)") +
  theme_bw() +
  theme(legend.position = 'none')

#combine figure 7a and 7b
Figure_7a <- ggplot(Density_flip_no_outliers, aes(x = Density, y = Flip_time)) +
  stat_boxplot(geom = 'errorbar', width = 0.2) +
  geom_boxplot() +
  geom_jitter(aes(colour = Density, shape = Density), size = 2) +
  labs(x = "Larval density", y = "Flip time", subtitle = "(A)") +
  scale_shape_manual(values = c(15, 16, 17, 18)) +
  scale_colour_manual(values = c("#006b3c", "#D2BAAD", "#9F8FAC", "#f3be77")) +
  theme_bw() +
  theme(legend.position = "none")

Figure_7 <- Figure_7a + Figure_7b

ggsave('Figure 7.png', Figure_7, path = "/Users/loganstory/Desktop/SSI/Data/Cupid Cohort/Figures", width = 10, height = 5, units = c("in"), dpi = 300)
```

## Specimen models

*asymp modeling*
```{r asymp model and bootstrap}
specimens <- read.csv("Pycnopodia specimen measurements.csv")

#reordering
specimens$Specimen_type <- factor(
  specimens$Specimen_type, 
  levels = c("Specimen", "Aquarium display", "Cupid Cohort"), 
  labels = c("Museum specimen", "Alaskan stars", "Cupid cohort"))

#nls asymp model
asymp <- nls(Arm_count ~ SSasymp(Length_cm, Asym, R0, lrc), data = specimens)
summary(asymp)
coef(asymp) 

#creating seq of 100 evenly spaced body lengths in the observed body size range 
length_seq <- seq(min(specimens$Length_cm), max(specimens$Length_cm), length.out = 100)

#prediction grid for model from length_seq, usable for later model predictions
pred_grid <- data.frame(Length_cm = length_seq)

#bootstrap model fits
set.seed(123)  #for reproducibility, random number seed

#bootstrap model replicates
n_boot <- 1000

#bootstrap replicate
boot_preds <- replicate(n_boot, { #drawing bootstrap sample w/ replacement
  boot_sample <- specimens[sample(1:nrow(specimens), replace = TRUE), ]
  tryCatch({ #fitting model to each bootstrap sample, predict arm count across all lengths
    boot_model <- nls(Arm_count ~ SSasymp(Length_cm, Asym, R0, lrc), data = boot_sample)
    predict(boot_model, newdata = pred_grid)
  }, error = function(e) rep(NA, length(length_seq)))
})

#transpose, remove failed fits (NAs)
boot_preds <- t(boot_preds) 
   #transposes, rows=bootstrap reps; columns=predictions for length
boot_preds <- boot_preds[complete.cases(boot_preds), ] #remove failed fits

#calculate mean, conf interval
pred_grid$mean <- apply(boot_preds, 2, mean) #mean prediction at each length
pred_grid$lower <- apply(boot_preds, 2, quantile, probs = 0.025) #2.5th percentile confint
pred_grid$upper <- apply(boot_preds, 2, quantile, probs = 0.975) #97.5th percentile confint

#extract model coefs, rounds for notation and reporting on graph
coef_vals <- coef(asymp) #saving model coefs in coef_vals
asym <- round(coef_vals["Asym"], 2)
r0 <- round(coef_vals["R0"], 2)
lrc <- round(coef_vals["lrc"], 2)

#creating psuedo R2 from residuals of model and prediction grid
y_pred <- predict(asymp) #predicted y values from model
y_obs <- specimens$Arm_count #the actual observed arm counts
rss <- sum((y_obs - y_pred)^2)  #total model error (diff in observed vs predicted values)
tss <- sum((y_obs - mean(y_obs))^2)  #total variability in data from the mean

#psuedo R2, returning as a result of nrs() not natively calculating this for me :(
r_squared <- 1 - (rss / tss)
round(r_squared, 3) #pretty great fit, psuedo R2 = 0.881
```

*plotting*
```{r plotting}
#annotate with summary stats
label_text <- paste0(
  "Asym = ", asym, "\n",
  "R0 = ", r0, "\n",
  "lrc = ", lrc, "\n",
  "Pseudo R² = ", round(r_squared, 3))

#plot model with coefs
Figure_8a <- ggplot(specimens, aes(x = Length_cm, y = Arm_count)) +
  geom_point(aes(colour = Specimen_type, shape = Specimen_type)) + 
  geom_ribbon(data = pred_grid, aes(x = Length_cm, ymin = lower, ymax = upper),
              alpha = 0.5, fill = "grey",
              inherit.aes = FALSE) + #override aes mapping in call to not inherit aes
  geom_line(data = pred_grid, aes(x = Length_cm, y = mean), 
            color = "black", size = 0.7) +
  geom_vline(xintercept = 20, linetype = "dashed", color = "black") +
  annotate("text", x = 22, y = 10, 
           label = label_text, hjust = 0, size = 3.5) + #adding summary stats to plot
  scale_colour_manual(values = c("#66c2a5", "#fc8d62", "#8da0cb")) +
  coord_cartesian(ylim = c(5, 25)) +
  labs(x = "Length (cm)", y = "Number of arms", shape = "Specimen type",
    colour = "Specimen type", subtitle = "(A)") +
  theme_bw() +
  theme(legend.position = 'none')

ggsave('Figure 8a.png', Figure_8a, path = "/Users/loganstory/Desktop/SSI/Data/Cupid Cohort/Figures", width = 7, height = 5, units = c("in"), dpi = 300)
```

*allometric modelling*
```{r power-rule model and bootstrap}
#reordering/relabeling
specimens_alive$Specimen_type <- factor(
  specimens_alive$Specimen_type, 
  levels = c("Cupid Cohort", "Aquarium display"), 
  labels = c("Cupid cohort", "Alaskan stars"))

#weight as kg
specimens_alive$Weight_kg <- specimens_alive$Weight_g /1000

#model
lm_log <- lm(log(Weight_kg) ~ log(Length_cm), data = specimens_alive)
summary(lm_log)

#back-transform coefficients
a <- exp(coef(lm_log)[1])
b <- coef(lm_log)[2]
b

# Labels
a_label <- formatC(a, format = "e", digits = 2)
b_label <- round(b, 2)
a #4.921003e-05 
b #2.846814, indicates negative allometry

#bootstrapping
set.seed(123)

#number of bootstraps 
n_boot <- 1000

#storing boot coefs
boot_coefs <- map_dfr(1:n_boot, function(i) {
  boot_sample <- specimens_alive %>% sample_frac(replace = TRUE) #re-sample
  boot_model <- lm(log(Weight_kg) ~ log(Length_cm), data = boot_sample) #fit model
  tibble( #extract and return coefs
    iter = i,
    intercept = coef(boot_model)[1],
    slope = coef(boot_model)[2]
  )
})

#back transform coefs
boot_coefs <- boot_coefs %>%
  mutate(a = exp(intercept),
         b = slope)

#summarize w/ confidence intervals
boot_summary <- boot_coefs %>%
  summarise(
    a_mean = mean(a),
    a_ci_lower = quantile(a, 0.025),
    a_ci_upper = quantile(a, 0.975),
    b_mean = mean(b),
    b_ci_lower = quantile(b, 0.025),
    b_ci_upper = quantile(b, 0.975))

#summary, printing
print(boot_summary)

#sequence of Length values to predict over
length_seq <- seq(min(specimens_alive$Length_cm), max(specimens_alive$Length_cm), length.out = 100)

#generate predicted weight curves for each bootstrap replicate
boot_lines <- boot_coefs %>%
  mutate(iter = row_number()) %>%
  group_by(iter) %>%
  do({
    data.frame(
      Length_cm = length_seq,
      Weight_kg = exp(.$intercept + .$slope * log(length_seq)))
  })

boot_lines <- map_dfr(1:nrow(boot_coefs), function(i) {
  data.frame(
    Length_cm = length_seq,
    Weight_kg = exp(boot_coefs$intercept[i] + boot_coefs$slope[i] * log(length_seq)),
    iter = i)
  })

#predicted and observed values
y_pred <- predict(lm_log)
y_obs <- log(specimens_alive$Weight_kg)

#calculate pseudo-R²
rss <- sum((y_obs - y_pred)^2)
tss <- sum((y_obs - mean(y_obs))^2)
r_squared <- 1 - (rss / tss)
r2_label <- round(r_squared, 3)

#precomputed back-transformed values
a_label <- formatC(a, format = "e", digits = 2)
b_label <- round(b, 2)

label_text <- paste0(
  "Weight = ", a_label, " × Length^", b_label, "\n",
  "Pseudo R² = ", r2_label
)
```

*plotting*
```{r plotting}
#bootstrapped curves plot
Figure_8b <- ggplot(specimens_alive, aes(x = Length_cm, y = Weight_kg, 
                            colour = Specimen_type, shape = Specimen_type)) +
  geom_point() +
  geom_line(data = boot_lines, aes(x = Length_cm, y = Weight_kg),
            inherit.aes = FALSE, alpha = 0.2, size = 0.2, color = "grey") +
  stat_function(fun = function(x) a * x^b, color = "black", size = 0.6,
                inherit.aes = FALSE) +
  geom_vline(xintercept = 20, linetype = "dashed", color = "black") +
  annotate("text", x = 22, y = 4.5,
           label = label_text, hjust = 0, size = 3.5) +
  coord_cartesian(xlim = c(0,60)) +
  scale_shape_manual(values = c(15, 17)) +
  scale_colour_manual(values = c("#8da0cb", "#fc8d62")) +
  labs(x = "Length (cm)", y = "Weight (kg)", subtitle = "(B)") +
  theme_bw() +
  theme(legend.position = "none")

ggsave('Figure 8b.png', Figure_8b, path = "/Users/loganstory/Desktop/SSI/Data/Cupid Cohort/Figures", width = 7, height = 5, units = c("in"), dpi = 300)

Figure_8 <- Figure_8a + Figure_8b

ggsave('Figure 8.png', Figure_8, path = "/Users/loganstory/Desktop/SSI/Data/Cupid Cohort/Figures", width = 10, height = 5, units = c("in"), dpi = 300)

#non linear model, no text
ggplot(specimens_alive, aes(x = Length_cm, y = Weight_kg, 
                            colour = Specimen_type, shape = Specimen_type)) +
  geom_point() +
  geom_line(data = boot_lines, aes(x = Length_cm, y = Weight_kg),
            inherit.aes = FALSE, alpha = 0.2, size = 0.2, color = "grey") +
  stat_function(fun = function(x) a * x^b, color = "black", size = 0.6,
                inherit.aes = FALSE) +
  theme_bw() +
  labs(x = "Length (cm)", y = "Weight (kg)") +
  coord_cartesian(xlim = c(0,60)) +
  scale_shape_manual(values = c(15, 17)) +
  scale_colour_manual(values = c("#8da0cb", "#fc8d62")) +
  theme(legend.position = "none")
```

*plot comparison*
```{r allometric and asymptotic models compared}
a_plot <- ggplot(specimens, aes(x = Length_cm, y = Arm_count)) +
  geom_point(aes(colour = Specimen_type, shape = Specimen_type)) + 
  geom_ribbon(data = pred_grid, aes(x = Length_cm, ymin = lower, ymax = upper),
              alpha = 0.5, fill = "grey",
              inherit.aes = FALSE) + #override aes mapping in call to not inherit aes
  geom_line(data = pred_grid, aes(x = Length_cm, y = mean), 
            color = "black", size = 0.7) +
  geom_vline(xintercept = 20, linetype = "dashed", color = "black") +
  annotate("text", x = 22, y = 10, 
           label = label_text, hjust = 0, size = 3.5) + #adding summary stats to plot
  scale_colour_manual(values = c("#66c2a5", "#fc8d62", "#8da0cb")) +
  scale_shape_manual(values = c(19, 17, 15)) +
  coord_cartesian(ylim = c(5, 25)) +
  labs(x = "Length (cm)", y = "Number of arms", 
       shape = "Specimen type", colour = "Specimen type",
       subtitle = "(A)") +
  theme_bw() +
  theme(legend.position = 'none')

b_plot <- ggplot(specimens_alive, aes(x = Length_cm, y = Weight_kg, 
                            colour = Specimen_type, shape = Specimen_type)) +
  geom_point() +
  geom_line(data = boot_lines, aes(x = Length_cm, y = Weight_kg),
            inherit.aes = FALSE, alpha = 0.2, size = 0.2, color = "grey") +
  stat_function(fun = function(x) a * x^b, color = "black", size = 0.6,
                inherit.aes = FALSE) +
  geom_vline(xintercept = 20, linetype = "dashed", color = "black") +
  annotate("text", x = 22, y = 4.5,
           label = label_text, hjust = 0, size = 3.5) +
  theme_bw() +
  labs(x = "Length (cm)", y = "Weight (kg)",
       subtitle = "(B)") +
  coord_cartesian(xlim = c(0,60)) +
  scale_colour_manual(values = c("#8da0cb", "#fc8d62")) +
  scale_shape_manual(values = c(15, 17)) +
  theme(legend.position = "none")

a_plot / b_plot
```

## Juvenile diet and weight

*exploration and analysis* 
```{r statistical analysis}
#working df for all diet analyses
Diet <- fitness %>% 
  dplyr::select(Group, Individual, Settlement_diet, 
                Flip_time, Flip_time_seconds, Weight_g, Arm_count, Length_mm) %>%
  drop_na(Settlement_diet, Weight_g)

#diet/weight summary
Diet_weight <- Diet %>% 
  dplyr::select(Group, Individual, Settlement_diet, 
                Flip_time, Flip_time_seconds, Weight_g, Arm_count, Length_mm) %>%
  drop_na(Settlement_diet, Weight_g) %>% 
  group_by(Settlement_diet) %>% 
  summarise(mean_weight = mean(Weight_g),
    sd_weight = sd(Weight_g),
    lower_weight = mean(Weight_g) - sd(Weight_g), 
    upper_weight = mean(Weight_g) + sd(Weight_g))

#working df for length/weight analysis
Diet_size <- Diet %>% 
  dplyr::select(Group, Individual, Settlement_diet, Weight_g, Length_mm) %>%
  drop_na(Settlement_diet, Weight_g, Length_mm) %>%
  group_by(Settlement_diet) %>% 
   mutate(mean_length = mean(Length_mm),
         se_length = sd(Length_mm)/sqrt(length(Length_mm)),
         mean_weight = mean(Weight_g),
         se_weight = sd(Weight_g)/sqrt(length(Weight_g))) 

#fattest and skinniest star
Diet_size %>%
  filter(Weight_g == min(Weight_g)) #skinniest star 0.2g in Regulus/ktchn sink
Diet_size %>%
  filter(Weight_g == max(Weight_g)) #fattest star is 12.1g in Betelgeuse/ktchn sink

#plotted diff in mean length
ggplot(Diet_size, aes(x = Settlement_diet, y = mean_weight, group = 1)) +
  geom_line() +
  geom_point() +
  geom_errorbar(width = 0.1, colour = "red", 
                aes(ymin = mean_weight - se_weight, 
                    ymax = mean_weight + se_weight)) #MLML Urchins signf skinnier (lol)

#descriptive stats
by(Diet_size$Weight_g, Diet_size$Settlement_diet, stat.desc)

#identify outliers, save in object to view whole list
Diet_weight_outliers <- Diet_size %>% 
  group_by(Settlement_diet) %>%
  identify_outliers(Weight_g) #4 outliers, none extreme

#visualization of variation
ggqqplot(Diet_size, "Weight_g") +
  facet_grid(~Settlement_diet) + #multiple outliers in ktchn sink
  theme_bw() 

Diet_size$Settlement_diet <- as.factor(Diet_size$Settlement_diet)

#checking assumption of homogeniety 
leveneTest(Diet_size$Weight_g, 
           Diet_size$Settlement_diet, center = median) 
#appropriate, p = 0.3526

#shapiro test(s)
Diet_size %>% group_by(Settlement_diet) %>%
  shapiro_test(Weight_g) #ktchn sink non norm dist
Diet_size %>%
  ungroup() %>%
  shapiro_test(Weight_g) #overall normality is very inappropriate p < 0.001

#non parametric analysis, given above testing
#KW
dietw_kw <- Diet_size %>%
  ungroup() %>%
  kruskal_test(Weight_g ~ Settlement_diet) 
dietw_kw #p = 0.326

#pairwise comparisons
dietw_ph <- Diet_size %>% 
  ungroup() %>%
  dunn_test(Weight_g ~ Settlement_diet, p.adjust.method = 'bonferroni') 
dietw_ph #all non signf

#work flow for publication-ready plot
dw_ph <- dw_ph %>% arrange(p.adj) #arrange
dw_ph <- add_xy_position(test = dw_ph,
  data = Density_weight_flip_no_outliers %>% as_tibble(.name_repair = "unique"),
  formula = Weight_g ~ Density, fun = "mean", x = "Density")

```

```{r plotting}
#graph, no stats needed due to insignificance from above testing
ggplot(Diet, aes(x = Settlement_diet, y = Weight_g)) +
  stat_boxplot(geom = 'errorbar', width = 0.2) +
  geom_boxplot(colour = "black") +
  geom_jitter(aes(colour = Settlement_diet, shape = Settlement_diet)) +
  labs(x= "Settlement diet", y = "Weight (g)") +
  scale_shape_manual(values = c(3, 8, 4, 11)) +
  scale_colour_brewer(type = 'div', palette = 3) +
  theme_bw() +
  theme(legend.position = 'none')
```

## Juvenile diet and length

*exploration and anlsysis*
```{r statistical analysis}
#working df for length/weight analysis
Diet_size <- Diet %>% 
  dplyr::select(Group, Individual, Settlement_diet, Weight_g, Length_mm) %>%
  drop_na(Settlement_diet, Weight_g, Length_mm) %>%
  group_by(Settlement_diet) %>% 
   mutate(mean_length = mean(Length_mm),
         se_length = sd(Length_mm)/sqrt(length(Length_mm)),
         mean_weight = mean(Weight_g),
         se_weight = sd(Weight_g)/sqrt(length(Weight_g))) 

#fattest and skinniest star
Diet_size %>%
  filter(Length_mm == min(Length_mm)) #smallest star 12mm in Regulus/ktchn sink
Diet_size %>%
  filter(Length_mm == max(Length_mm)) #fattest star is 70mm in Betelgeuse/ktchn sink

#plotted diff in mean length
ggplot(Diet_size, aes(x = Settlement_diet, y = mean_length, group = 1)) +
  geom_line() +
  geom_point() +
  geom_errorbar(width = 0.1, colour = "red", 
                aes(ymin = mean_length - se_length, 
                    ymax = mean_length + se_length)) #MLML Urchins signf smaller

#descriptive stats
by(Diet_size$Length_mm, Diet_size$Settlement_diet, stat.desc)

#identify outliers, save in object to view whole list
Diet_length_outliers <- Diet_size %>% 
  group_by(Settlement_diet) %>%
  identify_outliers(Length_mm) #1 outlier, not extreme

#visualization of variation
ggqqplot(Diet_size, "Length_mm") +
  facet_grid(~Settlement_diet) + #fairly normal across the board
  theme_bw() 

#checking assumption of homogeniety 
leveneTest(Diet_size$Length_mm, 
           Diet_size$Settlement_diet, center = median) 
#appropriate, p = 0.3375

#shapiro test(s)
Diet_size %>% group_by(Settlement_diet) %>%
  shapiro_test(Length_mm) #all norm
Diet_size %>%
  ungroup() %>%
  shapiro_test(Length_mm) #overall normality appropriate p > 0.05

#parametric analysis, given above testing
#anova
dietl_aov <- Diet_size %>%
  ungroup() %>%
  anova_test(Length_mm ~ Settlement_diet) 
dietl_aov #p = 0.208

#pairwise comparisons
dietl_ph <- Diet_size %>% 
  ungroup() %>%
  tukey_hsd(Length_mm ~ Settlement_diet) 
dietl_ph #non signf, all ns
```

*plotting*
```{r plotting}
#graph, no stats
ggplot(Diet, aes(x = Settlement_diet, y = Length_mm)) +
  stat_boxplot(geom = 'errorbar', width = 0.2) +
  geom_boxplot(colour = "black") +
  geom_jitter(aes(colour = Settlement_diet, shape = Settlement_diet)) +
  labs(x= "Settlement diet", y = "Length (mm)") +
  scale_shape_manual(values = c(3, 8, 4, 11)) +
  scale_colour_brewer(type = 'div', palette = 3) +
  theme_bw() +
  theme(legend.position = 'none')
```

## Juvenile diet and flip time

*exploration and anlsysis*
```{r statistical analysis}
#working df for length/weight analysis
Diet_flip <- Diet %>% 
  dplyr::select(Group, Individual, Settlement_diet, Flip_time, Flip_time_seconds) %>%
  drop_na(Settlement_diet, Flip_time) %>%
  group_by(Settlement_diet) %>% 
   mutate(mean = mean(Flip_time_seconds),
         se = sd(Flip_time_seconds)/sqrt(length(Flip_time_seconds))) 

#slowest and fastest flips
Diet_flip %>%
  filter(Flip_time_seconds == min(Flip_time_seconds)) 
#fastest flip 49 secs from Quasar/ktchn sink star
Diet_flip %>%
  filter(Flip_time_seconds == max(Flip_time_seconds))
#slowest flip 30 mins from Electra stars (x2) both ktchn sink 

#flip time means across groups
ggplot(Diet_flip, aes(x = Settlement_diet, y = mean, group = 1)) +
  geom_line() +
  geom_point() +
  geom_errorbar(width = 0.1, colour = "red", 
                aes(ymin = mean - se, 
                    ymax = mean + se)) #juv stars/MLML urchins fast flippers

#descriptive stats
by(Diet_flip$Flip_time_seconds, Diet_flip$Settlement_diet, stat.desc)

#identify outliers, save in object to view whole list
Diet_flip_outliers <- Diet_flip %>% 
  group_by(Settlement_diet) %>%
  identify_outliers(Flip_time_seconds) #8 outliers, 4 extreme

#exclude extreme outliers
Diet_flip_no_outliers <- Diet_flip %>% 
   group_by(Settlement_diet) %>%
   filter(!Flip_time_seconds == 1800,
          !Flip_time_seconds == 1322,
          !Flip_time_seconds == 1235)

#visualization of variation
ggqqplot(Diet_flip_no_outliers, "Flip_time") +
  facet_grid(~Settlement_diet) + #ktchn sink skewed
  theme_bw() 

Diet_flip_no_outliers$Settlement_diet <- as.factor(Diet_flip_no_outliers$Settlement_diet)

#checking assumption of homogeniety 
leveneTest(Diet_flip_no_outliers$Flip_time_seconds, 
           Diet_flip_no_outliers$Settlement_diet, center = median) 
#appropriate, p = 0.3778

#shapiro test(s)
Diet_flip_no_outliers %>% group_by(Settlement_diet) %>%
  shapiro_test(Flip_time_seconds) #juv stars and ktchn sink non normal p < 0.01
Diet_flip_no_outliers %>%
  ungroup() %>%
  shapiro_test(Flip_time_seconds) #extremely non normal, p < 0.000001

#non parametric analysis, given above testing
#kw
dietf_kw <- Diet_flip_no_outliers %>%
  ungroup() %>%
  kruskal_test(Flip_time_seconds ~ Settlement_diet) 
dietf_kw #p = 0.723

#pairwise comparisons
dietf_ph <- Diet_flip_no_outliers %>% 
  ungroup() %>%
  dunn_test(Flip_time_seconds ~ Settlement_diet) 
dietf_ph #non signf, all ns
```

*plotting*
```{r plotting}
#graph, no stats
ggplot(Diet_flip_no_outliers, aes(x = Settlement_diet, y = Flip_time)) +
  stat_boxplot(geom = 'errorbar', width = 0.2) +
  geom_boxplot(colour = "black") +
  geom_jitter(aes(colour = Settlement_diet, shape = Settlement_diet)) +
  labs(x= "Settlement diet", y = "Flip time") +
  scale_shape_manual(values = c(3, 8, 4, 11)) +
  scale_colour_brewer(type = 'div', palette = 3) +
  theme_bw() +
  theme(legend.position = 'none')
```

## Settlement diet and juvenile length and flip time

```{r linear regression}
#working vector
Diet_size_flip <- fitness %>%
  dplyr::select(Group, Individual, Settlement_diet, 
                Length_mm, Flip_time, Flip_time_seconds) %>%
  drop_na(Group, Settlement_diet, Flip_time)

#find and remove outliers, also done in chunk/line (XX)
Diet_size_flip_outliers <- Diet_size_flip %>% 
  group_by(Settlement_diet) %>%
  identify_outliers(Flip_time_seconds) #8 outliers, 4 extreme

#exclude extreme outliers
Diet_size_flip_no_outliers <- Diet_size_flip %>% 
   group_by(Settlement_diet) %>%
   filter(!Flip_time_seconds == 1800,
          !Flip_time_seconds == 1322,
          !Flip_time_seconds == 1235)

Diet_size_flip_no_outliers$Settlement_diet <- as.factor(Diet_size_flip_no_outliers$Settlement_diet)

#simple linear regression (already computed above)
summary(slr)

#interacting model
slr_diet <- lm(Flip_time_seconds ~ Length_mm * Settlement_diet,
                        data = Diet_size_flip_no_outliers)
summary(slr_diet) #p = 0.3

#multiple lr with length + settlement diet
mlr_diet <- lm(Flip_time_seconds ~ Settlement_diet + Length_mm,
                  data = Diet_size_flip_no_outliers)
summary(mlr_diet) #p = 0.08
```

*plotting*
```{r plotting}
#plot with lm line
ggplot(Diet_size_flip_no_outliers, aes(x = Length_mm, y = Flip_time)) +
  geom_smooth(data = Diet_size_flip_no_outliers,
              method = lm, colour = "black", se = TRUE) +
  geom_point(aes(shape = Settlement_diet, colour = Settlement_diet), size = 2) +
  scale_shape_manual(values = c(0, 1, 2, 5)) +
  scale_colour_brewer(type = 'div', palette = 3) +
  labs(x = "Length (mm)",
       y = "Flip time") +
  theme_bw() +
  theme(legend.position = 'top')
```

## Eating trial

*plotting*
```{r}
##Did different diets effect eating?
Eat_diet <- fitness %>% drop_na(Group, Settlement_diet, Eating) %>%
  group_by(Settlement_diet, Eating) %>% 
  tally() %>% 
  mutate(pct = n/sum(n)) 

#organizing Eating times
Eat_diet$Eating <- factor(Eat_diet$Eating, 
                          levels=c('Hour 8', 'Hour 24', 'Hour 48', 'Did not eat'))

#bar plot with percent (proportion) to account for kitchen sink bias
ggplot(Eat_diet) + 
  geom_bar(stat="identity", position="dodge", width = 0.7,
           aes(x=Settlement_diet, y=pct, fill=Eating)) +
  scale_fill_manual(values= c("#7bb662", "#ffd301", "#E03C32", "#A0A0A0")) +
  scale_y_continuous(labels = scales::percent) +
  theme_bw() +
  labs(y = "Juveniles that ate (%)", x = "Settlement diet", fill = "Eating interval") +
  theme(legend.position = 'top')
```

