---
title: "Pycno Data (cleaned)"
author: "Tallulah Story"
date: "2025-06-12"
output: html_document
editor_options: 
  chunk_output_type: console
---
## Setup
```{r packages, include=FALSE}
knitr::opts_chunk$set(echo = T)

#Packages used
library(dplyr)
library(ggplot2)
library(tidyverse)
library(gt)               #Creating tables
library(lubridate)        #Formatting dates and times
library(patchwork)        #Combining plots
library(ggtext)           #Plot aesthetics
library(ggthemes)         #Additional ggplot themes
library(plotly)           #interactive 3d plot (chunk 22)
library(scatterplot3d)    #non-interactive 3d plot (chunk 22)
library(pastecs)          #for descriptive statistics relating to time-series data
library(car)              #statistical analyses/tools
library(ggpubr)           #plots containing statistical information
library(mgcv)             #mixed GAMM computational vehicle
library(marginaleffects)  #compute and plot predictions from models
library(rstatix)          #post-hoc tests for a pipe-friendly interface

getwd() 
setwd("/Users/loganstory/Desktop/SSI/Data")

```


.csv files cleaned and included in GitHub

- Manual cleaning of fitness data included:
    1) Replacement of rogue values (ie. random spaces, random non-numerical items)
    2) Replacing NA with N/A
    3) Excluding any data collected from "unknowns" with just NA
    4) Addition of "eating" and "Eating_score" columns
    5) Replacing incorrect NA values
    
- Manual cleaning of growth data set(s) included: 
    1) Removal of random commas/characters
    2) Removal of missing data points
    
- Manual cleaning of settlement data set(s) included:
    1) Replacing empty cells with 0
    2) Removal of cell totals
    3) Reorganization of counts
    4) Pivot long for totals
    5) Reformatting values
    6) Changing column headers
    7) Additional organization by larval grouping
  
- Manual cleaning of this data set included:
    1) Combining specimen and live star measurements
    2) Removing any data points that had NA measurements/counts

- Manual cleaning of the larval growth data set included:
    1) transitioning from wide to long format
    2) assigning density based on tank (ie. 1 larv/mL to AHQ15)
    3) Grouping together cone and bucket to 15-20 larvae/mL
```{r data import, include=FALSE}
## Fitness data set (experiments from 2/13/25 to 2/25/25)
fitness <- read.csv("Fitness Test.csv")


## Growth data (from measurements and counts taken 12/31/24 to 1/2/25)
growth <- read.csv("Growth Cruves Dec to Feb.csv") #wide format data
growth_combined <- read.csv("growth curves 2.csv") #long format data


## Settlement data (from settlement counts 3/26/24 to 7/9/24)
settlement_raw_wide <- read.csv("Settlement raw counts wide.csv")
settlement_raw_long <- read.csv ("Settlement raw count long.csv")
settlement_total <- read.csv("Settlement total counts.csv")
settlement_cohort <- read.csv("Settlement cupid cohort.csv")
juvenile_settlement <- read.csv("Juvenile settlment survival.csv")


## Specimen Measurements (from Measurements done on 7/7/2025 to 7/9/2025)
specimens <- read.csv("Pycnopodia specimen measurements.csv")
specimens_alive <- read.csv("Pycnopodia specimen measurements - Live Stars.csv")


## Larval growth data collected from ImageJ measurements taken of larvae DPF 4-47
larval_growth <- read.csv("Larval Growth Curves.csv")

```

```{r data cleaning, include = FALSE}
## Fixing measurements, dates, and times with lubridate
#growth data
growth$Measurement_date_1 <- mdy(growth$Measurement_date_1) #jan sampling
class(growth$Measurement_date_1) #check it worked, output should be "date"

growth$Measurement_date_2 <- mdy(growth$Measurement_date_2) #repeat for Feb sampling
class(growth$Measurement_date_2) 

#fitness data
fitness$Flip_time_seconds <- period_to_seconds(hms(fitness$Flip_time)) #time as seconds vs hms

fitness$Flip_time <- hms(fitness$Flip_time)
fitness$Flip_time <- hms::hms(fitness$Flip_time) #time as hours, mins, seconds
class(fitness$Flip_time) #should be hms, difftime

#Re-coding Density variable to make it better for plots
fitness$Densities <- factor(
  fitness$Density, 
  levels = c("1larva/ml", "2larva/ml", "5then2larva/ml", "Bucket"),
  labels = c("1 larvae/mL", "2 larvae/mL", "5 larvae/mL", "15-20 larvae/mL"))

#settlement data
settlement_raw_wide$Date <- mdy(settlement_raw_wide$Date) 
class(settlement_raw_wide$Date) 

settlement_raw_long$Date <- mdy(settlement_raw_long$Date)
class(settlement_raw_long$Date)

#specimen data
specimens$Date_measured <- mdy(specimens$Date_measured)
class(specimens$Date_measured)
```

## Summary tables - fitness
Used summary information here to guide questions, later visualizations, and analyses. 
```{r tables, include=FALSE}
## How many surviving stars were there from each bucket that were tested for the fitness test?
Group_summary_table <- fitness %>% drop_na(Density, Date) %>% #remove NA obvs
  group_by(Group) %>%
  summarise(Individuals = n(), 
            .groups = 'drop') %>% #.groups allows to group by multiple
  arrange(desc(Individuals)) #arranging by groups with most observations to least 

Group_summary_table %>% gt() %>% 
   tab_header(title = md("***P. helianthoides* Tested**"),
               subtitle = md("Number of surviving individuals from each Larval bucket group"))


## Replicates in each treatment group (Density, Settlement Diet, Group)
#Larval density and diet
Replicates_summary <- fitness %>% drop_na(Date, Density) %>%
  group_by(Group, Density, Settlement_diet) %>% 
  summarise(Replicates = n(), 
            .groups = 'drop') %>% 
  arrange(desc(Replicates)) 

Replicates_summary %>% gt() %>% 
  tab_header(title = md("***P. helianthoides* Tested**"),
  subtitle = md("Number of replicates within eat experimental group (Settlement diet and larval density)"))

#Density
Replicates_summary_density <- fitness %>% drop_na(Density) %>%
  group_by(Density) %>% 
  summarise(Replicates = n(), 
            .groups = 'drop') %>% 
  arrange(desc(Replicates)) 

Replicates_summary_density %>% gt() %>% 
  tab_header(title = md("***P. helianthoides* Tested**"),
  subtitle = md("Number of replicates within each density treatment"))

#Settlement diet
Replicates_summary_diet <- fitness %>% drop_na(Density) %>%
  group_by(Settlement_diet) %>% 
  summarise(Replicates = n(), 
            .groups = 'drop') %>% 
  arrange(desc(Replicates))  

Replicates_summary_diet %>% gt() %>% 
  tab_header(title = md("***P. helianthoides* Tested**"),
  subtitle = md("Number of replicates within each settlement diet treatment"))
  


## What diet had the highest count of NOT eating during the entire experiment?
No_eating_summary <- fitness %>% 
  group_by(Settlement_diet, 
           Ate_hour_8, Ate_hour_24, Ate_hours_48) %>% 
  filter(Ate_hour_8 %in% c('No'), 
         Ate_hour_24 %in% c('No'), 
         Ate_hours_48 %in% c('No')) %>% #only looking at NOT eating
  summarise(Count = n(), 
            .groups = 'drop') %>% 
  mutate(Proportion = Count / sum(Count)*100) %>% 
  arrange(desc(Count)) 
No_eating_summary %>% gt()

```


## Plotting -- Flip time (fitness data set)
Exploring results of flip time visually.
```{r DENSITY and FLIP TIME}
#working vector for  density/flip exploration
Density_flip <- fitness %>% drop_na(Densities, Flip_time) 

#without outliers
Density_flip_no_outliers <- fitness %>% 
  drop_na(Density, Flip_time_seconds) %>%
  filter(Flip_time_seconds < 500) #there are 13/87 observations with flip times above 500 seconds

#presentation plot
ggplot(Density_flip, aes(x = Density, y = Flip_time, fill = Density)) +
  stat_boxplot(geom = "errorbar", width = 0.2) +
  geom_boxplot(show.legend = FALSE) +
  labs(x="Larval density", y="Flip time", fill="Larval density") +
  scale_fill_brewer() +
  theme_clean()

#publication worthy plot
ggplot(Density_flip, aes(x = Densities, y = Flip_time)) +
  geom_boxplot() +
  geom_jitter(aes(colour = Densities, shape = Densities), size = 2) +
  stat_boxplot(geom = 'errorbar', width = 0.2) +
  labs(x="Larval density (/mL)", y="Flip time") +
  scale_shape_manual(values = c(15, 16, 17, 18)) +
  theme_clean() +
  theme(legend.position = "none", 
        axis.title.x = element_blank())

#plot (without outliers)
ggplot(Density_flip_no_outliers, aes(x=Density, y=Flip_time_seconds, fill=Density)) +
  stat_boxplot(geom = "errorbar", width = 0.2) +
  geom_boxplot(show.legend = FALSE) +
  labs(x="Larval density", y="Flip time (seconds)", fill="Larval density") +
  scale_fill_brewer() +
  theme_bw()

#descriptive stats
Mean_density_flip <- Density_flip %>% group_by(Density) %>% #w/ outliers
  summarise(mean = (mean(Flip_time)/60))
Mean_density_flip

Mean_density_flip_2 <- Density_flip_no_outliers %>% group_by(Density) %>% #w/out outliers
  summarise(mean = (mean(Flip_time)/60))
Mean_density_flip_2

#largest and smallest flip_time
min(Density_flip$Flip_time) #fastest flip was 49 seconds
max(Density_flip$Flip_time) #slowest flip was 30 minutes
max(Density_flip_no_outliers$Flip_time) #slowest flip w/out outliers was 8 minutes

#violin plot
ggplot(Density_flip, aes(x = Densities, y = Flip_time, fill = Densities)) +
  geom_violin() +
  geom_jitter(aes(shape = Densities)) +
  scale_shape_manual(values = c(15, 16, 17, 18)) +
  theme_bw()
```

```{r DIET and FLIP TIME}
#working vector (w/ outliers)
Settlement_flip <- fitness %>% drop_na(Settlement_diet, Flip_time)

#removing outliers
Settlement_flip_no_outliers <-  fitness %>% drop_na(Settlement_diet, Flip_time_seconds) %>%
  filter(Flip_time_seconds < 500) 

#plot 1, with outliers
ggplot(Settlement_flip, aes(x = Settlement_diet, y = Flip_time, fill = Settlement_diet)) +
  stat_boxplot(geom = "errorbar", width = 0.2) +
  geom_boxplot(show.legend = FALSE) +
  labs(x="Settlement diet", y="Flip time", fill="Settlement Diet") + 
  scale_fill_brewer() +
  theme_bw()

#plot 2, same thing, without outliers
ggplot(Settlement_flip_no_outliers, aes(x=Settlement_diet, y=Flip_time_seconds, fill=Settlement_diet)) + #w/out outliers
  stat_boxplot(geom = "errorbar", width = 0.2) +
  geom_boxplot() +
  labs(x="Settlement diet", y="Flip Time (seconds)", fill="Settlement Diet") +
  scale_fill_brewer() +
  theme_bw()
```

## SICB Poster Plot - LENGTH and FLIP TIME by DENSITY
```{r LENGTH and FLIP TIME POSTER PLOT}
fitness_poster <- ggplot(Density_flip, aes(x = Length_mm, y = Flip_time)) +
  geom_smooth(data = Density_flip, method = lm, formula = y ~ poly(x, 2), 
              color = "red", se = TRUE) +
  geom_point(aes(shape = Densities, colour = Densities), size = 2) +
  scale_shape_manual(values = c(15, 16, 17, 18),
                     name = "Larval Density") +
  scale_color_manual(values = c("#006b3c", "#D2BAAD", "#9F8FAC", "#f3be77"),
                     name = "Larval Density") +
  labs(x = "Length (mm)",
       y = "Flip Time") +
  theme_classic() +
  theme(legend.position = 'right')

ggsave('fitness.png', fitness_poster, bg='transparent')
```

```{r LENGTH and FLIP TIME}
#working data vectors
Length_flip <-  fitness %>% drop_na(Length_mm, Flip_time)

Length_flip_no_outliers <- fitness %>% drop_na(Length_mm, Flip_time_seconds) %>%
  filter(Flip_time_seconds < 500) #no outliers

#plot
ggplot(Length_flip, aes(x=Flip_time, y=Length_mm)) + 
  geom_point() +
  labs(title="Relationship between star length (mm) and Flip time", 
       x="Flip Time", y="Length (mm)") +
  geom_smooth(method=lm , color="red", se=TRUE) + #adding linear trend line and SE
  theme_classic()

#plot, no outliers
ggplot(Length_flip_no_outliers, aes(x=Flip_time_seconds, y=Length_mm)) + 
  geom_point() +
  labs(title="Star Length (mm) and Flip Time (no flip time outliers)", 
       x="Flip Time (seconds)", y="Length (mm)") +
  geom_smooth(method=lm , color="red", se=TRUE) + 
  theme_bw()
```


```{r WEIGHT and FLIP TIME}
#working vector
Weight_flip <-  fitness %>% drop_na(Weight_g, Flip_time) 

#plot
ggplot(Weight_flip, aes(x=Flip_time, y=Weight_g)) +
  geom_point() +
  labs(title="Star Weight (g) and Flip Time", 
       x="Flip Time (seconds)", y="Weight (g)") +
  geom_smooth(method=lm , color="red", se=TRUE) +
  theme_bw()
```


```{r ARM COUNT and FLIP TIME}
Arm_flip <- fitness %>% drop_na(Arm_count, Flip_time) 

#plot
ggplot(Arm_flip, aes(y = Arm_count, x = Flip_time, size = Arm_count)) +
  geom_point(alpha = 0.2) +
  labs(x = "Flip Time", y = "Number of Arms", title = "Flip Time according to the Number of Arms") +
  theme_bw()

```


## Plotting -- Star size metrics (fitness data)
```{r LENGTH and WEIGHT}
#working df
Length_weight <- fitness %>% drop_na(Weight_g, Length_mm, Group)

#plotting by group
ggplot(Length_weight, aes(x=Length_mm, y=Weight_g, colour=Group)) +
  geom_point() +
  geom_smooth(method="lm", 
              formula= y ~ x + I(x^2), #fitting quadratic line
              color="black", size = 0.5, se=TRUE) + 
  labs(x = "Length (mm)", y = "Weight (g)", fill="Larval Bucket Group") +
  theme_bw()

#mean length and weight by group
Avg_length_weight <- fitness %>% 
  drop_na(Weight_g, Length_mm, Group) %>%
  group_by(Group) %>%
  summarise(mean_length = mean(Length_mm),
            mean_weight = mean(Weight_g))

#plotting average length and weight by group
ggplot(Avg_length_weight, aes(x=mean_length, y=mean_weight, colour=Group)) +
  geom_point() +
  geom_smooth(method="lm", formula= y ~ x + I(x^2) ,color="black", size=0.5, se=TRUE) + 
  labs(x = "Length (mm)", y = "Weight (g)", fill="Larval Bucket Group") +
  theme_bw()

#weight and length relationship (expected to be fairly linear/exponential)
ggplot(Length_weight, aes(x = Length_mm, y = Weight_g, 
                          shape = Densities, colour = Densities)) +
  geom_point() +
  scale_shape_manual(values = c(15, 16, 17, 18)) +
  scale_fill_manual(values = c("#92fe6a", "#68ebff", "#936bff", "#d4646c")) +
  theme_bw()

#weight and length relationship with specimens
ggplot(specimens_alive, aes(x = Length_cm, y = Weight_g, 
                            colour = Specimen_type, shape = Specimen_type)) +
  geom_point() +
  scale_shape_manual(values = c(15, 16)) +
  scale_fill_manual(values = c("#92fe6a", "#68ebff")) +
  theme_bw()
```


```{r DENSITY and LENGTH/WEIGHT}
##Did any particular larval density lead to bigger/smaller stars? 

#working data vector for length and larval density 
Density_length <- fitness %>% drop_na(Densities, Length_mm) %>% 
  group_by(Densities) %>% 
  summarise(mean = mean(Length_mm), 
    sd = sd(Length_mm),
    lowerci = mean(Length_mm) - sd(Length_mm), #calculating intervals for error bars
    upper = mean(Length_mm) + sd(Length_mm))

#bar graph (density and length)
ggplot(Density_length) +
  geom_bar(aes(x = Densities, y = mean), 
           stat="identity", fill = "#9F68E2") +
  geom_errorbar(aes(x=Densities, ymin=lowerci, ymax=upper), 
                width=0.3, colour="black", size=0.5) +
  labs(x = "Larval Density", y = "Mean Length (mm)") +
  theme_bw()

#weight and density bar graph
Density_weight <- fitness %>% drop_na(Density, Weight_g) %>% #dropping NA values
  group_by(Density) %>% #summarizing length by Density group
  summarise(
    mean_weight = mean(Weight_g), #being more descriptive with my variable names now, hooray
    sd_weight = sd(Weight_g),
    lower_weight = mean(Weight_g) - sd(Weight_g), 
    upper_weight = mean(Weight_g) + sd(Weight_g))

#density and weight bar graph 
ggplot(Density_weight) +
  geom_bar(aes(x=Density, y=mean_weight), 
           stat="identity", fill = "#8cc5e3") +
  geom_errorbar(aes(x=Density, ymin=lower_weight, ymax=upper_weight), 
                width=0.3, colour="black", size=0.5) + 
  labs(x= "Larval Density", y = "Mean Weight (g)") +
  theme_bw()
```


```{r DIET and LENGTH/WEIGHT}
##Did any particular diet result in bigger/smaller stars?

#LENGTH and DIET bar graph working df
Diet_weight <- fitness %>% 
  drop_na(Settlement_diet, Weight_g) %>% 
  group_by(Settlement_diet) %>% 
  summarise(mean_weight = mean(Weight_g),
    sd_weight = sd(Weight_g),
    lower_weight = mean(Weight_g) - sd(Weight_g), 
    upper_weight = mean(Weight_g) + sd(Weight_g))

#graph
Diet_weight_plot <- ggplot(Diet_weight) +
  geom_bar(aes(x=Settlement_diet, y=mean_weight), 
           stat="identity", fill = "#f55f74") +
  geom_errorbar(aes(x=Settlement_diet, ymin=lower_weight, ymax=upper_weight), 
                width=0.3, colour="black", size=0.5) + #error bars
  labs(x= "Larval Density", y = "Mean Weight (g)") +
  theme_bw()
Diet_weight_plot

#LENGTH and DIET bar graph working df
Diet_length <- fitness %>% 
  drop_na(Settlement_diet, Length_mm) %>% 
  group_by(Settlement_diet) %>% 
  summarise(mean_length = mean(Length_mm),
    sd_length = sd(Length_mm),
    lower_length = mean(Length_mm) - sd(Length_mm), 
    upper_length = mean(Length_mm) + sd(Length_mm))

#graph
ggplot(Diet_length) +
  geom_bar(aes(x=Settlement_diet, y=mean_length), 
           stat="identity", fill = "#0d7d87") +
  geom_errorbar(aes(x=Settlement_diet, ymin=lower_length, ymax=upper_length), 
                width=0.3, colour="black", size=0.5) +
  labs(x= "Larval Density", y = "Mean Length (mm)") +
  theme_clean()
```


```{r EATING and LENGTH/WEIGHT}
##Are smaller stars hungrier? (exploring SIZE and Ate (yes/no))
Eat_length <- fitness %>% drop_na(Group, Length_mm, Eating)

#density plot
ggplot(Eat_length, aes(x=Length_mm, fill=Eating)) +
  geom_density(alpha=0.5) + #alpha=transparency of fill
  labs(title="Star Length and Eating at Hour 8",
       x="Length (mm)",
       y="Frequency") +
  theme_bw()

#histogram with y-axis as count rather than frequency
ggplot(fitness, aes(x = Length_mm, fill = Eating)) +
  geom_histogram(alpha = 0.5, colour = "black", lwd=0.75, linetype = 1) +
  theme_bw()

#4 different histograms of eat, knit together 
hour8 <- fitness %>% dplyr::select(Length_mm, Ate_hour_8) %>%
  filter(Ate_hour_8=='Yes') %>%
  drop_na(Length_mm, Ate_hour_8)

hour24 <- fitness %>% dplyr::select(Length_mm, Ate_hour_24) %>%
  filter(Ate_hour_24=='Yes') %>%
  drop_na(Length_mm, Ate_hour_24)

hour48 <- fitness %>% dplyr::select(Length_mm, Ate_hours_48) %>%
  filter(Ate_hours_48=='Yes') %>%
  drop_na(Length_mm, Ate_hours_48)

no_eat <- fitness %>% dplyr::select(Length_mm, Ate_hours_48) %>%
  filter(Ate_hours_48=='No') %>%
  drop_na(Length_mm, Ate_hours_48)

#knit 4 plots (base r)
par(mfrow=c(2,2)) #2 rows and 2 columns of plots

hist(hour8$Length_mm, xlab = "Count", ylab = "Length (mm)", main = "Stars that ate at hour 8", breaks = 10, ylim = c(0,12), xlim = c(0,70))

hist(hour24$Length_mm, xlab = "Count", ylab = "Length (mm)", main = "Stars that ate at hour 24", breaks = 10, ylim = c(0,12), xlim = c(0,70))

hist(hour48$Length_mm, xlab = "Count", ylab = "Length (mm)", main = "Stars that ate at hour 48", breaks = 10, ylim = c(0,12), xlim = c(0,70))

hist(no_eat$Length_mm, xlab = "Count", ylab = "Length (mm)", main = "Stars that did not eat", breaks = 10, ylim = c(0,12), xlim = c(0,70))

par(mfrow=c(1,1)) #reset plot parameters
```


```{r EATING and DIET}
##Did different diets effect eating?
Eat_diet <- fitness %>% drop_na(Group, Settlement_diet, Eating) %>%
  group_by(Settlement_diet, Eating) %>% 
  tally() %>% 
  mutate(pct = n/sum(n)) 

#organizing Eating times
Eat_diet$Eating <- factor(Eat_diet$Eating, 
                          levels=c('Hour 8', 'Hour 24', 'Hour 48', 'Did not eat'))

#bar plot with raw counts
ggplot(Eat_diet) + 
  geom_bar(stat="identity", position="dodge", width = 0.7,
           aes(x=Settlement_diet, y=n, fill=Eating)) +
  scale_fill_manual(values= c("#7bb662", "#ffd301", "#E03C32", "#A0A0A0")) +
  labs(y = "Juveniles that ate (%)", x = "Settlement diet", fill = "Eating Interval")
  theme_bw() 


#bar plot with percent (proportion) to account for kitchen sink bias
ggplot(Eat_diet) + 
  geom_bar(stat="identity", position="dodge", width = 0.7,
           aes(x=Settlement_diet, y=pct, fill=Eating)) +
  scale_fill_manual(values= c("#7bb662", "#ffd301", "#E03C32", "#A0A0A0")) +
  scale_y_continuous(labels = scales::percent) +
  theme_bw() +
  labs(y = "Juveniles that ate (%)", x = "Settlement diet", fill = "Eating interval") +
  theme(legend.position = 'top')
```


```{r DIET and ARM COUNT}
#Did any particular diet result in stars with more/less arms?
Diet_arm <- fitness %>% drop_na(Settlement_diet, Arm_count) %>%
  group_by(Settlement_diet, Arm_count) %>% 
  summarise(Count = n(), .groups = 'drop')


ggplot(Diet_arm, aes(x = Settlement_diet, y = Arm_count)) +
  geom_bar(stat = "identity")

```


## Plotting -- percent growth (growth data set)

```{r LENGTH and WEIGHT}
#(between January and February)
#relationship between length and weight (December) plot
Length_weight_plot_1 <- ggplot(growth, aes(x=Length_mm_1, y=Weight_g_1)) +
  geom_point() +
  geom_smooth(method=lm , color="red", se=TRUE) + 
  labs(title = "January 2025", x = "Length (mm)", y = "Weight (g)") +
  theme_bw()
Length_weight_plot_1

#relationship between length and weight (February) plot
Length_weight_plot_2 <- ggplot(growth, aes(x=Length_mm_2, y=Weight_g_2)) +
  geom_point() +
  geom_smooth(method=lm , color="red", se=TRUE) + #add trend line and CE
  labs(title = "February 2025", x = "Length (mm)", y = "Weight (g)") +
  theme_bw()
Length_weight_plot_2

#patchwork, compare
Length_weight_plot_1 + Length_weight_plot_2

#combine plot, by month
ggplot(growth_combined, aes(x = Length_mm, y = Weight_g, colour = Measurement_month)) +
  geom_point() +
  geom_smooth(method = lm, colour="black", se=TRUE) +
  labs(x = "Length (mm)", y = "Weight (g)", title = "Size Measurements (January and February)") +
  scale_colour_manual(values = c("#f55f74", "#8cc5e3")) +
  theme_clean()
```


```{r pct GROWTH and DIET}
#percent growth from DEC to FEB
Percent_growth_summary <- growth %>% 
  dplyr::select(Settlement_diet, 
                Length_mm_1, Length_mm_2, 
                Measurement_date_1, Measurement_date_2) %>% 
  drop_na() %>% 
  group_by(Settlement_diet) %>% 
  mutate(Pct_growth = ((Length_mm_2 - Length_mm_1) / Length_mm_1)) %>% #calc percent growth
  summarize(mean_growth_length = mean(Pct_growth),
            min_growth = min(Pct_growth),
            n_obvs = length(Pct_growth),
            stdev = sd(Pct_growth, na.rm=T))

#Re-mutate and add additional columns (only way I could get this to work)
Percent_growth_summary2 <- Percent_growth_summary %>% 
  mutate(se = stdev/sqrt(n_obvs), #standard error equation: σ (stdev) / √n (# of obs)
         lower_ci = mean_growth_length - (se), #upper and lower "confidence interval"
         upper_ci = mean_growth_length + (se)) #not an actual confidence interval, I just use this shorthand to keep track of my upper and lower variables when graphing! 

#graph mean percent growth by diet group from DEC to FEB (with error bars!)
growth_length_plot <- ggplot(Percent_growth_summary2) +
  geom_bar(aes(x=Settlement_diet, y=mean_growth_length), 
           stat="summary", fill= "#8cc5e3") +
  geom_errorbar(aes(x=Settlement_diet, ymin=lower_ci, ymax=upper_ci), 
                width=0.4, colour = "black", alpha=1, size=0.5) + #error bar aesthetics <3
  scale_y_continuous(labels = scales::percent) + #scaling to percentage on y-axis
  labs(x = "Settlement Diet", y = "Mean Percent Length Growth (mm)") +
  theme_bw()
growth_length_plot

#comparing WEIGHT as the growth factor using the same exact methodology/coding as above
Percent_growth_summary_weight <- growth %>% 
  dplyr::select(Settlement_diet, 
                Weight_g_1, Weight_g_2, 
                Measurement_date_1, Measurement_date_2) %>% 
  drop_na() %>% 
  group_by(Settlement_diet) %>% 
  mutate(Pct_growth = ((Weight_g_2 - Weight_g_1) / Weight_g_1)) %>% 
  summarize(mean_growth_weight = mean(Pct_growth),
            min_growth = min(Pct_growth),
            n_obvs = length(Pct_growth),
            stdev = sd(Pct_growth, na.rm=T)) #growth mean by Settlement diet

#Re-mutate and add additional columns (a NOT parsimonious way to organize this data... oops)
Percent_growth_summary_weight2 <- Percent_growth_summary_weight %>% 
  mutate(se = stdev/sqrt(n_obvs),
         lower_ci = mean_growth_weight - (se), 
         upper_ci = mean_growth_weight + (se))  

#graph mean percent growth by diet group from DEC to FEB (with error bars!)
growth_weight_plot <- ggplot(Percent_growth_summary_weight2) +
  geom_bar(aes(x=Settlement_diet, y=mean_growth_weight), 
           stat="summary", fill ="#f55f74") +
  geom_errorbar(aes(x=Settlement_diet, ymin=lower_ci, ymax=upper_ci), 
                width=0.4, colour = "black", alpha=1, size=0.5) + 
  scale_y_continuous(labels = scales::percent) +
  labs(x = "Settlement Diet", y = "Mean Percent Weight Growth (g)") +
  theme_bw()
growth_weight_plot

#combine and compare weight and length plots
growth_length_plot + growth_weight_plot
```


```{r WEIGHT/LENGTH and ARM COUNT}
#length plots
ggplot(growth_combined, aes(x = Length_mm, y = Arm_count, colour = Measurement_month)) +
  geom_point() +
  labs(x = "Lenght (mm)", y = "Number of Arms", colour = "Measurement Month") +
  scale_colour_manual(values = c("#f55f74", "#8cc5e3")) +
  theme_bw()

#points where flip time is higher than 500 seconds
fitness$Highlight <- ifelse(fitness$Flip_time_seconds >= 500, ">=500s", "<500s")

#plot where flip time outliers are highlighted in 
ggplot(fitness, aes(x = Length_mm, y = Arm_count, colour = Highlight)) +
  geom_point() +
  labs(x = "Length (mm)", y = "Number of Arms", colour = "Flip Time") +
  scale_colour_manual(values = c(">=500s" = "red", "<500s" = "black")) +
  theme_bw()

#trend lines, by month
trend_1 <- growth_combined[1:98,] #january measurements
trend_2 <- growth_combined[99:196,] #febraury measurements

ggplot(growth_combined, aes(x = Length_mm, y = Arm_count, colour = Measurement_month)) +
  geom_point() +
  geom_smooth(data = trend_1, 
              aes(x=Length_mm, y = Arm_count), method = "lm", se = TRUE, color = "lightblue") +
  geom_smooth(data = trend_2, 
              aes(x=Length_mm, y = Arm_count), method = "lm", se = TRUE, color = "tomato") +
  labs(x = "Length (mm)", y = "Number of Arms", colour = "Measurement Month") +
  scale_colour_manual(values = c("tomato", "lightblue")) +
  theme_bw()

#weight plots
ggplot(growth_combined, aes(x = Weight_g, y = Arm_count, shape = Measurement_month)) +
  geom_point() +
  labs(x = "Weight_g", y = "Number of Arms", shape = "Measurement Month") +
  theme_bw()

#trend lines, by month
ggplot(growth_combined, aes(x = Weight_g, y = Arm_count, shape = Measurement_month)) +
  geom_point() +
  geom_smooth(data = trend_1, 
              aes(x=Weight_g, y = Arm_count), method = "lm", se = TRUE, color = "blue") +
  geom_smooth(data = trend_2, 
              aes(x=Weight_g, y = Arm_count), method = "lm", se = TRUE, color = "red") +
  labs(x = "Weight (g)", y = "Number of Arms", shape = "Measurement Month") +
  scale_colour_manual(values = c("red", "blue")) +
  theme_clean()
```




## Plotting -- Specimen measurements data
```{r LENGTH and ARM COUNT}
##Does the length of a star determine how many arms it has? At what length do arms stop developing? 
specimens <- read.csv("Pycnopodia specimen measurements.csv")

#Re-ordering for plot aesthetics
specimens$Specimen_type <- factor(
  specimens$Specimen_type, 
  levels = c("Cupid Cohort", "Aquarium display", "Specimen"), 
  labels = c("Cupid cohort", "Alaskan stars", "Museum specimens"))

#length and arms by specimen type
ggplot(specimens, aes(x = Length_cm, y = Arm_count, 
                      shape = Specimen_type, colour = Specimen_type)) +
  geom_point() +
  scale_colour_manual(values = c("#66c2a5", "#fc8d62", "#8da0cb")) + 
  ylim(5, 25) +  #ylim 25
  labs(x = "Length (cm)", y = "Number of Arms", 
       shape = "Specimen type", 
       colour = "Specimen type") +
  theme_bw() +
  theme(legend.title = element_markdown())
```


```{r LENGTH and WEIGHT}
##Are weight and length completely linear between different age classes (juvenile vs. adult)

#regular y-axis
ggplot(specimens_alive, aes(x = Length_cm, y = Weight_kg,
                            colour = Specimen_type, shape = Specimen_type)) +
  geom_point() +
  scale_colour_manual(values = c("#66c2a5", "#fc8d62")) + 
  labs(x = "Length (cm)", y = "Weight (kg)", 
       shape = "*P. helianthoides* specimen type", 
       colour = "*P. helianthoides* specimen type") + 
  theme_bw() +
  theme(legend.title = element_markdown())

#log transform y-axis
ggplot(specimens_alive, aes(x = Length_cm, y = Weight_kg, 
                            colour = Specimen_type, shape = Specimen_type)) +
  geom_point() +
  scale_y_log10(breaks = c(0, 0.01, 0.05, 0.1, 0.25, 0.5, 1, 2, 4, 6), 
                labels = scales::comma_format())  +
  scale_colour_manual(values = c("#66c2a5", "#fc8d62")) + 
  geom_smooth(aes(group = 1), color = "red", size = 0.7, se = TRUE) +
  labs(x = "Length (cm)", y = "Weight (kg, log scale)", 
       shape = "*P. helianthoides* specimen type", 
       colour = "*P. helianthoides* specimen type") + 
  theme_bw() +
  theme(legend.title = element_markdown())
```


```{r intercative 3d plot LENGTH/WEIGHT and ARM COUNT}
#interactive 3-d plot
plot_ly(data = specimens_alive, x = ~Weight_kg, y = ~Length_cm, z = ~Arm_count, 
        type = "scatter3d", mode = "markers", color = ~Specimen_type, 
        symbol = ~Specimen_type, 
        colors = c("Cupid cohort" = "#66c2a5", "Alaskan stars" = "#fc8d62"),
        marker = list(size = 5)) %>%
  layout(scene = list(xaxis = list(title = "Weight (kg)"), yaxis = list(title = "Length (cm)"),zaxis = list(title = "Number of Arms")))

test <- specimens_alive

#shapes by specimen type
shapes = c(16, 17) 
shapes <- shapes[as.numeric(test$Specimen_type)]

#color by specimen type
colors = c("#66c2a5", "#fc8d62")
colors <- colors[as.factor(test$Specimen_type)]

#normal, publishable 3-d plot
cool_plot <- scatterplot3d(x = test$Length_cm, y = test$Weight_kg, 
              z = test$Arm_count, color = colors, pch = shapes,
              xlab = "Weight (kg)", ylab = "Length (cm)", zlab = "Number of Arms", 
              angle = 45)

#add a legend
legend(cool_plot$xyz.convert(10, 5, 9.5), legend = c("Alaskan stars", "Cupid cohort"),  
       col = c("#66c2a5", "#fc8d62"), pch = c(16, 17), title = "Specimen Type") 
```


```{r WEIGHT and ARMS}
##As above, at what weight does limb development slow? How does this compare in juvnile vs. full-grown stars?
specimens_alive <- read.csv("Pycnopodia specimen measurements - Live Stars.csv")

#Re-ordering for plot aesthetics
specimens_alive$Specimen_type <- factor(
  specimens_alive$Specimen_type, 
  levels = c("Cupid Cohort", "Aquarium display"), 
  labels = c("Cupid cohort", "Alaskan stars"))

#Weight and arms plot (aquarium stars and cupid cohort only)
ggplot(specimens_alive, aes(x = Weight_g, y = Arm_count, 
                            colour = Specimen_type, shape = Specimen_type)) +
  geom_point() +
  geom_smooth(aes(group = 1), color = "red", size = 0.7, se = TRUE) +
  scale_colour_manual(values = c("#66c2a5", "#fc8d62")) + 
  labs(x = "Weight (g)", y = "Number of Arms", 
       shape = "*P. helianthoides* specimen type", 
       colour = "*P. helianthoides* specimen type") + 
  theme_bw() +
  theme(legend.title = element_markdown())

#trying the plot with weight in kg to see if visualization improves
specimens_alive$Weight_kg <- specimens_alive$Weight_g /1000

#re-plot, scale x accordingly
ggplot(specimens_alive, aes(x = Weight_kg, y = Arm_count, 
                            colour = Specimen_type, shape = Specimen_type)) +
  geom_point() +
  scale_x_log10(breaks = c(0, 0.01, 0.05, 0.1, 0.25, 0.5, 1, 2, 4, 6), 
                labels = scales::comma_format())  +
  scale_colour_manual(values = c("#66c2a5", "#fc8d62")) + 
  geom_smooth(aes(group = 1), color = "red", size = 0.7, se = TRUE) +
  labs(x = "Weight (kg, log scale)", y = "Number of Arms", 
       shape = "*P. helianthoides* specimen type", 
       colour = "*P. helianthoides* specimen type") + 
  coord_cartesian(ylim = c(5, 25)) +
  theme_bw() +
  theme(legend.title = element_markdown())
```




## Non-linear asymptotic model
- This model assumes that arm count increases with length, but eventually levels off. This is based on visual analysis of the graph above, which is the rationale for why I chose to do a self-starting non-linear model.
  - Arm count begins near 3.4, grows rapidly, then levels out at ~19 arms. 
  - lrc is at a constant of -1.9 as it nears the asymptote. 
  - All three parameters are significant ([Asym p = <2e-16], [R0 p = 5.2e-06], [lrc p = <2e-16]) and the model confidently identifies a saturating relationship between arm count and length, while each parameter contributes meaningfully to the fit of the model. 
  
  - A convergence over 6 iterations means the model converges quickly and reliably. Interpretation: As P. helianthoides individuals grow, their number of arms increases quickly early in development but eventually plateaus near 19 arms. The model fits this relationship well, with statistically significant parameters and low prediction error.
```{r model and plotting}
Arm_asymp_model <- nls(Arm_count ~ SSasymp(Length_cm, Asym, R0, lrc), data = specimens)

summary(Arm_asymp_model)
coef(Arm_asymp_model) 

#creating seq of 100 evenly spaced body lengths in the observed body size range 
length_seq <- seq(min(specimens$Length_cm), max(specimens$Length_cm), length.out = 100)

#create a prediction grid for model from length_seq, usable for later model predictions
pred_grid <- data.frame(Length_cm = length_seq)

#bootstrap model fits
set.seed(123)  #for reproducibility, random number seed

#bootstrap model replicates
n_boot <- 1000

#bootstrap replicate
boot_preds <- replicate(n_boot, { #drawing bootstrap sample w/ replacement
  boot_sample <- specimens[sample(1:nrow(specimens), replace = TRUE), ]
  tryCatch({ #fitting model to each bootstrap sample, predict arm count across all lengths
    boot_model <- nls(Arm_count ~ SSasymp(Length_cm, Asym, R0, lrc), data = boot_sample)
    predict(boot_model, newdata = pred_grid)
  }, error = function(e) rep(NA, length(length_seq)))
})

#transpose, remove failed fits (NAs)
boot_preds <- t(boot_preds) #transposes, rows=bootstrap reps; columns=predictions for length
boot_preds <- boot_preds[complete.cases(boot_preds), ] #remove failed fits

#calculate mean, conf interval
pred_grid$mean <- apply(boot_preds, 2, mean) #mean prediction at each length
pred_grid$lower <- apply(boot_preds, 2, quantile, probs = 0.025) #2.5th percentile confint
pred_grid$upper <- apply(boot_preds, 2, quantile, probs = 0.975) #97.5th percentile confint

#extract model coefs, rounds for notation and reporting on graph
coef_vals <- coef(Arm_asymp_model) #saving model coefs in coef_vals
asym <- round(coef_vals["Asym"], 2)
r0 <- round(coef_vals["R0"], 2)
lrc <- round(coef_vals["lrc"], 2)

#creating psuedo R2 from residuals of model and prediction grid
y_pred <- predict(Arm_asymp_model) #predicted y values from model
y_obs <- specimens$Arm_count #the actual observed arm counts
rss <- sum((y_obs - y_pred)^2)  #total model error (diff in observed vs predicted values)
tss <- sum((y_obs - mean(y_obs))^2)  #total variability in data from the mean

#psuedo R2, returning as a result of nrs() not natively calculating this for me :(
r_squared <- 1 - (rss / tss)
round(r_squared, 3) #pretty great fit, psuedo R2 = 0.881

#annotate with summary stats
label_text <- paste0(
  "Asym = ", asym, "\n",
  "R0 = ", r0, "\n",
  "lrc = ", lrc, "\n",
  "Pseudo R² = ", round(r_squared, 3))

#plot, length and arms by specimen type with self-starting asymp model line
asym_poster <- 
  ggplot(specimens, aes(x = Length_cm, y = Arm_count)) +
  geom_point(aes(colour = Specimen_type, shape = Specimen_type)) + #aesthetics by specimen
  geom_line(data = pred_grid, aes(x = Length_cm, y = mean), color = "black", size = 0.7) +
  geom_ribbon(data = pred_grid, aes(x = Length_cm, ymin = lower, ymax = upper),
              alpha = 0.2, fill = "gray",
              inherit.aes = FALSE) + #override aes mapping in call to not inherit aes
  annotate("text", x = 20, y = 10, 
           label = label_text, hjust = 0, size = 3.5) + #adding summary stats to plot
  scale_colour_manual(values = c("#66c2a5", "#fc8d62", "#8da0cb")) +
  coord_cartesian(ylim = c(5, 25)) +
  labs(x = "Length (cm)", y = "Number of Arms", shape = "Specimen type",
    colour = "Specimen type") +
  theme_classic() +
  theme(legend.title = element_markdown(),
        legend.position = 'top') #italicize in legend title

ggsave('asym_model.png', asym_poster, bg='transparent')

#plot no text
ggplot(specimens, aes(x = Length_cm, y = Arm_count)) +
  geom_point(aes(colour = Specimen_type, shape = Specimen_type)) + #aesthetics by specimen
  geom_line(data = pred_grid, aes(x = Length_cm, y = mean), color = "black", size = 0.7) +
  geom_ribbon(data = pred_grid, aes(x = Length_cm, ymin = lower, ymax = upper),
              alpha = 0.2, fill = "gray",
              inherit.aes = FALSE) + #override aes mapping in call to not inherit aes
  scale_colour_manual(values = c("#66c2a5", "#fc8d62", "#8da0cb")) +
  coord_cartesian(ylim = c(5, 25), xlim = c(0,60)) +
  labs(x = "Length (cm)", y = "Number of Arms", shape = "Specimen type",
    colour = "Specimen type") +
  theme_bw() +
  theme(legend.title = element_markdown()) #italicize in legend title
```



## Non-linear allometric growth model, W=a x L^b
- Where:
    W is the weight
    L is the length
    a is a constant that varies between 
    b the scaling exponent 
        - which dictates how weight changes with length
        - b is also known as the allometric constant
        - b = 3 indicates isometric growth
        - 'b<3' means negative allometry
        - 'b>3' positive allometry

```{r model and plotting}
#model
lm_log <- lm(log(Weight_kg) ~ log(Length_cm), data = specimens_alive)
summary(lm_log)

#back-transform coefficients
a <- exp(coef(lm_log)[1])
b <- coef(lm_log)[2]
b

# Labels
a_label <- formatC(a, format = "e", digits = 2)
b_label <- round(b, 2)
a #4.921003e-05 
b #2.846814, indicates negative allometry

#bootstrapping
set.seed(123)

#number of bootstraps 
n_boot <- 1000

#storing boot coefs
boot_coefs <- map_dfr(1:n_boot, function(i) {
  boot_sample <- specimens_alive %>% sample_frac(replace = TRUE) #re-sample
  boot_model <- lm(log(Weight_kg) ~ log(Length_cm), data = boot_sample) #fit model
  tibble( #extract and return coefs
    iter = i,
    intercept = coef(boot_model)[1],
    slope = coef(boot_model)[2]
  )
})

#back transform coefs
boot_coefs <- boot_coefs %>%
  mutate(a = exp(intercept),
         b = slope)

#summarize w/ confidence intervals
boot_summary <- boot_coefs %>%
  summarise(
    a_mean = mean(a),
    a_ci_lower = quantile(a, 0.025),
    a_ci_upper = quantile(a, 0.975),
    b_mean = mean(b),
    b_ci_lower = quantile(b, 0.025),
    b_ci_upper = quantile(b, 0.975))

#summary, printing
print(boot_summary)

#sequence of Length values to predict over
length_seq <- seq(min(specimens_alive$Length_cm), max(specimens_alive$Length_cm), length.out = 100)

#generate predicted weight curves for each bootstrap replicate
boot_lines <- boot_coefs %>%
  mutate(iter = row_number()) %>%
  group_by(iter) %>%
  do({
    data.frame(
      Length_cm = length_seq,
      Weight_kg = exp(.$intercept + .$slope * log(length_seq)))
  })

boot_lines <- map_dfr(1:nrow(boot_coefs), function(i) {
  data.frame(
    Length_cm = length_seq,
    Weight_kg = exp(boot_coefs$intercept[i] + boot_coefs$slope[i] * log(length_seq)),
    iter = i)
  })

#predicted and observed values
y_pred <- predict(lm_log)
y_obs <- log(specimens_alive$Weight_kg)

#calculate pseudo-R²
rss <- sum((y_obs - y_pred)^2)
tss <- sum((y_obs - mean(y_obs))^2)
r_squared <- 1 - (rss / tss)
r2_label <- round(r_squared, 3)

#precomputed back-transformed values
a_label <- formatC(a, format = "e", digits = 2)
b_label <- round(b, 2)

label_text <- paste0(
  "Weight = ", a_label, " × Length^", b_label, "\n",
  "Pseudo R² = ", r2_label
)

#adding bootstrapped curves to my plot
ggplot(specimens_alive, aes(x = Length_cm, y = Weight_kg, 
                            colour = Specimen_type, shape = Specimen_type)) +
  geom_point() +
  geom_line(data = boot_lines, aes(x = Length_cm, y = Weight_kg),
            inherit.aes = FALSE, alpha = 0.2, size = 0.2, color = "grey") +
  stat_function(fun = function(x) a * x^b, color = "black", size = 0.6,
                inherit.aes = FALSE) +
  annotate("text", x = 6, y = 2.5,
           label = label_text, hjust = 0, size = 3.5) +
  theme_bw() +
  labs(x = "Length (cm)", y = "Weight (kg)") +
  coord_cartesian(xlim = c(0,60)) +
  scale_colour_manual(values = c("#66c2a5", "#fc8d62")) +
  theme(legend.position = "none")

#non linear model, no text
ggplot(specimens_alive, aes(x = Length_cm, y = Weight_kg, 
                            colour = Specimen_type, shape = Specimen_type)) +
  geom_point() +
  geom_line(data = boot_lines, aes(x = Length_cm, y = Weight_kg),
            inherit.aes = FALSE, alpha = 0.2, size = 0.2, color = "grey") +
  stat_function(fun = function(x) a * x^b, color = "black", size = 0.6,
                inherit.aes = FALSE) +
  theme_clean() +
  labs(x = "Length (cm)", y = "Weight (kg)") +
  coord_cartesian(xlim = c(0,60)) +
  scale_colour_manual(values = c("#66c2a5", "#fc8d62")) +
  theme(legend.position = "none")
```




## Statistical analyses (fitness data)
Packages used:
    1) "pastecs" for stat.desc() descriptive statistics in relation to time series (flip time) data
    2) "car" for leveneTest()
    3) "rstatix" for dunn_test() and tukey_hsd() post-hoc anal
    4) "WRS2" for med1way() performing heteroscedastic one-way ANOVA for medians in robust ANOVA testing 
    5) base r "stats" for kruskal.test() and aov() for primary statistical approaches
        
```{r DIET and FLIP TIME}
##Is there a difference in the mean Flip_time between different settlement diets?

Diet_flip <- fitness %>% 
  drop_na(Settlement_diet, Flip_time_seconds) %>%
  group_by(Settlement_diet) %>% 
   mutate(mean_time = mean(Flip_time_seconds),
         se = sd(Flip_time_seconds)/sqrt(length((Flip_time_seconds))))

Diet_flip_no_outliers <- fitness %>% 
  drop_na(Settlement_diet, Flip_time_seconds) %>%
  filter(Flip_time_seconds < 500) %>%
  group_by(Settlement_diet) %>% 
  mutate(mean_time = mean(Flip_time_seconds),
         se = sd(Flip_time_seconds)/sqrt(length((Flip_time_seconds))))

#visually checking normality of FLIP TIME
hist(Diet_flip$Flip_time_seconds) #right skew, non normal dist
qqnorm(Diet_flip$Flip_time_seconds) #right skew, light tailed
qqline(Diet_flip$Flip_time_seconds)

#visually check differences in mean
ggplot(Diet_flip, aes(x = Settlement_diet, y = mean_time, group = 1)) +
  geom_line() +
  geom_point() +
  geom_errorbar(width = 0.1, colour = "blue", 
                aes(ymin = mean_time - se, ymax = mean_time + se)) +
  labs(x = "Settlement Diet", y = "Flip Time (seconds)") +
  theme_bw()

#test normality
shapiro.test(Diet_flip$Flip_time_seconds) #p = 1.838e-12, not normal

#descriptive stats
by(Diet_flip$Flip_time_seconds, Diet_flip$Settlement_diet, stat.desc)

#checking variances
leveneTest(Diet_flip$Flip_time_seconds, Diet_flip$Settlement_diet, center = median)
##appropriate: p = 0.4559, fulfills assumptions

#checking normality visually (again)
ggplot(Diet_flip, aes(sample = Flip_time_seconds)) +
     geom_qq() +
     geom_qq_line() +
     facet_wrap(~Settlement_diet, scales = "free_y")
##non-normal in ALL groups

#shapiro.test by group
test <- Diet_flip %>% dplyr::select(Settlement_diet, Flip_time_seconds) %>%
  filter(Settlement_diet == "Juv. stars")
shapiro.test(test$Flip_time_seconds) #juv stars p = 0.01525, NON NORMAL

test <- Diet_flip %>% dplyr::select(Settlement_diet, Flip_time_seconds) %>%
  filter(Settlement_diet == "Kitchen sink")
shapiro.test(test$Flip_time_seconds) #kitchen sink, p = <0.01, NON NORMAL

test <- Diet_flip %>% dplyr::select(Settlement_diet, Flip_time_seconds) %>%
  filter(Settlement_diet == "MLML Urchins")
shapiro.test(test$Flip_time_seconds) #MLML urchins, p = 0.4, normal

test <- Diet_flip %>% dplyr::select(Settlement_diet, Flip_time_seconds) %>%
  filter(Settlement_diet == "Worms & Diatoms")
shapiro.test(test$Flip_time_seconds) #worms and diatoms, p = 0.01
##3/4 non-normal, proceeding with kruskal-wallis 

par(mfrow = c(1, 1)) #reset plot params

#kruskal-wallis 
Diet_flip_kruskal <- kruskal.test(Flip_time_seconds ~ Settlement_diet, data = Diet_flip)
Diet_flip_kruskal #p = 0.5087

#post-hoc tests
Diet_flip_posthoc <- dunn_test(data = fitness, Flip_time_seconds ~ Settlement_diet, p.adjust.method = "bonferroni")
Diet_flip_posthoc
Diet_flip_posthoc <- Diet_flip_posthoc %>% arrange(p.adj)
Diet_flip_posthoc <- Diet_flip_posthoc %>% add_xy_position(x = "Settlement_diet")

#plot 
ggboxplot(Diet_flip, x = "Settlement_diet", y = "Flip_time_seconds",
          color = "Settlement_diet", add = "jitter", shape = "Settlement_diet") + 
  stat_pvalue_manual(Diet_flip_posthoc, label = "p.adj.signif",hide.ns = TRUE) +
  stat_compare_means(label.y = 2000) + #add global p-value from KW
  labs(x = "Settlement Diet", y = "Flip Time (seconds)") +
  theme(legend.position = "none")
```
A Kruskal-Wallis non-parametric test was performed to compare the effect of settlement diet on star flip time performed during the February fitness test. The Kruskal-Wallis test revealed that there WAS NOT a statistically significant difference in flip time between at least two groups (KW Chi2(3) = 2.32, p = 0.5087)). A non-parametric Dunn's Test for multiple comparisons found that the mean value of flip time was not significantly different between or within any of the diet groups. 


```{r DIET and FLIP TIME no outliers}
##Is there a difference in the mean Flip_time between different settlement diets when excluding the 13 flip time outliers?

Diet_flip_no_outliers <- fitness %>% drop_na(Settlement_diet, Flip_time_seconds) %>%
  filter(Flip_time_seconds < 500) %>%
  group_by(Settlement_diet) %>% 
  mutate(mean_time = mean(Flip_time_seconds),
         se = sd(Flip_time_seconds)/sqrt(length((Flip_time_seconds))))

#visually checking normality of FLIP TIME
hist(Diet_flip_no_outliers$Flip_time_seconds) #mostly normal
qqnorm(Diet_flip_no_outliers$Flip_time_seconds) #light tailed skew
qqline(Diet_flip_no_outliers$Flip_time_seconds)

#visually check differences in mean
ggplot(Diet_flip_no_outliers, aes(x = Settlement_diet, y = mean_time, group = 1)) +
  geom_line() +
  geom_point() +
  geom_errorbar(width = 0.1, colour = "blue", 
                aes(ymin = mean_time - se, ymax = mean_time + se)) +
  labs(x = "Settlement Diet", y = "Flip Time (≤500 seconds)") +
  theme_bw()

#test normality
shapiro.test(Diet_flip_no_outliers$Flip_time_seconds) #p = 0.0035, still not normal

#descriptive stats
by(Diet_flip_no_outliers$Flip_time_seconds, Diet_flip_no_outliers$Settlement_diet, stat.desc)

#checking variances
leveneTest(Diet_flip_no_outliers$Flip_time_seconds, Diet_flip_no_outliers$Settlement_diet, center = median)
##appropriate: p = 0.589, fulfills assumptions

#checking normality (again)
ggplot(Diet_flip_no_outliers, aes(sample = Flip_time_seconds)) +
     geom_qq() +
     geom_qq_line() +
     facet_wrap(~Settlement_diet, scales = "free_y")
##deviation in all, light skew in all groups

#shapiro.test by group
test <- Diet_flip_no_outliers %>% dplyr::select(Settlement_diet, Flip_time_seconds) %>%
  filter(Settlement_diet == "Juv. stars")
shapiro.test(test$Flip_time_seconds) #juv stars p = 0.3313

test <- Diet_flip_no_outliers %>% dplyr::select(Settlement_diet, Flip_time_seconds) %>%
  filter(Settlement_diet == "Kitchen sink")
shapiro.test(test$Flip_time_seconds) #kitchen sink, p = <0.01, NON NORMAL

test <- Diet_flip_no_outliers %>% dplyr::select(Settlement_diet, Flip_time_seconds) %>%
  filter(Settlement_diet == "MLML Urchins")
shapiro.test(test$Flip_time_seconds) #MLML urchins, p = 0.4, normal

test <- Diet_flip_no_outliers %>% dplyr::select(Settlement_diet, Flip_time_seconds) %>%
  filter(Settlement_diet == "Worms & Diatoms")
shapiro.test(test$Flip_time_seconds) #worms and diatoms, p = 0.938
##3/4 are normal, however given between group SW rejecttion, proceeding with kruskal-wallis 

par(mfrow = c(1, 1)) #reset plot params

#kruskal-wallis 
Diet_flip_no_kruskal <- kruskal.test(Flip_time_seconds ~ Settlement_diet, data = Diet_flip_no_outliers)
Diet_flip_no_kruskal #p = 0.5898

#ungroup
Diet_flip_no_outliers <- Diet_flip_no_outliers %>% ungroup()

#post-hoc tests
Diet_flip_no_posthoc <- dunn_test(data = Diet_flip_no_outliers, Flip_time_seconds ~ Settlement_diet, p.adjust.method = "bonferroni")
Diet_flip_no_posthoc
Diet_flip_no_posthoc <- Diet_flip_no_posthoc %>% arrange(p.adj)
Diet_flip_no_posthoc <- Diet_flip_no_posthoc %>% add_xy_position(x = "Settlement_diet")

### **NEED TO FIX, code currently NOT WORKING ----
ggboxplot(test, x = "Settlement_diet", y = "Flip_time_seconds",
          color = "Settlement_diet", add = "jitter", shape = "Settlement_diet") +
  stat_pvalue_manual(Diet_flip_no_posthoc, label = "p.adj.signif", hide.ns = TRUE) +
  stat_compare_means(label.y = 1000) + #add global p-value from KW
  labs(x = "Settlement Diet", y = "Flip Time (seconds)")
```
A Kruskal-Wallis non-parametric test was performed to compare the effect of settlement diet on star flip time performed during the February fitness test. The Kruskal-Wallis test revealed that there WAS NOT a statistically significant difference in flip time between at least two groups (KW Chi2(3) = 2.32, p = 0.5087)). A non-parametric Dunn's Test for multiple comparisons found that the mean value of flip time was not significantly different between or within any of the diet groups.


```{r DENSITY and FLIP TIME}
## Is there a difference between larval density groups and flip time?

#working data frame for this question:
Density_flip <- fitness %>% drop_na(Densities, Flip_time_seconds) %>%
  group_by(Density) %>% 
   mutate(mean_time = mean(Flip_time_seconds),
         se = sd(Flip_time_seconds)/sqrt(length((Flip_time_seconds))))

#visually check differences in mean
ggplot(Density_flip, aes(x = Density, y = mean_time, group = 1)) +
  geom_line() +
  geom_point() +
  geom_errorbar(width = 0.1, colour = "blue", 
                aes(ymin = mean_time - se, ymax = mean_time + se)) +
  labs(x = "Density", y = "Flip Time (seconds)") +
  theme_classic2()

#descriptive stats
by(Density_flip$Flip_time_seconds, Density_flip$Density, stat.desc)

#checking variances
leveneTest(Density_flip$Flip_time_seconds, Density_flip$Density, center = median)
##appropriate: p = 0.4529, fulfills assumptions

#checking normality (again)
ggplot(Density_flip, aes(sample = Flip_time_seconds)) +
     geom_qq() +
     geom_qq_line() +
     facet_wrap(~Density, scales = "free_y") 
##deviation within groups, check normality by group

#shapiro test, by group
test <- Density_flip %>% dplyr::select(Density, Flip_time_seconds) %>%
  filter(Density == "1larva/ml")
shapiro.test(test$Flip_time_seconds) #1 larva/ml p = <0.01, NOT NORMAL

test <- Density_flip %>% dplyr::select(Density, Flip_time_seconds) %>%
  filter(Density == "2larva/ml")
shapiro.test(test$Flip_time_seconds) #2 larva/ml p = <0.01, NOT NORMAL

test <- Density_flip %>% dplyr::select(Density, Flip_time_seconds) %>%
  filter(Density == "5then2larva/ml")
shapiro.test(test$Flip_time_seconds) #5 then 2 larva/ml p = <0.01, NOT NORMAL

test <- Density_flip %>% dplyr::select(Density, Flip_time_seconds) %>%
  filter(Density == "Bucket")
shapiro.test(test$Flip_time_seconds) #bucket p = <0.01, NOT NORMAL
##all groups not normal, proceed with KW

par(mfrow = c(1, 1)) #reset plot params

#kruskal-wallis 
Density_flip_kruskal <- kruskal.test(Flip_time_seconds ~ Densities, data = Density_flip)
Density_flip_kruskal #p = 0.2961

#post-hoc tests
Density_flip_posthoc <- dunn_test(data = fitness, Flip_time_seconds ~ Densities, p.adjust.method = "bonferroni")
Density_flip_posthoc
Density_flip_posthoc <- Density_flip_posthoc %>% arrange(p.adj)
Density_flip_posthoc <- Density_flip_posthoc %>% add_xy_position(x = "Density")

ggboxplot(Density_flip, x = "Densities", y = "Flip_time_seconds") + 
  geom_jitter(aes(colour = Densities, shape = Densities), size = 2) +
  stat_boxplot(geom = 'errorbar', width = 0.2) +
  stat_pvalue_manual(Density_flip_posthoc, label = "p.adj.signif", hide.ns = TRUE) +
  stat_compare_means(label.y = 2005) +
  labs(x = "Larval Density", y = "Flip Time (seconds)") +
  scale_shape_manual(values = c(15, 16, 17, 18)) +
  scale_color_manual(values = c("#92fe6a", "#68ebff", "#936bff", "#d4646c")) +
  theme_bw() +
  theme(legend.position = "none", axis.title.x = element_blank())
```
A Kruskal-Wallis non-parametric test was performed to compare the effect of pre-settlement larval density on star flip time--performed during the February fitness test. The Kruskal-Wallis test revealed that there WAS NOT a statistically significant difference in flip time between at least two groups (KW Chi2(3) = 3.6968, p = 0.2961)). A non-parametric Dunn's Test for multiple comparisons found that the mean value of flip time was not significantly different between or within any of the diet groups. 


```{r pct GROWTH and DIET}
##Is there a difference in growth rates between settlement diet (dec to feb)

Growth_diet <- growth %>% 
  dplyr::select(Settlement_diet, 
                Length_mm_1, Length_mm_2, 
                Measurement_date_1, Measurement_date_2) %>% 
  drop_na(Settlement_diet, Length_mm_1, Length_mm_2) %>% 
  group_by(Settlement_diet) %>% 
  mutate(Pct_growth = ((Length_mm_2 - Length_mm_1) / Length_mm_1)*100,
         mean_growth = mean(Pct_growth),
         se = sd(Pct_growth)/sqrt(length((Pct_growth)))) %>% #pct growth variable
  filter(Pct_growth >= 0) #exclude negative growth values, now 83 obvs

#visually check differences in mean
ggplot(Growth_diet, aes(x = Settlement_diet, y = mean_growth, group = 1)) +
  geom_line() +
  geom_point() +
  geom_errorbar(width = 0.1, colour = "blue", 
                aes(ymin = mean_growth - se, ymax = mean_growth + se)) +
  labs(x = "Settlement Diet", y = "Mean Percent Growth (January to February)") +
  theme_classic2()

#descriptive stats
by(Growth_diet$Pct_growth, Growth_diet$Settlement_diet, stat.desc)

#checking variances
leveneTest(Growth_diet$Pct_growth, Growth_diet$Settlement_diet, center = median)
##appropriate: p = 0.9389, fulfills assumptions

#checking normality (again)
ggplot(Growth_diet, aes(sample = Pct_growth)) +
     geom_qq() +
     geom_qq_line() +
     facet_wrap(~Settlement_diet, scales = "free_y")
##deviation within all groups

#shapiro.test by group
test <- Growth_diet %>% dplyr::select(Settlement_diet, Pct_growth) %>%
  filter(Settlement_diet == "Kitchen sink")
shapiro.test(test$Pct_growth) #kitchen sink, p = <0.01, NON NORMAL

test <- Growth_diet %>% dplyr::select(Settlement_diet, Pct_growth) %>%
  filter(Settlement_diet == "MLML Urchins")
shapiro.test(test$Pct_growth) #MLML urchins, p = <0.01, NON NORMAL

test <- Growth_diet %>% dplyr::select(Settlement_diet, Pct_growth) %>%
  filter(Settlement_diet == "Worms & Diatoms")
shapiro.test(test$Pct_growth) #worms and diatoms, p = 0.01487
##ALL VARBS non-normal, proceeding with kruskal-wallis 

par(mfrow = c(1, 1)) #reset plot params

#kruskal-wallis 
Growth_diet_kruskal <- kruskal.test(Pct_growth ~ Settlement_diet, data = Growth_diet)
Growth_diet_kruskal #p = 0.355

growth <- growth %>% drop_na() %>%
  mutate(Pct_growth = ((Length_mm_2 - Length_mm_1) / Length_mm_1)*100)

#post-hoc tests
Growth_diet_posthoc <- dunn_test(data = growth, Pct_growth ~ Settlement_diet, p.adjust.method = "bonferroni")
Growth_diet_posthoc
Growth_diet_posthoc <- Growth_diet_posthoc %>% arrange(p.adj)
Growth_diet_posthoc <- Growth_diet_posthoc %>% add_xy_position(x = "Settlement_diet")

#plotting w result
ggboxplot(Growth_diet, x = "Settlement_diet", y = "Pct_growth",
          color = "Settlement_diet", add = "jitter", shape = "Settlement_diet") + 
  stat_pvalue_manual(Growth_diet_posthoc, label = "p.adj.signif",hide.ns = TRUE) +
  stat_compare_means(label.y = 200) + #add global p-value from KW
  labs(x = "Settlement Diet", y = "Percent Growth (January to February)")
```
A Kruskal-Wallis non-parametric test was performed to compare the effect of settlement diet on percent growth from January to February. The Kruskal-Wallis test revealed that there WAS NOT a statistically significant difference in the percent growth between at least two groups (KW Chi2(3) = 3.25, p = 0.355)). A non-parametric Dunn's Test for multiple comparisons found that the mean value of flip time was not significantly different between or within any of the diet groups. 


```{r pct GROWTH and DENSITY}
##Is there a difference in growth rates between larval density treatments (dec to feb)?

Growth_density <- growth %>% 
  dplyr::select(Density, 
                Length_mm_1, Length_mm_2, 
                Measurement_date_1, Measurement_date_2) %>% 
  drop_na(Density, Length_mm_1, Length_mm_2) %>% 
  group_by(Density) %>% 
  mutate(Pct_growth = ((Length_mm_2 - Length_mm_1) / Length_mm_1)*100,
         mean_growth = mean(Pct_growth),
         se = sd(Pct_growth)/sqrt(length((Pct_growth)))) %>%
  filter(Pct_growth >= 0) 

#visually check differences in mean
ggplot(Growth_density, aes(x = Density, y = mean_growth, group = 1)) +
  geom_line() +
  geom_point() +
  geom_errorbar(width = 0.1, colour = "blue", 
                aes(ymin = mean_growth - se, ymax = mean_growth + se)) +
  labs(x = "Density", y = "Mean Percent Growth (January to February)") +
  theme_classic2()

#descriptive stats
by(Growth_density$Pct_growth, Growth_density$Density, stat.desc)

#checking variances
leveneTest(Growth_density$Pct_growth, Growth_density$Density, center = median)
##appropriate: p = 0.5826, fulfills assumptions

#checking normality (again)
ggplot(Growth_density, aes(sample = Pct_growth)) +
     geom_qq() +
     geom_qq_line() +
     facet_wrap(~Density, scales = "free_y")
##deviation in all groups, check by group

#normality between groups
shapiro.test(Growth_density$Pct_growth)

#shapiro test, by group
test <- Growth_density %>% dplyr::select(Density, Pct_growth) %>%
  filter(Density == "1larva/ml")
shapiro.test(test$Pct_growth) #1 larva/ml p = <0.01, NOT NORMAL

test <- Growth_density %>% dplyr::select(Density, Pct_growth) %>%
  filter(Density == "2larva/ml")
shapiro.test(test$Pct_growth) #2 larva/ml p = 0.6915, normal

test <- Growth_density %>% dplyr::select(Density, Pct_growth) %>%
  filter(Density == "5then2larva/ml")
#shapiro.test(test$Pct_growth) ,sample to small to calculate? 

test <- Growth_density %>% dplyr::select(Density, Pct_growth) %>%
  filter(Density == "Bucket")
#shapiro.test(test$Pct_growth) #error message, assuming non-normal
##all groups not normal, proceed with KW

par(mfrow = c(1, 1)) #reset plot params

#kruskal-wallis 
Growth_density_kruskal <- kruskal.test(Pct_growth ~ Density, data = Growth_density)
Growth_density_kruskal #p = 0.2961

#ungroup for analysis
Growth_density <- Growth_density %>% ungroup()

#post-hoc tests
Growth_density_posthoc <- dunn_test(data = Growth_density, Pct_growth ~ Density, p.adjust.method = "bonferroni")
Growth_density_posthoc
Growth_density_posthoc <- Growth_density_posthoc %>% arrange(p.adj)
Growth_density_posthoc <- Growth_density_posthoc %>% add_xy_position(x = "Density")

#plot
ggboxplot(Growth_density, x = "Density", y = "Pct_growth",
          color = "Density", add = "jitter", shape = "Density") + 
  stat_pvalue_manual(Growth_density_posthoc, label = "p.adj.signif", hide.ns = TRUE) +
  stat_compare_means(label.y = 200) + #add global p-value from KW
  labs(x = "Larval Density", y = "Percent Growth (January to February)")
```
A Kruskal-Wallis non-parametric test was performed to compare the effect of pre-settlement larval density on percent growth between the January and Febrauary measurments. The Kruskal-Wallis test revealed that there WAS NOT a statistically significant difference in flip time between at least two groups (KW Chi2(3) = 4.166, p = 0.2441)). A non-parametric Dunn's Test for multiple comparisons found that the mean value of flip time was not significantly different between or within any of the diet groups. 


```{r DESNITY and LENGTH}
##Did the pre-settlement density impact the growth of settled stars?

fitness$Densities <- factor(
  fitness$Density, 
  levels = c("1larva/ml", "2larva/ml", "5then2larva/ml", "Bucket"),
  labels = c("1 larvae/mL", "2 larvae/mL", "5 larvae/mL", "15-20 larvae/mL"))

#LENGTH df 
Density_length <- fitness %>% 
  drop_na(Densities, Flip_time, Flip_time_seconds) %>%
  group_by(Density) %>% #add descriptive stats by group, w/out removing other data points
   mutate(mean_length = mean(Length_mm),
         se = sd(Length_mm)/sqrt(length((Length_mm)))) 

#visually check differences in mean
ggplot(Density_length, aes(x = Density, y = mean_length, group = 1)) +
  geom_line() +
  geom_point() +
  geom_errorbar(width = 0.1, colour = "blue", 
                aes(ymin = mean_length - se, ymax = mean_length + se)) +
  labs(x = "Larval Density", y = "Mean Length") +
  theme_classic2()

#descriptive stats
by(Density_length$Length_mm, Density_length$Density, stat.desc)

#checking variances
leveneTest(Density_length$Length_mm, Density_length$Density, center = median)
##appropriate! p = 0.82, fulfills assumptions

#checking normality
ggplot(Density_length, aes(sample = Length_mm)) +
     geom_qq() +
     geom_qq_line() +
     facet_wrap(~Density, scales = "free_y")
shapiro.test(Density_length$Length_mm)

#shapiro.test by group
test <- Density_length %>% dplyr::select(Density, Length_mm) %>%
  filter(Density == "1larva/ml")
shapiro.test(test$Length_mm) #1 larva/ml p = 0.997

test <- Density_length %>% dplyr::select(Density, Length_mm) %>%
  filter(Density == "2larva/ml")
shapiro.test(test$Length_mm) #2 larva/ml p = 0.629

test <- Density_length %>% dplyr::select(Density, Length_mm) %>%
  filter(Density == "5then2larva/ml")
shapiro.test(test$Length_mm) #5 then 2 larva/ml p = 0.3097

test <- Density_length %>% dplyr::select(Density, Length_mm) %>%
  filter(Density == "Bucket")
shapiro.test(test$Length_mm) #bucket p = 0.4559
##all approximately normal, assumptions fufilled, proceed with regular ANOVA! 

par(mfrow = c(1, 1))

#anova 
Density_length_aov <- aov(Length_mm ~ Densities, data = Density_length)
summary(Density_length_aov) #p = 0.000138***
plot(Density_length_aov)

#robust anova
t1way(Length_mm ~ Density, data = Density_length) #default 20%
##p = 0.00122
t1way(Length_mm ~ Density, data = Density_length, tr = .1) #trim only 10%
##p = 0.00079
med1way(Length_mm ~ Density, data = Density_length)
##p = 0.001

#post-hoc tests
Density_length_posthoc <- Density_length_aov %>% tukey_hsd() #tukey's test
Density_length_posthoc #view results

#work flow for publication-ready plot
Density_length_posthoc <- Density_length_posthoc %>% arrange(p.adj) #arrange
Density_length_posthoc <- add_xy_position(test = Density_length_posthoc,
  data = Density_length %>% as_tibble(.name_repair = "unique"),  #unique column names
  formula = Length_mm ~ Density, fun = "mean", x = "Densities")

#factor levels in the order used in the plot
x_levels <- levels(factor(Density_length$Densities))

#vector to convert group names to x-axis positions
x_map <- setNames(seq_along(x_levels), x_levels)

#map to assign numeric values to xmin and xmax
Density_length_posthoc <- Density_length_posthoc %>%
  mutate(xmin = x_map[group1], xmax = x_map[group2])

#adjusting y.position to fit the signf brackets
Density_length_posthoc <- Density_length_posthoc %>% mutate(y.position = y.position + 40)
Density_length_posthoc$y.position[2] <- 95 #specific y position of 2nd signf bracket

#this plot took so long to trouble shoot.... genuinely... so long... its a MIRACLE it works
ggboxplot(Density_length, x = "Densities", y = "Length_mm", color = "black") + 
  geom_jitter(aes(colour = Densities, shape = Densities), size = 2) +
  stat_boxplot(geom='errorbar', width = 0.2) +
  stat_pvalue_manual(Density_length_posthoc, label = "p.adj.signif", hide.ns = TRUE) +
  stat_compare_means(method = "anova", label.y = 101, label.x = "1 larvae/mL") +
  labs(x = "Larval density", y = "Length (mm)") +
  coord_cartesian(ylim = c(0, 105)) +  
  scale_shape_manual(values = c(15, 16, 17, 18)) +
  scale_color_brewer(type = "qual", palette = 6) +
  theme_bw() +
  theme(legend.position = "none", axis.title.x = element_blank())

report(Density_length_aov)
```
A one-way ANOVA was performed to compare the effect of Density on Length (mm).A one-way ANOVA revealed that there WAS a statistically significant difference in Length between at least two groups (F(3, 83) = 7.67, p = <0.001)). Tukey’s HSD Test for multiple comparisons found that the mean value of Length was significantly different between Bucket and 5 then 2 larva/mL (p = 0.002, 95% C.I. = -21.5, -3.5) and significantly different between Bucket and 1 larva/mL (p = <0.001, 95% C.I. = -23.4, -6.4).There was no statistically significant difference between 2 larva/mL and 1 larva/mL/Bucket (p = 0.5 / 0.2) and no statistically significant difference between 5 then 2 larva/mL and 1 larva/mL/2 larva (p = 0.8 / 0.9).

```{r DENSITY and LENGTH POSTER PLOT}
#LENGTH df 
Density_length <- fitness %>% 
  drop_na(Densities, Flip_time, Flip_time_seconds) %>%
  group_by(Density) %>% #add descriptive stats by group, w/out removing other data points
   mutate(mean_length = mean(Length_mm),
         se = sd(Length_mm)/sqrt(length((Length_mm))))

density_length_poster <- ggboxplot(Density_length, x = "Densities", y = "Length_mm", color = "black") + 
  geom_jitter(aes(colour = Densities, shape = Densities), size = 2) +
  stat_boxplot(geom='errorbar', width = 0.2) +
  stat_pvalue_manual(Density_length_posthoc, label = "p.adj.signif", hide.ns = TRUE) +
  stat_compare_means(method = "anova", label.y = 101, label.x = "1 larvae/mL") +
  labs(x = "Larval density", y = "Length (mm)") +
  coord_cartesian(ylim = c(0, 105)) +  
  scale_shape_manual(values = c(15, 16, 17, 18)) +
  scale_color_manual(values = c("#006b3c", "#D2BAAD", "#9F8FAC", "#f3be77")) +
  theme(legend.position = "none", axis.title.x = element_blank())


density_length_poster +
  theme(
    panel.background = element_rect(fill='transparent'), #transparent panel bg
    plot.background = element_rect(fill='transparent', color=NA), #transparent plot bg
    panel.grid.major = element_blank(), #remove major gridlines
    panel.grid.minor = element_blank(), #remove minor gridlines
    legend.background = element_rect(fill='transparent'), #transparent legend bg
    legend.box.background = element_rect(fill='transparent') #transparent legend panel
  )

ggsave('myplot.png', density_length_poster, bg='transparent')
```

```{r}
library(ggpubr)
library(ggplot2)

density_length_poster <- 
  ggboxplot(
    Density_length,
    x = "Densities",
    y = "Length_mm",
    color = "white",     # box outlines in white
    fill = NA
  ) +
  geom_jitter(
    aes(colour = Densities, shape = Densities),
    size = 2
  ) +
  stat_boxplot(
    geom = "errorbar",
    width = 0.2,
    colour = "white"
  ) +
  stat_pvalue_manual(
    Density_length_posthoc,
    label = "p.adj.signif",
    hide.ns = TRUE,
    color = "white"
  ) +
  stat_compare_means(
    method = "anova",
    label.y = 101,
    label.x = "1 larvae/mL",
    color = "white"
  ) +
  labs(
    y = "Length (mm)",
    x = NULL
  ) +
  coord_cartesian(ylim = c(0, 105)) +
  scale_shape_manual(values = c(15, 16, 17, 18)) +
  scale_color_manual(values = c("#006b3c", "#D2BAAD", "#9F8FAC", "#f3be77")) +
  theme(
    # transparent backgrounds
    panel.background = element_rect(fill = "transparent", colour = NA),
    plot.background  = element_rect(fill = "transparent", colour = NA),
    legend.background = element_rect(fill = "transparent"),
    legend.box.background = element_rect(fill = "transparent"),

    # remove gridlines
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),

    # white text everywhere
    text = element_text(color = "white"),
    axis.text = element_text(color = "white"),
    axis.title = element_text(color = "white"),

    # white axes and ticks
    axis.line = element_line(color = "white"),
    axis.ticks = element_line(color = "white"),

    legend.position = "none"
  )

ggsave(
  filename = "density_length_transparent.png",
  plot = density_length_poster,
  bg = "transparent",
  dpi = 300
)

```


```{r DENSITY and WEIGHT}
##Did the pre-settlement density impact the growth of settled stars?

#WEIGHT df 
Density_weight <- fitness %>% 
  drop_na(Densities, Weight_g) %>%
  group_by(Densities) %>% 
   mutate(mean_weight = mean(Weight_g),
         se = sd(Weight_g)/sqrt(length((Weight_g)))) 

#visually check differences in mean
ggplot(Density_weight, aes(x = Density, y = mean_weight, group = 1)) +
  geom_line() +
  geom_point() +
  geom_errorbar(width = 0.1, colour = "blue", 
                aes(ymin = mean_weight - se, ymax = mean_weight + se)) +
  labs(x = "Larval Density", y = "Mean Weight") +
  theme_classic2()

#descriptive stats
by(Density_weight$Weight_g, Density_weight$Density, stat.desc)

#checking variances
leveneTest(Density_weight$Weight_g, Density_weight$Density, center = median)
##appropriate! p = 0.35, fulfills assumptions

#checking normality
ggplot(Density_weight, aes(sample = Weight_g)) +
     geom_qq() +
     geom_qq_line() +
     facet_wrap(~Density, scales = "free_y") +
  labs(x = "Theoretical Quantiles", y = "Sample Quantiles") +
  theme_few()
#deviation in 3/4 of groups, proceed with shapiro to confirm normality 

#shapiro.test by group
shapiro.test(Density_weight$Weight_g) #p = 3.469e-5, NOT NORMAL
test <- Density_weight %>% dplyr::select(Density, Weight_g) %>%
  filter(Density == "1larva/ml")

shapiro.test(test$Weight_g) #1 larva/ml p = 0.00172, NOT NORMAL
test <- Density_weight %>% dplyr::select(Density, Weight_g) %>%
  filter(Density == "2larva/ml")

shapiro.test(test$Weight_g) #2 larva/ml p = 0.3793, NORMAL
test <- Density_weight %>% dplyr::select(Density, Weight_g) %>%
  filter(Density == "5then2larva/ml")

shapiro.test(test$Weight_g) #5 then 2 larva/ml p = 0.0269, NOT NORMAL
test <- Density_weight %>% dplyr::select(Density, Weight_g) %>%
  filter(Density == "Bucket")

shapiro.test(test$Weight_g) #bucket p = 0.01868, NOT NORMAL
##3/4 not normal, assumptions of normality not fulfilled, proceed with kruskal-wallis

par(mfrow = c(1, 1)) #reset plot params

#kruskal-wallis 
Density_weight_kruskal <- kruskal.test(Weight_g ~ Densities, data = Density_weight)
Density_weight_kruskal #p = 0.001987

#post-hoc tests
Density_weight_posthoc <- dunn_test(data = fitness, Weight_g ~ Densities, p.adjust.method = "bonferroni")
Density_weight_posthoc
Density_weight_posthoc <- Density_weight_posthoc %>% arrange(p.adj)
Density_weight_posthoc <- Density_weight_posthoc %>% add_xy_position(x = "Densities")

#plot
ggboxplot(Density_weight, x = "Densities", y = "Weight_g", color = "black") + 
  geom_jitter(aes(colour = Densities, shape = Densities), size = 2) +
  stat_boxplot(geom = 'errorbar', width = 0.2) +
  stat_pvalue_manual(Density_weight_posthoc, label = "p.adj.signif",hide.ns = TRUE) +
  stat_compare_means(label.y = 16) +
  labs(x = "Larval Density", y = "Weight (g)") +
  scale_shape_manual(values = c(15, 16, 17, 18)) +
  scale_color_manual(values = c("#92fe6a", "#68ebff", "#936bff", "#d4646c")) +
  theme_clean() +
  theme(legend.position = "none", axis.title.x = element_blank())

#length plot for comparison
ggboxplot(Density_length, x = "Densities", y = "Length_mm", color = "black") + 
  geom_jitter(aes(colour = Densities, shape = Densities), size = 2) +
  stat_boxplot(geom='errorbar', width = 0.2) +
  stat_pvalue_manual(Density_length_posthoc, label = "p.adj.signif", hide.ns = TRUE) +
  stat_compare_means(method = "anova", label.y = 101, label.x = "1 larvae/mL") +
  labs(x = "Larval density", y = "Length (mm)") +
  coord_cartesian(ylim = c(0, 105)) +  
  scale_shape_manual(values = c(15, 16, 17, 18)) +
  scale_color_manual(values = c("#92fe6a", "#68ebff", "#936bff", "#d4646c")) +
  theme_clean() +
  theme(legend.position = "none", axis.title.x = element_blank())
```
A Kruskal-Wallis non-parametric test was performed to compare the effect of larval density on star weight (g) post-settlement. The Kruskal-Wallis test revealed that there WAS a statistically significant difference in weight between at least two groups (KW Chi2(3) = 14.81, p = 0.002)). A non-parametric Dunn's Test for multiple comparisons found that the mean value of weight was significantly different between Bucket and 5 then 2 larva/mL (p = 0.002) and significantly different between Bucket and 1 larva/mL (p = 0.0004).There was no statistically significant difference between 2 larva/mL and 1 larva/mL and Bucket (p = 0.16 / 0.02) and no statistically significant difference between 5 then 2 larva/mL and 1 larva/mL and 2 larva (p = 0.8 / 0.2).



```{r DENSITY and ARM COUNT}
##Did the larval density impact the number of arms a star has?

Density_arm <- fitness %>% 
  drop_na(Density, Arm_count) %>%
  group_by(Density) %>% 
   mutate(mean_arm = mean(Arm_count),
         se = sd(Arm_count)/sqrt(length((Arm_count)))) 

#visually check differences in mean
ggplot(Density_arm, aes(x = Density, y = mean_arm, group = 1)) +
  geom_line() +
  geom_point() +
  geom_errorbar(width = 0.1, colour = "blue", 
                aes(ymin = mean_arm - se, ymax = mean_arm + se)) +
  labs(x = "Density", y = "Mean number of arms") +
  theme_classic2()

#descriptive stats
by(Density_arm$Arm_count, Density_arm$Density, stat.desc)

#checking variances
leveneTest(Density_arm$Arm_count, Density_arm$Density, center = median)
##appropriate! p = 0.63

#checking normality
ggplot(Density_arm, aes(sample = Arm_count)) +
     geom_qq() +
     geom_qq_line() +
     facet_wrap(~Density, scales = "free_y") +
  labs(x = "Theoretical Quantiles", y = "Sample Quantiles") +
  theme_few()
#data is largely bi-modal in distribution

#shapiro.test between and within groups
shapiro.test(Density_arm$Arm_count) #overall weight, p = 3.469e-05

test <- Density_arm %>% dplyr::select(Density, Arm_count) %>%
  filter(Density == "1larva/ml")
shapiro.test(test$Arm_count) #1 larva/ml p = 0.002856, NOT NORMAL

test <- Density_arm %>% dplyr::select(Density, Arm_count) %>%
  filter(Density == "2larva/ml")
shapiro.test(test$Arm_count) #2 larva/ml p = 0.151, NORMAL

test <- Density_arm %>% dplyr::select(Density, Arm_count) %>%
  filter(Density == "5then2larva/ml")
shapiro.test(test$Arm_count) #5 then 2 larva/ml p = 0.0003, NOT NORMAL

test <- Density_arm %>% dplyr::select(Density, Arm_count) %>%
  filter(Density == "Bucket")
shapiro.test(test$Arm_count) #bucket p = 0.1941, NORMAL
##2/4 not normal, assumptions of normality not fully fulfilled, proceed with kruskal-wallis

par(mfrow = c(1, 1)) #reset plot params

#kruskal-wallis 
Density_arms_kruskal <- kruskal.test(Arm_count ~ Density, data = Density_arm)
Density_arms_kruskal #p = 0.1229

#anova, to be robust given normality of 1/2 of groups 
Density_arms_aov <- aov(Arm_count ~ Density, data = Density_arm)
summary(Density_arms_aov) #p = 0.203
plot(Density_arms_aov)

#robust anova
t1way(Arm_count ~ Density, data = Density_arm) #default 20%
##p = 0.16899
t1way(Arm_count ~ Density, data = Density_arm, tr = .1) #trim only 10%
##p = 0.17096
med1way(Arm_count ~ Density, data = Density_arm)
##p = 0.143

#post-hoc tests
Density_arm_posthoc <- dunn_test(data = fitness, Arm_count ~ Density, p.adjust.method = "bonferroni")
Density_arm_posthoc
Density_arm_posthoc <- Density_arm_posthoc %>% arrange(p.adj)
Density_arm_posthoc <- Density_arm_posthoc %>% add_xy_position(x = "Density")

#plot
ggboxplot(drop_na(fitness), x = "Density", y = "Arm_count",
          color = "black", add = "jitter", shape = "Density") + 
  stat_pvalue_manual(Density_arm_posthoc, label = "p.adj.signif", hide.ns = TRUE) +
  stat_compare_means(label.y = 15) + #add global p-value
  labs(x = "Larval density", y = "Arm count") +
  theme(legend.position = "none")
```
A Kruskal-Wallis non-parametric test was performed to compare the effect of the various settlement diet treatments on star weight (g) post-settlement. The Kruskal-Wallis test and anova revealed that there was not a statistically significant difference in weight between at least two groups (KW Chi2(3) = 3.462, p = 0.33)). A non-parametric Dunn's Test for multiple comparisons found that the mean value of weight was no statistically significant difference between any particular settlement diet groups.



```{r DIET and WEIGHT}
##Did the post-settlement diet impact the growth of settled stars?

#WEIGHT df, same as previously used diet/weight df, overriding previous work to the df
Diet_weight <- fitness %>% 
  drop_na(Settlement_diet, Weight_g) %>%
  group_by(Settlement_diet) %>% 
   mutate(mean_weight = mean(Weight_g),
         se = sd(Weight_g)/sqrt(length((Weight_g)))) 

#visually check differences in mean
ggplot(Diet_weight, aes(x = Settlement_diet, y = mean_weight, group = 1)) +
  geom_line() +
  geom_point() +
  geom_errorbar(width = 0.1, colour = "blue", 
                aes(ymin = mean_weight - se, ymax = mean_weight + se)) +
  labs(x = "Settlement Diet", y = "Mean Weight") +
  theme_classic2()

#descriptive stats
by(Diet_weight$Weight_g, Diet_weight$Settlement_diet, stat.desc)

#checking variances
leveneTest(Diet_weight$Weight_g, Diet_weight$Settlement_diet, center = median)
##appropriate! p = 0.35, fulfills assumptions

#checking normality
ggplot(Diet_weight, aes(sample = Weight_g)) +
     geom_qq() +
     geom_qq_line() +
     facet_wrap(~Settlement_diet, scales = "free_y") +
  labs(x = "Theoretical Quantiles", y = "Sample Quantiles") +
  theme_few()
#deviation in 1/4 of groups

#shapiro.test between and within groups
shapiro.test(Diet_weight$Weight_g) #overall weight, p = 3.469e-05

test <- Diet_weight %>% dplyr::select(Settlement_diet, Weight_g) %>%
  filter(Settlement_diet == "Juv. stars")
shapiro.test(test$Weight_g) #juv stars p = 0.7887

test <- Diet_weight %>% dplyr::select(Settlement_diet, Weight_g) %>%
  filter(Settlement_diet == "Kitchen sink")
shapiro.test(test$Weight_g) #kitchen sink, p = <0.01, NON NORMAL

test <- Diet_weight %>% dplyr::select(Settlement_diet, Weight_g) %>%
  filter(Settlement_diet == "MLML Urchins")
shapiro.test(test$Weight_g) #MLML urchins, p = 0.4, normal

test <- Diet_weight %>% dplyr::select(Settlement_diet, Weight_g) %>%
  filter(Settlement_diet == "Worms & Diatoms")
shapiro.test(test$Weight_g) #worms and diatoms, p = 0.9872
##1/4 not normal, assumptions of normality not fully fulfilled, proceed with both kruskal-wallis and anova to be robust!

par(mfrow = c(1, 1)) #reset plot params

#kruskal-wallis 
Diet_weight_kruskal <- kruskal.test(Weight_g ~ Settlement_diet, data = Diet_weight)
Diet_weight_kruskal #p = 0.3257

#anova, as a result of only kitchen sink showing deviation from normality 
Diet_weight_aov <- aov(Weight_g ~ Settlement_diet, data = Diet_weight)
summary(Diet_weight_aov) #p = 0.366
plot(Density_length_aov)

#robust anova
t1way(Weight_g ~ Settlement_diet, data = Diet_weight) #default 20%
##p = 0.21566
t1way(Weight_g ~ Settlement_diet, data = Diet_weight, tr = .1) #trim only 10%
##p = 0.23943
med1way(Weight_g ~ Settlement_diet, data = Diet_weight)
##p = 0.457

#post-hoc tests
Diet_weight_posthoc <- dunn_test(data = fitness, Weight_g ~ Settlement_diet, p.adjust.method = "bonferroni")
Diet_weight_posthoc
Diet_weight_posthoc <- Diet_weight_posthoc %>% arrange(p.adj)
Diet_weight_posthoc <- Diet_weight_posthoc %>% add_xy_position(x = "Settlement_diet")

#plotting
ggboxplot(Diet_weight, x = "Settlement_diet", y = "Weight_g",
          color = "Settlement_diet", add = "jitter", shape = "Settlement_diet") + 
  stat_pvalue_manual(Diet_weight_posthoc, label = "p.adj.signif", hide.ns = TRUE) +
  stat_compare_means(label.y = 15) + #add global p-value
  labs(x = "Settlement Diet", y = "Weight (g)")
```
A Kruskal-Wallis non-parametric test was performed to compare the effect of the various settlement diet treatments on star weight (g) post-settlement. The Kruskal-Wallis test and anova revealed that there was not a statistically significant difference in weight between at least two groups (KW Chi2(3) = 3.462, p = 0.33)). A non-parametric Dunn's Test for multiple comparisons found that the mean value of weight was no statistically significant difference between any particular settlement diet groups.



```{r DIET and LENGTH}
##Did the post-settlement diet impact the growth of settled stars?

#LENGTH df, exact same workflow as above
Diet_length <- fitness %>% 
  drop_na(Settlement_diet, Length_mm) %>%
  group_by(Settlement_diet) %>% 
   mutate(mean_length = mean(Length_mm),
         se = sd(Length_mm)/sqrt(length((Length_mm)))) 

Diet_length_no_outliers <- fitness %>% 
  drop_na(Settlement_diet, Length_mm) %>%
  filter(Flip_time_seconds < 500) %>%
  group_by(Settlement_diet) %>% 
  mutate(mean_length = mean(Length_mm),
         se = sd(Length_mm)/sqrt(length((Length_mm)))) 

#visually check differences in mean
ggplot(Diet_length, aes(x = Settlement_diet, y = mean_length, group = 1)) +
  geom_line() +
  geom_point() +
  geom_errorbar(width = 0.1, colour = "blue", 
                aes(ymin = mean_length - se, ymax = mean_length + se)) +
  labs(x = "Settlement Diet", y = "Mean Length (mm)") +
  theme_classic2()

#descriptive stats
by(Diet_length$Length_mm, Diet_length$Settlement_diet, stat.desc)

#checking variances
leveneTest(Diet_length$Length_mm, Diet_length$Settlement_diet, center = median)
##appropriate! p = 0.33, fulfills assumptions

#checking normality
ggplot(Diet_length, aes(sample = Length_mm)) +
     geom_qq() +
     geom_qq_line() +
     facet_wrap(~Settlement_diet, scales = "free_y") +
  labs(x = "Theoretical Quantiles", y = "Sample Quantiles") +
  theme_few()
#deviation in 1/4 of groups (worms and diatoms seems to be bimodal is dist)

#shapiro.test between and within groups
shapiro.test(Diet_length$Length_mm) #overall weight, p = 0.5018

test <- Diet_length %>% dplyr::select(Settlement_diet, Length_mm) %>%
  filter(Settlement_diet == "Juv. stars")
shapiro.test(test$Length_mm) #juv stars p = 0.9154, normal

test <- Diet_length %>% dplyr::select(Settlement_diet, Length_mm) %>%
  filter(Settlement_diet == "Kitchen sink")
shapiro.test(test$Length_mm) #kitchen sink, p = 0.3949, normal

test <- Diet_length %>% dplyr::select(Settlement_diet, Length_mm) %>%
  filter(Settlement_diet == "MLML Urchins")
shapiro.test(test$Length_mm) #MLML urchins, p = 0.9354, normal

test <- Diet_length %>% dplyr::select(Settlement_diet, Length_mm) %>%
  filter(Settlement_diet == "Worms & Diatoms")
shapiro.test(test$Length_mm) #worms and diatoms, p = 0.3751
##all good here, proceeding with regular and robust anova tests

par(mfrow = c(1, 1)) #reset plot params

#anova, as a result of only kitchen sink showing deviation from normality 
Diet_length_aov <- aov(Length_mm ~ Settlement_diet, data = Diet_length)
summary(Diet_length_aov) #p = 0.366
plot(Diet_length_aov)

#robust anova
t1way(Length_mm ~ Settlement_diet, data = Diet_length) #default 20%
##p = 0.07849
t1way(Length_mm ~ Settlement_diet, data = Diet_length, tr = .1) #trim only 10%
##p = 0.0998
med1way(Length_mm ~ Settlement_diet, data = Diet_length)
##p = 0.128

#post-hoc tests
Diet_length_posthoc <- Diet_length_aov %>% tukey_hsd() #tukey's test
Diet_length_posthoc #view results

#work flow for publication-ready plot
Diet_length_posthoc <- Diet_length_posthoc %>% arrange(p.adj) #arrange
Diet_length_posthoc <- add_xy_position(test = Diet_length_posthoc,
  data = Diet_length %>% as_tibble(.name_repair = "unique"),  #unique column names
  formula = Length_mm ~ Density, fun = "mean", x = "Settlement_diet")

#factor levels in the order used in the plot
x_levels <- levels(factor(Diet_length$Settlement_diet))

#vector to convert group names to x-axis positions
x_map <- setNames(seq_along(x_levels), x_levels)

#map to assign numeric values to xmin and xmax
Diet_length_posthoc <- Diet_length_posthoc %>%
  mutate(xmin = x_map[group1], xmax = x_map[group2])

#adjusting y.position to fit the signf brackets
Diet_length_posthoc <- Diet_length_posthoc %>% mutate(y.position = y.position + 40)
Diet_length_posthoc$y.position[2] <- 95 #specific y position of 2nd signf bracket

#now that i know the workflow for this, its a walk in the park
ggboxplot(Diet_length, x = "Settlement_diet", y = "Length_mm", 
          color = "Settlement_diet", add = "jitter", shape = "Settlement_diet") + 
  stat_pvalue_manual(Diet_length_posthoc, label = "p.adj.signif", hide.ns = TRUE) +
  stat_compare_means(method = "anova", label.y = 100, label.x = "Kitchen sink") +
  labs(x = "Settlement Diet", y = "Length (mm)") +
  coord_cartesian(ylim = c(0, 105))  #extend y-axis to fit signf brackets

report(Diet_length_aov)
```
A one-way ANOVA was performed to compare the effect of settlement diet treatment on Length (mm).A one-way ANOVA revealed that there was no statistically significant difference in Length between at least two groups (F(3, 83) = 1.55, p = 0.208)). Tukey’s HSD Test for multiple comparisons found that the mean value of Length was not significantly different between any of the settlement diet groups.




## Linear regression(s)
Asking: how might we be able to apply analysis or statistics to variables in flip time? Could this help us determine a particular trend in linearity between variables? 
Used base r "stats" package to run lm()

```{r LENGTH and FLIP TIME}
##Do longer stars flip over slower? 

Size_flip <- fitness %>% 
  drop_na(Length_mm, Weight_g, Flip_time, Flip_time_seconds)

Size_flip_no_outliers <- fitness %>% 
  drop_na(Length_mm, Weight_g, Flip_time, Flip_time_seconds) %>%
  filter(Flip_time_seconds < 1000) #13/88 observations with flip times above 500 seconds

#test normality visually (qq-plot/hist)
hist(Size_flip$Length_mm) #fairly normal
hist(Size_flip$Flip_time_seconds) #right skew, non-normal
qqnorm(Size_flip$Length_mm) #length is normal
qqnorm(Size_flip$Flip_time_seconds) #flip time is NOT normal

#also check normality numerically with shapiro (just simply for my learning)
shapiro.test(Size_flip$Length_mm) #p=0.44, no deviation from normality
shapiro.test(Size_flip$Flip_time_seconds) #p=1.411e-12, HIGH deviation from normality

#running a linear regression model
Flip_length_lm <- lm(formula = Flip_time_seconds ~ Length_mm, data = Size_flip)
Flip_length_lm
summary(Flip_length_lm)
sqrt(0.03947) #pearson coefficient is = 0.1986
AIC(Flip_length_lm) #AIC = 1277.18

#flip_time ~ length model diagnostics
model.diag.metrics <- augment(Flip_length_lm)
head(model.diag.metrics)

#plotted residuals error (red) between observed values and fitted regression line
ggplot(model.diag.metrics, aes(Length_mm, Flip_time_seconds)) +
  geom_point() +
  stat_smooth(method = lm, se = FALSE) +
  geom_segment(aes(xend = Length_mm, yend = .fitted), color = "red", size = 0.3) +
  labs(x = "Length (mm)", y = "Flip Time (seconds)") +
  theme_classic2()
```
Simple linear regression was used to test if Length (mm) significantly predicted P. helianothoides flip time from the February fitness test. The fitted regression model was: [Flip time = 549 x -5.424*(Length)]. The overall regression was statistically insignificant (R2 = 0.039, F(1, 86) = 3.534, p = 0.0635).It was found that length is not an accurate predictor of flip time (β = -5.242, p = 0.0635).


```{r WEIGHT and FLIP TIME }
##Do small stars maneuver or flip faster?

#using same working df's from above question!

#test normality visually (qq-plot/hist)
hist(Size_flip$Weight_g) #right skew
qqnorm(Size_flip$Weight_g) #non-normal
qqline(Size_flip$Weight_g) #non-normal

#confirm with shapiro (we already know Flip time is not normal, given above)
shapiro.test(Size_flip$Weight_g) #p=0.00007 weight is NOT normal... apparently

#running a linear regression model
Flip_weight_lm <- lm(formula = Flip_time_seconds ~ Weight_g, data = Size_flip)
summary(Flip_weight_lm) #p = 0.1189
sqrt(0.02804) #pearson coefficient is = 0.1674515

#flip_time ~ length model diagnostics
model.diag.metrics <- augment(Flip_weight_lm)
head(model.diag.metrics)

#plotted residuals error (red) between observed values and fitted regression line
ggplot(model.diag.metrics, aes(Weight_g, Flip_time_seconds)) +
  geom_point() +
  stat_smooth(method = lm, se = FALSE) +
  geom_segment(aes(xend = Weight_g, yend = .fitted), color = "red", size = 0.3) +
  labs(x = "Weight (g)", y = "Flip Time (seconds)") +
  theme_classic2()
```
Weight of a sea-star can account for 2.8% (R2 = 0.028) of the variation in sea-star flip time (given outliers are included in the analysis). This model, similar to the first length model, DOES NOT accurately predict the variation in sea-star flip time given the variation in star length (F=2.481, p=0.1189).


```{r WEIGHT and FLIP TIME no outliers}
#linear regression with NON outlier data

#normality tests with NON outlier data (qq-plot/hist)
hist(Size_flip_no_outliers$Weight_g) #still right skew
qqnorm(Size_flip_no_outliers$Weight_g) 
qqline(Size_flip_no_outliers$Weight_g) #better fit, but still non-normal

#shapiro test
shapiro.test(Size_flip_no_outliers$Weight_g) #still non-normal without outliers (p = 0.00027)

#regression with NO OUTLIER data
Flip_weight_lm_2 <- lm(formula = Flip_time_seconds ~ Weight_g, data = Size_flip_no_outliers)
summary(Flip_weight_lm_2) #signf, (p = 0.002)
sqrt(0.05619) #pearson coefficient = 0.237
```
Weight of a sea-star can account for 5.6% (R2 = 0.05619) of the variation in sea-star flip time (given outliers are NOT included in the analysis). This model, similar to the first CAN accurately predict the variation in sea-star flip time given the variation in star length (F = 4.822, p = 0.03).



```{r ARM COUNT and FLIP TIME}
##Does the number of arms a star have correlate to the speed of flip time?

#normality check
hist(Size_flip$Arm_count) #bi-modal dist
qqnorm(Size_flip$Arm_count) #VERY bi-modal
qqline(Size_flip$Arm_count)

#normality test
shapiro.test(Size_flip$Arm_count) #non-normal, p = 1.3e-5

#regression w/ flip time outliers
Flip_arms_1 <- lm(formula = Flip_time_seconds ~ Arm_count, data = Size_flip)
summary(Flip_arms_1)
sqrt(0.04663) #pearson coef = 0.2159
```
Results...


```{r ARM COUNT and FLIP TIME no outliers}
#regression with NON outlier data

#normality check
hist(Size_flip_no_outliers$Arm_count) 
qqnorm(Size_flip_no_outliers$Arm_count) #still very bi-modal
qqline(Size_flip_no_outliers$Arm_count)

#normality test
shapiro.test(Size_flip_no_outliers$Arm_count) #non-normal, p = 1.924e-5

#regression w/out flip time outliers
Flip_arms_2 <- lm(formula = Flip_time_seconds ~ Arm_count, data = Size_flip_no_outliers)
summary(Flip_arms_2)

```
Results...




## Multiple regression analyses
```{r checking homo/heteroscedasticity}
#using same working df's as above, testing for homoscedasticity in predictor variables
Multi_model_test <- lm(Flip_time_seconds ~ Length_mm + Weight_g + Arm_count + Eating_score + Wellness_score, data = Size_flip) #choosing these 5 predictor variables for the model, checking assumptions below

ncvTest(Multi_model_test) 
##data is not fulfilling assumptions, showing heteroscedasticity (p= 5.12e-8)

#testing predictors individually
ncvTest(Multi_model_test, ~ Length_mm) #p=0.16, good
ncvTest(Multi_model_test, ~ Weight_g) #p=0.62, good
ncvTest(Multi_model_test, ~ Arm_count) #p= 0.0016, NOT GOOD
ncvTest(Multi_model_test, ~ Eating_score) #p=7.51e-6, NOT GOOD 
ncvTest(Multi_model_test, ~ Wellness_score) #p=0.08, good
ncvTest(Multi_model_test, ~ Length_mm + Weight_g) #p=0.11, good
ncvTest(Multi_model_test, ~ Length_mm + Weight_g + Arm_count) #p=0.004, NOT GOOD
ncvTest(Multi_model_test, ~ Length_mm + Weight_g + Wellness_score) #p=0.04, NOT GOOD
ncvTest(Multi_model_test, ~ Length_mm + Weight_g + Eating_score) #p=1.49e-5, good
ncvTest(Multi_model_test, ~ Length_mm + Weight_g + Eating_score + Arm_count) #p=1.78e-6, NOT GOOD
ncvTest(Multi_model_test, ~ Length_mm + Weight_g + Eating_score + Wellness_score) #p=bad
ncvTest(Multi_model_test, ~ Length_mm + Wellness_score) #p=0.06, good
ncvTest(Multi_model_test, ~ Weight_g + Wellness_score) #p=0.18, better than length
ncvTest(Multi_model_test, ~ Weight_g + Wellness_score + Eating_score) #p=bad
```
Arm_count AND Eating score are heteroscedastic, therefore will not be included in the model, however wellness score is heteroscedastic when with weight/length



```{r multiple regression model with FLIP TIME by WEIGHT and WELLNESS SCORE}
#model without Arm_count, Eating_score, and without Length_mm (given result of heteroscedasticity pairing above)
Multi_model_1 <- lm(Flip_time_seconds ~ Weight_g + Wellness_score, data = Size_flip)
summary(Multi_model_1)
```
This model can explain 3.5% of the variation in Flip time (R2 = 0.035), as opposed to 3.9% in just the model fit with Length.
    Non significant model results (F = 1.543, p = 0.2196)
    No particular parameter is making a significant contribution to the model given Pr(>|t|) all are >0.05


```{r model diagnostics}
#reporting standardized ß measurements
#lm.beta(Multi_model_1) #code no longer working, ignore for now
##Given standardized units, as weight increases or changes, flip time will decrease and as Wellness_sore increases, flip time will also increase by units according to the standardized ß scores (Weight_gß = -0.17, Wellness_scoreß = 0.08). 

#confidence intervals of regression model
confint(Multi_model_1)

#Residuals for multi-model 1 (model with outliers)
Size_flip$Residuals <- resid(Multi_model_1)
Size_flip$Standard_residuals <- rstandard(Multi_model_1)
Size_flip$Studentized_residuals <- rstudent(Multi_model_1)
Size_flip$Cooks_distance <- cooks.distance(Multi_model_1)
Size_flip$Df_beta <- dfbeta(Multi_model_1)
Size_flip$Df_fit <- dffits(Multi_model_1)
Size_flip$Leverage <- hatvalues(Multi_model_1)
Size_flip$Covariance_ratios <- covratio(Multi_model_1)

#large residuals search; every case when residual is less than -2 or greater than 
Size_flip$Large_residuals <- Size_flip$Standard_residuals > 2 | 
  Size_flip$Standard_residuals < -2 
sum(Size_flip$Large_residuals) #6 cases with a large residual

#identifying large residuals
Size_flip[Size_flip$Large_residuals, c("Weight_g", "Wellness_score", "Standard_residuals")] 

#looking at other values where large residuals is TRUE
Size_flip[Size_flip$Large_residuals, c("Cooks_distance", "Leverage", "Covariance_ratios")] 
```
    none of the cooks distance numbers are greater than one, so these cases are not having an undue influence on the model. 
    No cases have a higher average leverage. 
    NONE of the cases here have CVR values within the bounds (1.06-0.94). There is undue influence from all of these cases according to the CVR


```{r model diagnostics}
#Durban Watson test of independence
dwt(Multi_model_1)
##D-W stat = 1.32, semi-close to 2, and a p = 0.008, so we cannot assume independence ???

#plotting studentized residuals, fitted values, and theoretical values of the model
Size_flip$Fitted <- Multi_model_1$fitted.values

#histogram with dnorm line
histogram.residuals <- ggplot(Size_flip, aes(Studentized_residuals)) +
  geom_histogram(aes(y = ..density..), colour = "black", fill = "white") +
  labs(x = "Studentized Residuals", y = "Density") +
  theme_clean()
histogram.residuals <- histogram.residuals + stat_function(fun = dnorm, args = list(
  mean = mean(Size_flip$Studentized_residuals, na.rm = TRUE),
  sd = sd(Size_flip$Studentized_residuals, na.rm = TRUE)), 
  colour = "red", size = 0.5) +
  theme_clean()

#Q-Q plot
qqplot.residuals <- ggplot() +
  geom_qq(data = Size_flip, aes(sample = Studentized_residuals)) + 
  labs(x = "Theoretical Values", y = "Observed Values") +
  theme_clean()

#scatter
scatter.residuals <- ggplot(Size_flip, aes(Fitted, Studentized_residuals)) +
  geom_point() + 
  geom_smooth(method = lm, colour = "blue") +
  labs(x = "Fitted Values", y = "Stundentized Residual") +
  theme_clean()

#patchwork graphs for viewing
histogram.residuals + qqplot.residuals + scatter.residuals
```
model is fairly (kinda?) accurate for our population, but this model cannot be generalized to a greater population in better determining responses of sea-stars to environmental stimuli (ie. being pushed upside down, then righting oneself)


```{r multiple regression model with FLIP TIME by WEIGHT and WELLNESS SCORE no outliers}
#Multi model without flip time outliers 
Multi_model_1_no <- lm(Flip_time_seconds ~ Weight_g + Wellness_score, data = Size_flip_no_outliers)
summary(Multi_model_1_no)
```
This model explains 12.8% in the variation (R2 = 0.165) of Flip time as opposed to 11.2% (R2 = 0.112) variation from just length without flip time outliers simple model. 
There is a significant relationship in this model (p = 0.007)
Adjusted R2 is different by xxx, so cross validity of the model is... okay

```{r model diagnostics}
#standardized beta scores
#lm.beta(Multi_model_1_no) 
##Given standardized units, as weight AND wellness score increase or change, flip time will decrease according to the standardized ß scores (Weight_gß = -0.32, Wellness_scoreß = 0.11). 

#confidence intervals of regression model
confint(Multi_model_1_no)

#Residuals for multi-model 1 (model with outliers)
Size_flip_no_outliers$Residuals <- resid(Multi_model_1_no)
Size_flip_no_outliers$Standard_residuals <- rstandard(Multi_model_1_no)
Size_flip_no_outliers$Studentized_residuals <- rstudent(Multi_model_1_no)
Size_flip_no_outliers$Cooks_distance <- cooks.distance(Multi_model_1_no)
Size_flip_no_outliers$Df_beta <- dfbeta(Multi_model_1_no)
Size_flip_no_outliers$Df_fit <- dffits(Multi_model_1_no)
Size_flip_no_outliers$Leverage <- hatvalues(Multi_model_1_no)
Size_flip_no_outliers$Covariance_ratios <- covratio(Multi_model_1_no)

#large residuals search  
Size_flip_no_outliers$Large_residuals <- Size_flip_no_outliers$Standard_residuals > 2 |
  Size_flip_no_outliers$Standard_residuals < -2  
sum(Size_flip_no_outliers$Large_residuals) #5 cases with a large residual vs. 6 WITH outliers

#identifying large residuals
Size_flip_no_outliers[Size_flip_no_outliers$Large_residuals, 
                      c("Weight_g", "Wellness_score", "Standard_residuals")]
Size_flip_no_outliers[Size_flip_no_outliers$Large_residuals, 
                      c("Cooks_distance", "Leverage", "Covariance_ratios")]
```
    none of the cooks distance numbers are greater than one, so these cases are not having an undue influence on the model. 
    instance 2 has a notable larger cooks distance than the others
    No cases have a higher average leverage. instance 2 has another high leverage score
    NONE of the cases here have CVR values within the bounds (1.06-0.94). 
    There is undue influence from all of these cases according to the CVR, excpet for case 2, which has a CVR = 0.95. 

```{r model diagnostics}
#Durban Watson test of independence
dwt(Multi_model_1)
##D-W stat = 1.32, semi-close to 2, and a p = 0.006, so we cannot assume independence ???

#plotting studentized residuals, fitted values, and theoretical values of the model
Size_flip_no_outliers$Fitted <- Multi_model_1_no$fitted.values

#histogram with dnorm line
histogram.residuals <- ggplot(Size_flip_no_outliers, aes(Studentized_residuals)) +
  geom_histogram(aes(y = ..density..), colour = "black", fill = "white") +
  labs(x = "Studentized Residuals", y = "Density") +
  theme_clean()
histogram.residuals + stat_function(fun = dnorm, args = list(
  mean = mean(Size_flip_no_outliers$Studentized_residuals, na.rm = TRUE),
  sd = sd(Size_flip_no_outliers$Studentized_residuals, na.rm = TRUE)), 
  colour = "red", size = 0.5) +
  theme_clean()

#Q-Q plot
qqplot.residuals <- ggplot() +
  geom_qq(data = Size_flip_no_outliers, aes(sample = Studentized_residuals)) + 
  labs(x = "Theoretical Values", y = "Observed Values") +
  theme_clean()

#scatter
scatter.residuals <- ggplot(Size_flip_no_outliers, aes(Fitted, Studentized_residuals)) +
  geom_point() + 
  geom_smooth(method = lm, colour = "blue") +
  labs(x = "Fitted Values", y = "Stundentized Residual") +
  theme_clean()

#patchwork graphs for viewing
histogram.residuals + qqplot.residuals + scatter.residuals
```
Again, model is fairly (kinda?) accurate for our population, but this model still cannot be generalized to a greater population (even without extreme cases)



```{r MLR with FLIP TIME by LENGTH and WELLNESS SCORE}
Multi_model_2 <- lm(Flip_time_seconds ~ Length_mm + Wellness_score, data = Size_flip)
summary(Multi_model_2)
```
This model can explain 4.7% in the variation (R2 = 0.0477) of Flip time as opposed to 3.5% of the variation in the previous mutli model (flip time ~ weight + wellness score) with outliers. 
There is NO significant relationship in this model (p = 0.1252)
Adjusted R2 is different by 0.02241, so cross validity of the model is okay, but still not strong


```{r model diagnostics}
#standardized beta scores
#lm.beta(Multi_model_2) 
##Given standardized units, length and wellness score have opposite effects on flip time, as length increases flip time decreases (Length_mmß = -0.21) and as wellness score increases (Wellness_scoreß = 0.09), flip time increases? TBH this is a super weird result and I don't truly trust the validity of this model 

#confidence intervals of regression model
confint(Multi_model_2)
##

#Residuals for multi-model 2 (model without outliers)
Size_flip$Residuals <- resid(Multi_model_2)
Size_flip$Standard_residuals <- rstandard(Multi_model_2)
Size_flip$Studentized_residuals <- rstudent(Multi_model_2)
Size_flip$Cooks_distance <- cooks.distance(Multi_model_2)
Size_flip$Df_beta <- dfbeta(Multi_model_2)
Size_flip$Df_fit <- dffits(Multi_model_2)
Size_flip$Leverage <- hatvalues(Multi_model_2)
Size_flip$Covariance_ratios <- covratio(Multi_model_2)

#save new data
write.table(Size_flip, "Multi model 2 with diagnostics.dat", sep = "\t", row.names = FALSE)

#large residuals search  
Size_flip$Large_residuals <- 
  Size_flip$Standard_residuals > 2 |
  Size_flip$Standard_residuals < -2 # 
sum(Size_flip$Large_residuals) #6 cases with a large residual 

#identifying large residuals
Size_flip[Size_flip$Large_residuals, 
                      c("Weight_g", "Wellness_score", "Standard_residuals")]
Size_flip[Size_flip$Large_residuals, 
                      c("Cooks_distance", "Leverage", "Covariance_ratios")]
```
none of the cooks distance numbers are greater than one, so these cases are not having an undue influence on the model. instance 23 has a notably higher cooks distance than the others
No cases have a higher average leverage
NONE of the cases have a CVR value within the bounds (1.06-0.94). 
There is undue influence from 6 of these cases according to the CVR 


```{r model diagnostics}
#Durban Watson test of independence
dwt(Multi_model_2)
##D-W stat = 1.29, semi-close to 2, and a p = 0.004, same value as Multi_model_1

#plotting studentized residuals, fitted values, and theoretical values of the model
Size_flip$Fitted <- Multi_model_2$fitted.values

#histogram with dnorm line
histogram.residuals <- ggplot(Size_flip, aes(Studentized_residuals)) +
  geom_histogram(aes(y = ..density..), colour = "black", fill = "white") +
  labs(x = "Studentized Residuals", y = "Density") +
  theme_clean()
histogram.residuals <- histogram.residuals + stat_function(fun = dnorm, args = list(
  mean = mean(Size_flip$Studentized_residuals, na.rm = TRUE),
  sd = sd(Size_flip$Studentized_residuals, na.rm = TRUE)), 
  colour = "red", size = 0.5) +
  theme_clean()

#Q-Q plot
qqplot.residuals <- ggplot() +
  geom_qq(data = Size_flip, aes(sample = Studentized_residuals)) + 
  labs(x = "Theoretical Values", y = "Observed Values") +
  theme_clean()

#scatter
scatter.residuals <- ggplot(Size_flip, aes(Fitted, Studentized_residuals)) +
  geom_point() + 
  geom_smooth(method = lm, colour = "blue") +
  labs(x = "Fitted Values", y = "Stundentized Residual") +
  theme_clean()

#patchwork graphs for viewing
histogram.residuals + qqplot.residuals + scatter.residuals
```
Same as above results



```{r MRM with FLIP TIME by LENGTH and WELLNESS SCORE no outliers}
Multi_model_2_no <- lm(Flip_time_seconds ~ Length_mm + Wellness_score, data = Size_flip_no_outliers)
summary(Multi_model_2_no)
```
This model can explain 12.5% in the variation (R2 = 0.1254) of Flip time as opposed to the 12.8% of the variation in the previous multi model (flip time ~ weight + wellness score) run without outliers. 
There is NO significant relationship in this model (p = 0.1252)
Adjusted R2 is different by 0.02241, so cross validity of the model is okay, but still not strong


```{r model diagnostics}
#standardized beta scores
#lm.beta(Multi_model_2_no) 
##Given standardized units, length and wellness score have both negative ß scores (Length_mmß = -0.322) and as wellness score increases (Wellness_scoreß = -0.116)

#confidence intervals of regression model
confint(Multi_model_2_no)

#Residuals for multi-model 2 (model without outliers)
Size_flip_no_outliers$Residuals <- resid(Multi_model_2_no)
Size_flip_no_outliers$Standard_residuals <- rstandard(Multi_model_2_no)
Size_flip_no_outliers$Studentized_residuals <- rstudent(Multi_model_2_no)
Size_flip_no_outliers$Cooks_distance <- cooks.distance(Multi_model_2_no)
Size_flip_no_outliers$Df_beta <- dfbeta(Multi_model_2_no)
Size_flip_no_outliers$Df_fit <- dffits(Multi_model_2_no)
Size_flip_no_outliers$Leverage <- hatvalues(Multi_model_2_no)
Size_flip_no_outliers$Covariance_ratios <- covratio(Multi_model_2_no)

#large residuals search  
Size_flip_no_outliers$Large_residuals <- Size_flip_no_outliers$Standard_residuals > 2 |
  Size_flip_no_outliers$Standard_residuals < -2  
sum(Size_flip_no_outliers$Large_residuals) #4 cases with a large residual 

#identifying large residuals
Size_flip_no_outliers[Size_flip_no_outliers$Large_residuals, 
                      c("Weight_g", "Wellness_score", "Standard_residuals")]
Size_flip_no_outliers[Size_flip_no_outliers$Large_residuals, 
                      c("Cooks_distance", "Leverage", "Covariance_ratios")]
```
none of the cooks distance numbers are greater than one, so these cases are not having an undue influence on the model. instance 2 and 42 have notably higher ratios than the other values
No cases have a higher average leverage
NONE of the cases have a CVR value within the bounds (1.06-0.94) but instance 2 and 42 have higher values (approaching 0.95). There is undue influence from 2/4 of these cases according to the CVR 


```{r model diagnostics}
#Durban Watson test of independence
dwt(Multi_model_2_no)
##D-W stat = 1.59, much closer to 2 than any of the other models, and a p = 0.064, so independence is not assumed and likely violated

#still plotting studentized residuals, fitted values, and theoretical values of the model to assess fit of residuals 
Size_flip_no_outliers$Fitted <- Multi_model_2_no$fitted.values

#histogram with dnorm line
histogram.residuals <- ggplot(Size_flip_no_outliers, aes(Studentized_residuals)) +
  geom_histogram(aes(y = ..density..), colour = "black", fill = "white") +
  labs(x = "Studentized Residuals", y = "Density") +
  theme_clean()
histogram.residuals <- histogram.residuals + stat_function(fun = dnorm, args = list(
  mean = mean(Size_flip_no_outliers$Studentized_residuals, na.rm = TRUE),
  sd = sd(Size_flip_no_outliers$Studentized_residuals, na.rm = TRUE)), 
  colour = "red", size = 0.5) +
  theme_clean()

#Q-Q plot
qqplot.residuals <- ggplot() +
  geom_qq(data = Size_flip_no_outliers, aes(sample = Studentized_residuals)) + 
  labs(x = "Theoretical Values", y = "Observed Values") +
  theme_clean()

#scatter
scatter.residuals <- ggplot(Size_flip_no_outliers, aes(Fitted, Studentized_residuals)) +
  geom_point() + 
  geom_smooth(method = lm, colour = "blue") +
  labs(x = "Fitted Values", y = "Stundentized Residual") +
  theme_clean()

#patchwork graphs for viewing
histogram.residuals + qqplot.residuals + scatter.residuals
```


## Plotting -- settlement data
```{r cumulative SETTLEMENT overtime}
#fixing labeling for plots
settlement_raw_long$Densities <- factor(
  settlement_raw_long$Density, 
  levels = c("1larva/ml", "2larva/ml", "5then2larva/ml", "Bucket"),
  labels = c("1 larvae/mL", "2 larvae/mL", "5 larvae/mL", "15-20 larvae/mL"))

#cumulative settlement by larval density, df 
Density_settlement <- settlement_raw_long %>% 
  group_by(Densities, DPF) %>% #initially group by DPF and density
  summarise(Settlers = sum(Settlers)) %>% #sum the settler count by day for each density
  arrange(Densities, DPF) %>% #arrange df by density and DPF
  group_by(Densities) %>% #cumulative sum only by density 
  mutate(Cum_sum = cumsum(Settlers)) #cumulative sum by density group over-time

#line plot!
ggplot(Density_settlement, aes(x = DPF, y = Cum_sum, colour = Densities)) +
  geom_line(size = 0.7) +
  labs(x = "Days post fertilization (DPF)",y = "Approximate larval settlers", 
       colour = "Larval density") +
  scale_color_manual(values = c("#92fe6a", "#68ebff", "#936bff", "#d4646c")) +
  scale_x_continuous(breaks = seq(0, 150, by = 10)) +
  theme_bw()

#cum settlement by density normalized according to # of replicates in density groups
replicate_counts <- c("1 larvae/mL" = 8, #replicates in each density trtment
                      "2 larvae/mL" = 4, 
                      "5 larvae/mL" = 4, 
                      "15-20 larvae/mL" = 4)

#working df
Density_settlement_normalized <- settlement_raw_long %>%
  group_by(Densities, DPF) %>% #summarize by density and DPF
  summarise(Daily_settlers = sum(Settlers), .groups = "drop") %>% #sum by daily settlers
  arrange(Densities, DPF) %>% #rearrange df
  group_by(Densities) %>% #calculate and mutate by density groups
  mutate(Cum_sum = cumsum(Daily_settlers),
    Normalized_cum_sum = Cum_sum / replicate_counts[Densities]) #divide cumsum by # replicates

#plot
ggplot(Density_settlement_normalized, aes(x = DPF, y = Normalized_cum_sum, colour = Densities)) +
  geom_line(size = 0.7) +
  labs(x = "Days post fertilization (DPF)", y = "Normalized cumulative daily settlement",
       colour = "Larval density") +
  scale_color_manual(values = c("#92fe6a", "#30d5c8", "#c0c0ff", "#f07177")) +
  scale_x_continuous(breaks = seq(0, 150, by = 10)) +
  theme_bw()

#Surviving Juveniles in cohort overtime, df
test <- settlement_cohort %>% arrange(match(Date_interval, c("Month 3", "Month 6", "Month 9", "Month 12", "Month 15")))
```

```{r normalized CUMULATIVE settlement POSTER PLOT}
normalized_settlement_poster <- 
  ggplot(Density_settlement_normalized, 
         aes(x = DPF, y = Normalized_cum_sum, colour = Densities)) +
  geom_line(size = 0.7) +
  labs(x = "Days post fertilization (DPF)", 
       y = "Normalized cumulative daily settlement",
       colour = "Larval Density") +
  scale_color_manual(values = c("#006b3c", "#D2BAAD", "#9F8FAC", "#f3be77")) +
  scale_x_continuous(breaks = seq(0, 150, by = 10)) +
  theme(legend.position = 'top',
        panel.background = element_rect(fill='transparent'), #transparent panel bg
        plot.background = element_rect(fill='transparent', color=NA), #transparent plot bg
        panel.grid.major = element_blank(), #remove major gridlines
        panel.grid.minor = element_blank(), #remove minor gridlines
        legend.background = element_rect(fill='transparent'), #transparent legend bg
        legend.box.background = element_rect(fill='transparent')) +
  theme_classic() +
  theme(legend.position = 'top')

ggsave('normalized_settlement.png', normalized_settlement_poster, bg='transparent')
```

```{r}
normalized_settlement_poster <- 
  ggplot(
    Density_settlement_normalized,
    aes(x = DPF, y = Normalized_cum_sum, colour = Densities)
  ) +
  geom_line(size = 0.7) +
  labs(
    x = "Days post fertilization (DPF)",
    y = "Normalized cumulative daily settlement",
    colour = "Larval Density"
  ) +
  scale_color_manual(values = c("#006b3c", "#D2BAAD", "#9F8FAC", "#f3be77")) +
  scale_x_continuous(breaks = seq(0, 150, by = 10)) +
  theme(
    # transparent backgrounds
    panel.background = element_rect(fill = "transparent", colour = NA),
    plot.background  = element_rect(fill = "transparent", colour = NA),
    legend.background = element_rect(fill = "transparent"),
    legend.box.background = element_rect(fill = "transparent"),

    # remove gridlines
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),

    # white text everywhere
    text = element_text(color = "white"),
    axis.text = element_text(color = "white"),
    axis.title = element_text(color = "white"),
    legend.text = element_text(color = "white"),
    legend.title = element_text(color = "white"),

    # axes lines and ticks in white (important on dark backgrounds)
    axis.line = element_line(color = "white"),
    axis.ticks = element_line(color = "white"),

    legend.position = "top"
  )

ggsave(
  filename = "normalized_settlement_transparent.png",
  plot = normalized_settlement_poster,
  bg = "transparent",
  dpi = 300
)
```


```{r survivorship by DENSITY}
survivorship <- settlement_total

#density labels for plot
survivorship$Densities <- factor(
  survivorship$Density, 
  levels = c("1larva/ml", "2larva/ml", "5then2larva/ml", "Bucket"),
  labels = c("1 larvae/mL", "2 larvae/mL", "5 larvae/mL", "15-20 larvae/mL"))

#plot
ggplot(survivorship, aes(x = Densities, y = Survivorship, 
                         shape = Densities)) +
  stat_boxplot(geom = 'errorbar', width = 0.2) +
  geom_boxplot() +
  geom_jitter(aes(colour = Densities)) +
  scale_y_continuous(labels = scales::percent) +
  scale_shape_manual(values = c(15, 16, 17, 18)) +
  scale_color_manual(values = c("#92fe6a", "#30d5c8", "#c0c0ff", "#f07177")) +
  theme_classic2() +
  theme(legend.position = "none")
```





## Statistical analysis -- settlement and survival (settlement data)
```{r DENSITY and SETTLEMENT}
##Did the pre-settlement density impact the number of settled stars?

Density_settlement_normalized_group <- settlement_raw_long %>%
  group_by(Densities, DPF, Group) %>% 
  summarise(Daily_settlers = sum(Settlers), 
            .groups = "drop") %>%
  arrange(Densities, DPF) %>%
  group_by(Group) %>% 
  mutate(Cum_sum = cumsum(Daily_settlers),
    Normalized_cum_sum = Cum_sum / replicate_counts[Densities])

#df for tests
Normalized_sum <- Density_settlement_normalized_group %>% 
  dplyr::filter(DPF >= 146) %>%
  group_by(Densities) %>%
  mutate(se = sd(Normalized_cum_sum)/sqrt(length(Normalized_cum_sum)))

#mean for below graph
Normalized_mean_sum <- Density_settlement_normalized_group %>% 
  dplyr::filter(DPF >= 146) %>%
  group_by(Densities) %>%
  summarise(Mean_settle = mean(Normalized_cum_sum),
            se = sd(Normalized_cum_sum)/sqrt(length(Normalized_cum_sum)))

#visually check differences in mean
ggplot(Normalized_mean_sum, aes(x = Densities, y = Mean_settle, group = 1)) +
  geom_line() +
  geom_point() +
  geom_jitter(data = Normalized_sum, aes(y = Normalized_cum_sum, 
                                         colour = Densities, shape = Densities)) +
  geom_errorbar(width = 0.1, colour = "red", 
                aes(ymin = Mean_settle - se, ymax = Mean_settle + se)) +
  labs(x = "Larval Density", y = "Mean normalized cumulative settlement") +
  scale_shape_manual(values = c(15, 16, 17, 18)) +
  scale_color_manual(values = c("#92fe6a", "#30d5c8", "#c0c0ff", "#f07177"))  +
  theme_bw() +
  theme(axis.title.x = element_blank())

by(Normalized_sum$Normalized_cum_sum, Normalized_sum$Densities, stat.desc)

leveneTest(Normalized_sum$Normalized_cum_sum, Normalized_sum$Densities, center = median) #appropriate, p = 0.88

ggplot(Normalized_sum, aes(sample = Normalized_cum_sum)) +
     geom_qq() +
     geom_qq_line() +
     facet_wrap(~Densities, scales = "free_y") +
  theme_minimal()

shapiro.test(Normalized_sum$Normalized_cum_sum) 
#p = 0.623, overall normal, proceed with anova also given normality from NON normalized data

#ANOVA
Density_settlement_normal_aov <- aov(Normalized_cum_sum ~ Densities, data = Normalized_sum)
summary(Density_settlement_normal_aov) #p = 0.000431
plot(Density_settlement_normal_aov)

#robust anova
t1way(Normalized_cum_sum ~ Densities, data = Normalized_sum) #default 20%
##p = 0.01
t1way(Normalized_cum_sum ~ Densities, data = Normalized_sum, tr = .1) #trim only 10%
##p = 0.006
med1way(Normalized_cum_sum ~ Densities, data = Normalized_sum)
##p = 0.008

Density_settlement_normal_posthoc <- Density_settlement_normal_aov %>% tukey_hsd()
Density_settlement_normal_posthoc #view results

Density_settlement_normal_posthoc <- Density_settlement_normal_posthoc %>% arrange(p.adj)
Density_settlement_normal_posthoc <- add_xy_position(test = Density_settlement_normal_posthoc,
  data = Normalized_sum %>% as_tibble(.name_repair = "unique"),  #unique column names
  formula = Normalized_cum_sum ~ Densities, fun = "mean", x = "Densities")

x_levels <- levels(factor(Normalized_sum$Densities))

x_map <- setNames(seq_along(x_levels), x_levels)

Density_settlement_normal_posthoc <- Density_settlement_normal_posthoc %>%
  mutate(xmin = x_map[group1], xmax = x_map[group2])

Density_settlement_normal_posthoc <- Density_settlement_normal_posthoc
Density_settlement_normal_posthoc$y.position[1] <- 630 
Density_settlement_normal_posthoc$y.position[2] <- 570

#this plot took so long to trouble shoot.... genuinely... so long... its a MIRACLE it works
ggboxplot(Normalized_sum, x = "Densities", y = "Normalized_cum_sum", 
          color = "black", shape = "Densities") + 
  stat_boxplot(geom = "errorbar", width = 0.1) +
  geom_jitter(aes(x = Densities, y = Normalized_cum_sum, 
                  color = Densities, shape = Densities),
              data = Normalized_sum, width = 0.2, alpha = 0.6, size = 2) +
  stat_pvalue_manual(Density_settlement_normal_posthoc, 
                     label = "p.adj.signif", hide.ns = TRUE) +
  stat_compare_means(method = "anova", label.y = 680, label.x = "1 larvae/mL") +
  labs(x = "Larval density", y = "Normalized cumulative settlement") +
  scale_shape_manual(values = c(15, 16, 17, 18)) +
  scale_color_manual(values = c("#92fe6a", "#30d5c8", "#c0c0ff", "#f07177"))  +
  coord_cartesian(ylim = c(0, 700)) +  #extend y-axis to fit signf brackets
  theme_bw() +
  theme(legend.position = "none", axis.title.x = element_blank())

report(Density_settlement_normal_aov)
```
A one-way ANOVA was performed to compare the effect of Density on cumulative larval settlement.A one-way ANOVA revealed that there was a statistically significant difference in cumulative settlement of larvae between at least two groups (F(3, 16) = 10.65, p = <0.01)). Tukey’s HSD Test for multiple comparisons found that the mean value of Length was statistically significantly different between the 1 larva/mL and the 5 and 15-20 groups.


```{r DENSITY and SURVIVORSHIP kruskal}
##Did different larval densities have different levels of survivorship to settlement?
##Given that: each group is in an 8L container and larval densities are approximate, not precise.

#density labels for plot
survivorship$Densities <- factor(
  survivorship$Density, 
  levels = c("1larva/ml", "2larva/ml", "5then2larva/ml", "Bucket"),
  labels = c("1 larvae/mL", "2 larvae/mL", "5 larvae/mL", "15-20 larvae/mL"))

#calculate survivorship as % rather than decimal
survivorship$Survivorship <- survivorship$Survivorship * 100

#descriptive stats
#by(survivorship$Survivorship, survivorship$Densities, stat.desc) 
##mean/sd:
#1 larva/ml = 13.11% / 8.4
#2 larva/ml = 6.3% / 2.9
#5 larva/ml = 3.4% / 1
#Bucket = 1.3% / 0.2

#variance test, levene test
#leveneTest(survivorship$Survivorship, survivorship$Densities, center = median) 
#non appropriate, p = 0.01

#checking normality within groups
ggplot(survivorship, aes(sample = Survivorship)) +
     geom_qq() +
     geom_qq_line() +
     facet_wrap(~Density, scales = "free_y")
shapiro.test(survivorship$Survivorship) #p = <0.01, non-normal

#shapiro.test by group
test <- survivorship %>% dplyr::select(Density, Survivorship) %>%
  filter(Density == "1larva/ml")
shapiro.test(test$Survivorship) #1 larva/ml p = 0.116

test <- survivorship %>% dplyr::select(Density, Survivorship) %>%
  filter(Density == "2larva/ml")
shapiro.test(test$Survivorship) #2 larva/ml p = 0.6345

test <- survivorship %>% dplyr::select(Density, Survivorship) %>%
  filter(Density == "5then2larva/ml")
shapiro.test(test$Survivorship) #5 then 2 larva/ml p = 0.035, NON NORMAL

test <- survivorship %>% dplyr::select(Density, Survivorship) %>%
  filter(Density == "Bucket")
shapiro.test(test$Survivorship) #bucket p = 0.611
##3/4 normal, do both KW and ANOVA to be robust

Settlement_survivorship_kruskal <- kruskal.test(Survivorship ~ Densities, data = survivorship)
Settlement_survivorship_kruskal #p = 0.002632

Settlement_survivorship_posthoc <- dunn_test(data = survivorship, Survivorship ~ Densities, p.adjust.method = "bonferroni")
Settlement_survivorship_posthoc

Settlement_survivorship_posthoc <- Settlement_survivorship_posthoc %>% arrange(p.adj)
Settlement_survivorship_posthoc <- Settlement_survivorship_posthoc %>% add_xy_position(x = "Density")

#super awesome plot
ggplot(survivorship, aes(x = Densities, y = Survivorship)) +
  stat_boxplot(geom = 'errorbar', width = 0.2) +
  geom_boxplot() +
  geom_jitter(aes(shape = Densities, colour = Densities)) +
  stat_pvalue_manual(Settlement_survivorship_posthoc, label = "p.adj.signif", 
                     hide.ns = TRUE, y.position = "y.position") +
  stat_compare_means(label.y = 31, label.x = "1 larva/mL") +
  scale_y_continuous(labels = function(x) paste0(x, "%")) +
  scale_shape_manual(values = c(15, 16, 17, 18)) +
  scale_color_manual(values = c("#92fe6a", "#30d5c8", "#c0c0ff", "#f07177"))  +
  theme_classic2() +
  labs(y = "Larval settlement rate (%)") +
  theme_bw() +
  theme(legend.position = "none", axis.title.x = element_blank()) 

Settlement_survivorship_kruskal
```
A Kruskal-Wallis non-parametric test was performed to compare the effect of larval density on percent survivorship through settlement and metamorphosis. The Kruskal-Wallis test revealed that there WAS a statistically significant difference in weight between at least two groups (KW Chi2(3) = 14.211, p = 0.002)). A non-parametric Dunn's Test for multiple comparisons found that the mean value of weight was significantly different between Bucket and 1 larva/mL (p = 0.002).There was no statistically significant difference between any of the other density treatments.


```{r DENSITY and SURVIVORSHIP aov}
Settlement_survivorship_aov <- aov(Survivorship ~ Densities, data = survivorship)
summary(Settlement_survivorship_aov) #p = 0.0145
plot(Settlement_survivorship_aov)

#robust anova
t1way(Survivorship ~ Density, data = survivorship) #default 20%
##p = 0.01022*
t1way(Survivorship ~ Density, data = survivorship, tr = .1) #trim only 10%
##p = 0.0052**
med1way(Survivorship ~ Density, data = survivorship)
##p = 0.008**

#post-hoc tests
Settlement_survivorship_posthoc_2 <- Settlement_survivorship_aov %>% 
  tukey_hsd()
Settlement_survivorship_posthoc_2 #view results

#work flow for publication-ready plot
Settlement_survivorship_posthoc_2 <- Settlement_survivorship_posthoc_2 %>% arrange(p.adj)
Settlement_survivorship_posthoc_2 <- add_xy_position(test = Settlement_survivorship_posthoc_2, data = survivorship %>% as_tibble(
  .name_repair = "unique"), formula = Survivorship ~ Densities, fun = "mean", x = "Densities")

x_levels <- levels(factor(survivorship$Densities))

x_map <- setNames(seq_along(x_levels), x_levels)

Settlement_survivorship_posthoc_2 <- Settlement_survivorship_posthoc_2 %>%
  mutate(xmin = x_map[group1], xmax = x_map[group2])

Settlement_survivorship_posthoc_2 <- Settlement_survivorship_posthoc_2 %>% 
  mutate(y.position = y.position + 40)

Settlement_survivorship_posthoc_2$y.position[1] <- 27 #specific y position of signf bracket

#anova results on plot
ggplot(survivorship, aes(x = Densities, y = Survivorship)) +
  stat_boxplot(geom = 'errorbar', width = 0.5) +
  geom_boxplot(aes(colour = "black")) +
  geom_jitter(aes(shape = Densities, colour = Densities)) +
  stat_pvalue_manual(Settlement_survivorship_posthoc_2, label = "p.adj.signif", 
                     hide.ns = TRUE) +
  stat_compare_means(method = "anova", label.y = 30) +
  scale_y_continuous(labels = function(x) paste0(x, "%")) +
  scale_shape_manual(values = c(15, 16, 17, 18)) +
  scale_fill_manual(values = c("#66c2a5", "#fc8d62", "#8da0cb", "#e78ac3")) +
  theme_bw() +
  labs(y = "% Larvae surviving to settlement") +
  theme(legend.position = "none", axis.title.x = element_blank())

report(Settlement_survivorship_aov)
```
A one-way ANOVA was performed to compare the effect of Density on cumulative larval settlement.A one-way ANOVA revealed that there was NOT a statistically significant difference in cumulative settlement of larvae between at least two groups (F(3, 16) = 2.97, p = 0.063)). Tukey’s HSD Test for multiple comparisons found that the mean value of Length was no statistically significantly difference between or within any of the density groups.

```{r DENSITY and SURVIVORSHIP POSTER PLOT}
density_survivorship_poster <- 
  ggboxplot(survivorship, x = "Densities", y = "Survivorship") +
  geom_jitter(aes(colour = Densities, shape = Densities), size = 2) +
  stat_boxplot(geom = 'errorbar', width = 0.2) +
  stat_pvalue_manual(Settlement_survivorship_posthoc_2, label = "p.adj.signif", 
                     hide.ns = TRUE, y.position = 27) +
  stat_compare_means(method = "anova", label.y = 30) +
  scale_y_continuous(labels = function(x) paste0(x, "%")) +
  scale_shape_manual(values = c(15, 16, 17, 18)) +
  scale_color_manual(values = c("#006b3c", "#D2BAAD", "#9F8FAC", "#f3be77")) +
  labs(y = "Percent (%) survival to settlement") +
  theme(legend.position = "none", axis.title.x = element_blank())

density_survivorship_poster +
  theme(
    panel.background = element_rect(fill='transparent'), #transparent panel bg
    plot.background = element_rect(fill='transparent', color=NA), #transparent plot bg
    panel.grid.major = element_blank(), #remove major gridlines
    panel.grid.minor = element_blank(), #remove minor gridlines
    legend.background = element_rect(fill='transparent'), #transparent legend bg
    legend.box.background = element_rect(fill='transparent') #transparent legend panel
  )

ggsave('density_survivorship.png', density_survivorship_poster, bg='transparent')
```

```{r}
density_survivorship_poster <- 
  ggboxplot(
    survivorship,
    x = "Densities",
    y = "Survivorship",
    color = "white",   # ← white box outline
    fill  = NA         # ← transparent IQR
  ) +
  geom_jitter(
    aes(colour = Densities, shape = Densities),
    size = 2
  ) +
  stat_boxplot(
    geom = "errorbar",
    width = 0.2,
    colour = "white"
  ) +
  stat_pvalue_manual(
    Settlement_survivorship_posthoc_2,
    label = "p.adj.signif",
    hide.ns = TRUE,
    y.position = 27,
    color = "white"
  ) +
  stat_compare_means(
    method = "anova",
    label.y = 30,
    color = "white"
  ) +
  scale_y_continuous(labels = function(x) paste0(x, "%")) +
  scale_shape_manual(values = c(15, 16, 17, 18)) +
  scale_color_manual(values = c("#006b3c", "#D2BAAD", "#9F8FAC", "#f3be77")) +
  labs(
    y = "Percent (%) survival to settlement",
    x = NULL
  ) +
  theme(
    panel.background = element_rect(fill = "transparent", colour = NA),
    plot.background  = element_rect(fill = "transparent", colour = NA),
    legend.background = element_rect(fill = "transparent"),
    legend.box.background = element_rect(fill = "transparent"),

    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),

    text = element_text(color = "white"),
    axis.text = element_text(color = "white"),
    axis.title = element_text(color = "white"),

    legend.position = "none"
  )


ggsave(
  filename = "density_survivorship_transparent.png",
  plot = density_survivorship_poster,
  bg = "transparent",
  dpi = 300
)
```


```{r GAM of SETTLEMENT by DENSITY over DPF non normalized}
##Do daily settlement rates differ by density groups over time?

Settlement_dpf <- settlement_raw_long %>%
  mutate(Density = as.factor(Density),
         Group = as.factor(Group),
         Settlers = as.integer(Settlers), 
         DPF_scaled = scale(DPF)) 

#GAM 
Settler_gam_model <- gam(Settlers ~ s(DPF, by = Density) + Density, 
                 family = nb(), data = Settlement_dpf)
summary(Settler_gam_model)

#new df for settlement prediction across DPF range for each density group
new_data <- Settlement_dpf %>% group_by(Density) %>%
  reframe(DPF = seq(min(DPF), max(DPF), length.out = 100)) 

#adding the predicted fits on a response scale
new_data$Predicted <- predict(Settler_gam_model, newdata = new_data, type = "response")

#plot it
ggplot(Settlement_dpf, aes(x = DPF, y = Settlers, color = Density)) +
  geom_point(alpha = 0.4) +
  geom_line(data = new_data, aes(x = DPF, y = Predicted, color = Density), size = 1.2) +
  labs(title = "GAM-Predicted Larval Settlement Trajectories by Density",
       y = "Daily Settler Count (Observed + Predicted)",
       x = "Days Post Fertilization") +
  theme_minimal(base_size = 14)

#plot GAM smooths apart from graphed data
plot(Settler_gam_model, pages = 1, se = TRUE, shade = TRUE)
```



```{r GAM of SETTLEMENT by DENSITY over DPF normalized}
##Do daily settlement rates differ by density groups over time? (with normalized data from density group replicates)

#Settlers by density normalized according to # of replicates in density groups
replicate_counts <- c("1 Larvae/mL" = 8, #replicates in each density treatment
                      "2 larvae/mL" = 4, 
                      "5 larvae/mL" = 4, 
                      "15-20 larvae/mL" = 4)

#working df
Settlement_dpf_normalized <- settlement_raw_long %>%
  group_by(Densities, DPF) %>% 
  arrange(Densities, DPF) %>% 
  group_by(Densities) %>% 
  mutate(Normalized_settlers = Settlers / replicate_counts[Densities], #div by # replicates
         Density = as.factor(Density),
         Group = as.factor(Group),
         Settlers = as.integer(Settlers))

#GAM 
Settlement_dpf_normalized$Densities <- as.factor(Settlement_dpf_normalized$Densities)
Settler_normalized_gam <- gam(Normalized_settlers ~ s(DPF, by = Densities) + Densities, 
                 family = nb(), data = Settlement_dpf_normalized)
summary(Settler_normalized_gam)
coef(Settler_normalized_gam)

#visualizing GAM outputs:
plot(Settler_normalized_gam, shade = TRUE) #4 splines output
abline(h = 0, lty = 'dashed') #adding line at zero

#plot predictions using marginal effects
#aggregated across density treatments
plot_predictions(Settler_normalized_gam, condition = 'DPF', type = 'link') +
  labs(y = "Linear predictor (link scale)", title = "Average effects of DPF", subtitle = "Aggregated across larval densities") +
  scale_x_continuous(breaks = seq(0, 150, by = 10)) +
  theme_clean()

#by density treatment
plot_predictions(Settler_normalized_gam, condition = c('DPF', 'Densities'), type = 'link') +
  labs(y = "Linear predictor (link scale)", title = "Average effects of DPF", subtitle = "Aggregated by larval density") +
  scale_x_continuous(breaks = seq(0, 150, by = 10)) +
  scale_color_manual(values = c("#92fe6a", "#30d5c8", "#c0c0ff", "#f07177")) +
  theme_clean()

#plotting GAM effect on response scale
#aggregated across larval densities
plot_predictions(Settler_normalized_gam, condition = 'DPF', type = 'response') +
  labs(y = "Expected response", title = "Average smooth effect of DPF on Settlement",
       subtitle = "Aggregated across all larval densities") +
  scale_x_continuous(breaks = seq(0, 150, by = 10)) +
  theme_classic2()

#by density treatment (without obvs)
plot_predictions(Settler_normalized_gam, condition = c('DPF', 'Densities'), type = 'response') +
  labs(y = "Expected response", title = "Expected effect of DPF on settlement",
       subtitle = "Aggregated by larval density") +
  scale_color_manual(values = c("#92fe6a", "#30d5c8", "#c0c0ff", "#f07177")) +
  scale_x_continuous(breaks = seq(0, 150, by = 10)) +
  theme_bw() +
  theme(legend.position = "none")

#plot with regular observations on a smooth curve
ggplot(Settlement_dpf_normalized, aes(x = DPF, y = Normalized_settlers, 
                                      colour = Densities)) +
  geom_smooth(aes(fill = Densities), alpha = 0.1, size = 0.5) +
   labs(y = "Observed response", title = "Observed effect of DPF on settlement",
       subtitle = "Aggregated by larval density") +
  scale_color_manual(values = c("#92fe6a", "#30d5c8", "#c0c0ff", "#f07177")) +
  scale_fill_manual(values = c("#92fe6a", "#30d5c8", "#c0c0ff", "#f07177")) +
  scale_x_continuous(breaks = seq(0, 150, by = 10)) +
  coord_cartesian(ylim = c(0,40)) + 
  theme_bw()

#by density treatment WITH obvs barcoded
plot_predictions(Settler_normalized_gam, condition = c('DPF', 'Densities'), type = 'response', points = 0.5, rug = TRUE) +
  labs(y = "Expected response", title = "Average smooth effect of DPF on settlement",
       subtitle = "Aggregated within larval density treatments") +
  theme_bw()

#new df for settlement prediction across DPF range for each density group
Normalized_settlement_prediction <- Settlement_dpf_normalized %>% 
  group_by(Densities) %>%
  reframe(DPF = seq(min(DPF), max(DPF), length.out = 100)) 

#adding the predicted fits on a response scale
Normalized_settlement_prediction$Predicted <- predict(
  Settler_normalized_gam, newdata = Normalized_settlement_prediction, type = "response")

#plot it: GAM-predicted larval settlement trajectories by density
ggplot(Settlement_dpf_normalized, aes(x = DPF, y = Normalized_settlers, 
                                      color = Densities)) +
  geom_point(alpha = 0.4) + #make actual obvs more translucent
  geom_line(data = Normalized_settlement_prediction, #plotting GAM predictions
            aes(x = DPF, y = Predicted, color = Densities), size = 1) +
  labs(y = "Normalized daily settler count (observed + predicted)",
       x = "Days post fertilization") +
  theme_clean(base_size = 14) 

#plot GAM smooths apart from graphed data
plot(Settler_normalized_gam, pages = 1, se = TRUE, shade = TRUE)

#simulating from fitted GAM
#using linear predictor matrix Xlp
xp <- predict.gam(Settler_normalized_gam, type = 'lpmatrix')
dim(xp) #dimensions of prediction- 40 columns of basis functions for smooth terms in GAM
colnames(xp) #see smooth terms of spline
```



## Survivorship from settlement to juvenilehood
```{r plotting jitter and boxplot}
#Density labels for nicer plots
juvenile_settlement$Densities <- factor(
  juvenile_settlement$Density, 
  levels = c("1larva/ml", "2larva/ml", "5then2larva/ml", "Bucket"),
  labels = c("1 Larvae/mL", "2 larvae/mL", "5 larvae/mL", "15-20 larvae/mL"))

#calculative settlement rate (#juveniles in feb 25 / #settlers in july 24)
juvenile_settlement <- juvenile_settlement %>% group_by(Densities) %>%
  mutate(Survival_rate = Juvenile_total / Larval_total)

#mean, sd, summary statis for survival
Survival_summary <- juvenile_settlement %>% group_by(Densities) %>%
  summarise(mean_survival = mean(Survival_rate),
            sd = sd(Survival_rate),
            n = n())

#summary stats labels for graph
Survival_summary <- Survival_summary %>% mutate(label = paste0(
      round(mean_survival * 100, 2), "% ± ",
      round(sd * 100, 2), "%"))

#horizontal jitter with error bars
ggplot(data = juvenile_settlement, 
       aes(x = Densities, y = Survival_rate, color = Densities)) +
  geom_jitter(size = 2, alpha = 0.7, aes(shape = Densities)) +
  stat_summary(fun = mean, geom = "crossbar", width = 0.2, size = 0.7,
               color = "black", fatten = 1) +
  geom_errorbar(data = Survival_summary, 
                aes(x = Densities, ymin = mean_survival - sd, ymax = mean_survival + sd), 
                width = 0.2, color = "black", inherit.aes = FALSE) +
  scale_y_continuous(labels = scales::percent) +
  scale_shape_manual(values = c(15, 16, 17, 18)) +
  scale_fill_manual(values = c("#66c2a5", "#fc8d62", "#8da0cb", "#e78ac3")) +
  labs(y = "% Settlers surviving to 8 months post-settlement") +
  theme_bw() +
  theme(legend.position = "none", axis.title.x = element_blank())

#geom text with summary stats if you want to add them
#geom_text(data = Survival_summary, aes(x = Densities, y = mean_survival + sd + 0.01, label = label),inherit.aes = FALSE, size = 3.5, color = "black") +

#box plot, if that's more your flavor
ggplot(juvenile_settlement, aes(x = Densities, y = Survival_rate, fill = Densities)) +
  geom_boxplot() +
  stat_boxplot(geom ='errorbar', width = 0.3) +
  geom_jitter(aes(shape = Densities)) +
  scale_y_continuous(labels = scales::percent) +
  scale_shape_manual(values = c(15, 16, 17, 18)) +
  scale_fill_manual(values = c("#66c2a5", "#fc8d62", "#8da0cb", "#e78ac3")) +
  theme_bw() +
  labs(y = "% Settlers surviving to 8 months post-settlement") +
  theme(legend.position = "none", axis.title.x = element_blank())
```


```{r jitter plot with point labels}
set.seed(123) #reproducibility
jitter_width <- 0.25 #spacing out jitters for label readability

#add numeric x-position and jitter to your data so jitter points are not random
juvenile_settlement <- juvenile_settlement %>%
  mutate(Density_num = as.numeric(Densities),
         jittered_x = Density_num + runif(n(), -jitter_width, jitter_width))

Survival_summary <- Survival_summary %>%
  mutate(Density_num = as.numeric(Densities))

#plot
ggplot() + 
  geom_point(data = juvenile_settlement,
             aes(x = jittered_x, y = Survival_rate, color = Densities, shape = Densities),
             size = 2, alpha = 0.7) +
  geom_text(data = juvenile_settlement, 
            aes(x = jittered_x, y = Survival_rate, label = Group), 
            size = 3, vjust = -1, show.legend = FALSE) +
 stat_summary(data = juvenile_settlement, aes(x = Density_num, y = Survival_rate),
               fun = mean, geom = "crossbar", width = 0.2, size = 0.7, 
              color = "black", fatten = 1) +
 geom_errorbar(data = Survival_summary, 
               aes(x = Density_num, ymin = mean_survival - sd, ymax = mean_survival + sd),
                width = 0.2, color = "black") +
  scale_y_continuous(labels = scales::percent) +
  scale_shape_manual(values = c(15, 16, 17, 18)) +
  scale_color_manual(values = c("#66c2a5", "#fc8d62", "#8da0cb", "#e78ac3")) +
  scale_x_continuous(breaks = 1:4, labels = levels(juvenile_settlement$Densities)) +
  labs(y = "% Settlers surviving to 8 months post-settlement") +
  theme_classic() +
  theme(legend.position = "none", axis.title.x = element_blank())

#reimport for cleaning purposes
juvenile_settlement <- read.csv("Juvenile settlment survival.csv")

ggplot(juvenile_settlement, aes(x = Settlement_survival, y = Juvenile_survival, colour = Density)) +
  geom_point() +
  geom_text(aes(label = Group)) +
  scale_y_continuous(labels = scales::percent) +
  scale_x_continuous(labels = scales::percent) +
  scale_color_manual(values = c("#66c2a5", "#fc8d62", "#8da0cb", "#e78ac3")) +
  theme_classic2()

#plot and colour by diet? 
ggplot(juvenile_settlement, aes(x = Settlement_survival, y = Juvenile_survival, colour = Diet)) +
  geom_point() +
  scale_y_continuous(labels = scales::percent) +
  scale_x_continuous(labels = scales::percent) +
  theme_classic2()
```


```{r linear regression JUVENILE and SETTLER SURVIVAL}
Settlement_survival_lm <- lm(formula = Juvenile_survival ~ Settlement_survival, data = juvenile_settlement)

Settlement_survival_lm
summary(Settlement_survival_lm)
sqrt(0.1136) #pearson coefficient is = 0.337046
AIC(Settlement_survival_lm) #AIC = -163.299

model.diag.metrics <- augment(Settlement_survival_lm)
head(model.diag.metrics)

#plotted residuals error (red) between observed values and fitted regression line
ggplot(model.diag.metrics, aes(Settlement_survival, Juvenile_survival)) +
  geom_point() +
  stat_smooth(method = lm, se = FALSE) +
  geom_segment(aes(xend = Settlement_survival, yend = .fitted), 
               color = "red", size = 0.3) +
  theme_classic2()
```


```{r pct SETTLEMENT over DPF}
##At what DPF, will x% of larva in x density treatment be settled? (predicted)

#prep data
Settlement_percentile <- Settlement_dpf_normalized %>% group_by(Group) %>% 
  group_by(Density) %>%
  arrange(DPF) %>%
  mutate(Cumulative_settlement = cumsum(Normalized_settlers),
         Percent_settled = Cumulative_settlement / max(Cumulative_settlement, na.rm = TRUE)) %>% ungroup()

#average percent settled across density treatments
Density_settlement_percentile <- Settlement_percentile %>% group_by(Density, DPF) %>%
  summarise(mean_percent_settled = mean(Percent_settled, na.rm = TRUE), .groups = 'drop')
  
#New gam with percentile data
Settler_percentile_gam <- gam(mean_percent_settled ~ s(DPF, by = Density) + Density,
                              family = nb(), data = Density_settlement_percentile)
summary(Settler_percentile_gam)
coef(Settler_percentile_gam)

#using gam to find the DPF at x% cumulative settlement
##generate predicted values from the GAM across DPF

#prediction grid, vary by 0.1 per day
pred_grid <- expand.grid(DPF = seq(min(settlement_raw_long$DPF),
                                   max(settlement_raw_long$DPF), by = 0.1),
                         Density = unique(Density_settlement_percentile$Density))

#predict values in prediction grid
pred_grid$Predicted_percent <- predict(Settler_percentile_gam, 
                                       newdata = pred_grid, type = "response")

#function, getting DPF at threshold
get_dpf_at_threshold <- function(df, threshold) {
  df %>% group_by(Density) %>%
    dplyr::filter(Predicted_percent >= threshold) %>%
    slice_min(DPF, n = 1) %>%
    mutate(Percentile = paste0(threshold * 100, "%")) %>%
    dplyr::select(Density, Percentile, DPF)
}

#retrieve the DPF for each percentile we want, i'm choosing 25%, 50%, 75%, and 90%
percentiles <- bind_rows(get_dpf_at_threshold(pred_grid, 0.25),
  get_dpf_at_threshold(pred_grid, 0.5),
  get_dpf_at_threshold(pred_grid, 0.9),
  get_dpf_at_threshold(pred_grid, 1.0))

print(percentiles) #retrieving DPF where density x will be at x%

percentiles$Percentile <- factor(percentiles$Percentile,
                                 levels = c("25%", "50%", "90%", "100%"))

#plot predictions with estimated DPF 
ggplot(pred_grid, aes(x = DPF, y = Predicted_percent, colour = Density)) +
  geom_line() +
  geom_vline(data = percentiles, aes(xintercept = DPF, colour = Density), 
             linetype = "dashed") +
  facet_wrap(~ Percentile, nrow = 2, ncol = 2) +
  scale_y_continuous(labels = scales::percent) +
  scale_color_manual(values = c("#92fe6a", "#30d5c8", "#c0c0ff", "#f07177")) +
  labs(y = "Predicted Settlement (%)", x = "Days Post Fertilization (DPF)") +
  theme_bw()
```




```{r stat anal for all that stuff}
## Did the larval density impact the survival of new settlers to juveniles 8 months post-settlement? 
juvenile_settlement$Densities <- factor(
  juvenile_settlement$Density, 
  levels = c("1larva/ml", "2larva/ml", "5then2larva/ml", "Bucket"),
  labels = c("1 Larvae/mL", "2 larvae/mL", "5 larvae/mL", "15-20 larvae/mL"))

#mean, sd, summary stats for survival for graph below :)
Survival_summary <- juvenile_settlement %>% group_by(Densities) %>%
  summarise(mean_survival = mean(Juvenile_survival),
            sd = sd(Juvenile_survival),
            n = n())

#visually checking the differences in mean survival %
ggplot(Survival_summary, aes(x = Densities, y = mean_survival, group = 1)) +
  geom_line() +
  geom_point() +
   geom_errorbar(width = 0.1, colour = "red", 
                aes(ymin = mean_survival - sd, ymax = mean_survival + sd)) +
  scale_y_continuous(labels = scales::percent) +
  labs(x = "Larval Density", y = "Mean % survival to juvenile-hood") +
  theme_classic2()

#descriptive stats
by(juvenile_settlement$Juvenile_survival, juvenile_settlement$Densities, stat.desc)

#checking variances
leveneTest(juvenile_settlement$Juvenile_survival, juvenile_settlement$Densities, center = median) #appropriate, p = 0.2

#checking normality
ggplot(juvenile_settlement, aes(sample = Juvenile_survival)) +
     geom_qq() +
     geom_qq_line() +
     facet_wrap(~Densities, scales = "free_y")
shapiro.test(juvenile_settlement$Juvenile_survival) #p = <0.01, overall NON NORMAL

#shapiro.test by group
test <- juvenile_settlement %>% dplyr::select(Density, Juvenile_survival) %>%
  filter(Density == "1larva/ml")
shapiro.test(test$Juvenile_survival) #1 larva/ml p = 0.29

test <- juvenile_settlement %>% dplyr::select(Density, Juvenile_survival) %>%
  filter(Density == "2larva/ml")
shapiro.test(test$Juvenile_survival) #2 larva/ml p = 0.7873

test <- juvenile_settlement %>% dplyr::select(Density, Juvenile_survival) %>%
  filter(Density == "5then2larva/ml")
shapiro.test(test$Juvenile_survival) #5 then 2 larva/ml p = 0.5533

test <- juvenile_settlement %>% dplyr::select(Density, Juvenile_survival) %>%
  filter(Density == "Bucket")
shapiro.test(test$Juvenile_survival) #bucket p = 0.3371
##3/4 appropriate, between group non appropriate, proceeding with both KW and aNOVa to be robust

##KW
#kruskal-wallis 
Settlement_survival_kruskal <- kruskal.test(Juvenile_survival ~ Densities, data = juvenile_settlement)
Settlement_survival_kruskal #p = 0.747

#post-hoc tests
Settlement_survival_kw_posthoc <- dunn_test(data = juvenile_settlement, Juvenile_survival ~ Densities, p.adjust.method = "bonferroni")
Settlement_survival_kw_posthoc
Settlement_survival_kw_posthoc <- Settlement_survival_kw_posthoc %>% arrange(p.adj)
Settlement_survival_kw_posthoc <- Settlement_survival_kw_posthoc %>% add_xy_position(x = "Density")



ggboxplot(juvenile_settlement, x = "Densities", y = "Juvenile_survival", color = "black") +
  geom_jitter(aes(colour = Densities, shape = Densities), size = 2) + 
  stat_pvalue_manual(Settlement_survival_kw_posthoc, label = "p.adj.signif", 
                     hide.ns = TRUE) +
  stat_boxplot(geom ='errorbar', width = 0.2) +
  stat_compare_means(label.y = 0.021, label.x = "1 Larvae/mL") +
  labs(x = "Larval density", y = "Larvae surviving to 8-months post-settlement (%)") +
  coord_cartesian(ylim = c(0, 0.021)) + 
  scale_y_continuous(labels = scales::percent) +
  scale_shape_manual(values = c(15, 16, 17, 18)) +
  scale_color_manual(values = c("#92fe6a", "#68ebff", "#936bff", "#d4646c")) +
  theme_bw() +
  theme(legend.position = "none", axis.title.x = element_blank()) 

#ANOVA
Settlement_survival_aov <- aov(Juvenile_survival ~ Densities, data = juvenile_settlement)
summary(Settlement_survival_aov) #p = 0.562
plot(Settlement_survival_aov)

#robust anova
t1way(Juvenile_survival ~ Densities, data = juvenile_settlement) #default 20%
##p = 0.0102
t1way(Juvenile_survival ~ Densities, data = juvenile_settlement, tr = .1) #trim only 10%
##p = 0.005
med1way(Juvenile_survival ~ Densities, data = juvenile_settlement)
##p = 0.008

#post-hoc tests
Settlement_survival_posthoc <- Settlement_survival_aov %>% tukey_hsd() #tukey's test
Settlement_survival_posthoc #view results, of course bucket and 1larva are signf....

#work flow for publication-ready plot
Settlement_survival_posthoc <- Settlement_survival_posthoc %>% arrange(p.adj) #arrange
Settlement_survival_posthoc <- add_xy_position(test = Settlement_survival_posthoc,
  data = juvenile_settlement %>% as_tibble(.name_repair = "unique"),  #unique column names
  formula = Juvenile_survival ~ Densities, fun = "mean", x = "Densities")

#factor levels in the order used in the plot
x_levels <- levels(factor(juvenile_settlement$Densities))

#vector to convert group names to x-axis positions
x_map <- setNames(seq_along(x_levels), x_levels)

#map to assign numeric values to xmin and xmax
Settlement_survival_posthoc <- Settlement_survival_posthoc %>%
  mutate(xmin = x_map[group1], xmax = x_map[group2])

#adjusting y.position to fit the signf brackets
Settlement_survival_posthoc <- Settlement_survival_posthoc %>% 
  mutate(y.position = y.position)
Settlement_survival_posthoc$y.position[1] <- 0.30 #specific y position of 2nd signf bracket

#this plot took so long to trouble shoot.... genuinely... so long... its a MIRACLE it works
ggboxplot(juvenile_settlement, x = "Densities", y = "Juvenile_survival", 
          color = "black") + 
  geom_jitter(aes(colour = Densities, shape = Densities), size = 2) +
  stat_boxplot(geom ='errorbar', width = 0.2) +
  stat_pvalue_manual(Settlement_survival_posthoc, label = "p.adj.signif", hide.ns = TRUE) +
  stat_compare_means(method = "anova", label.y = 0.02, label.x = "1 Larvae/mL") +
  labs(x = "Larval density", y = "% Larvae surviving to settlement") +
  coord_cartesian(ylim = c(0, 0.02)) + 
  scale_y_continuous(labels = scales::percent) +
  scale_shape_manual(values = c(15, 16, 17, 18)) +
  scale_color_manual(values = c("#92fe6a", "#68ebff", "#936bff", "#d4646c")) +
  theme(legend.position = "none", axis.title.x = element_blank())

report(Settlement_survival_aov)
```
A one-way ANOVA was performed to compare the effect of Density on cumulative larval settlement.A one-way ANOVA revealed that there was NOT a statistically significant difference in cumulative settlement of larvae between at least two groups (F(3, 16) = 2.97, p = 0.063)). Tukey’s HSD Test for multiple comparisons found that the mean value of Length was no statistically significantly difference between or within any of the density groups. 


## Larval Growth -- from ImageJ photos
```{r larval growth}
#first pass visualization
plot(data = larval_growth, Size ~ DPF)

#ordering densities
larval_growth$Density <- 
  factor(larval_growth$Density, levels = c("1 larvae/mL",
                                           "2 larvae/mL",
                                           "5 larvae/mL",
                                           "15-20 larvae/mL"))

#working df
Larval_growth <- larval_growth %>% arrange(desc(Density)) %>%
  group_by(DPF, Density) %>%
  summarise(mean_size = mean(Size))

#first pass plot
ggplot(Larval_growth, aes(x = DPF, y = mean_size, colour = Density)) +
  geom_point() +
  geom_line() +
  labs(y= "Average size (µm)", x = "Days post fertilization (DPF)") +
  theme_clean()

```

```{r larval growth POSTER PLOT}

Larval_growth <- Larval_growth %>% filter(!DPF == 42)

larval_growth_poster <- 
  ggplot(Larval_growth, aes(x = DPF, y = mean_size, colour = Density)) +
  geom_point() +
  geom_line() +
  scale_color_manual(values = c("#006b3c", "#D2BAAD", "#9F8FAC", "#f3be77")) +
  labs(y= "Average size (µm)", x = "Days post fertilization (DPF)",
       colour = "Larval Density") +
  theme_classic() +
  theme(legend.position = 'top')

ggsave('larval_growth.png', larval_growth_poster, bg='transparent')
```

```{r}
Larval_growth <- Larval_growth %>% 
  filter(DPF != 42)

larval_growth_poster <- 
  ggplot(Larval_growth, aes(x = DPF, y = mean_size, colour = Density)) +
  geom_point() +
  geom_line() +
  scale_color_manual(values = c("#006b3c", "#D2BAAD", "#9F8FAC", "#f3be77")) +
  labs(
    y = "Average size (µm)",
    x = "Days post fertilization (DPF)",
    colour = "Larval Density"
  ) +
  theme(
    # transparent backgrounds
    panel.background = element_rect(fill = "transparent", colour = NA),
    plot.background  = element_rect(fill = "transparent", colour = NA),
    legend.background = element_rect(fill = "transparent"),
    legend.box.background = element_rect(fill = "transparent"),

    # no gridlines
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),

    # white text everywhere
    text = element_text(color = "white"),
    axis.text = element_text(color = "white"),
    axis.title = element_text(color = "white"),
    legend.text = element_text(color = "white"),
    legend.title = element_text(color = "white"),

    # white axes and ticks
    axis.line = element_line(color = "white"),
    axis.ticks = element_line(color = "white"),

    legend.position = "top"
  )

ggsave(
  filename = "larval_growth_transparent.png",
  plot = larval_growth_poster,
  bg = "transparent",
  dpi = 300
)
```

